### 2021-09-24 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@0bc3b6c27c90321866495cce141c996ca59614b5 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Dissimilarity of species interaction networks:
quantifying the effect of turnover and rewiring}
\vskip 5em

%
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
Sciences

\textsuperscript{‡}\,These authors contributed equally to the work\\

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill
Despite having established its usefulness in the last ten years, the
decomposition of ecological networks in components allowing to measure
their \(eta\)-diversity retains some methodological ambiguities.
Notably, how to quantify the relative effect of mechanisms tied to
interaction rewiring \emph{vs.} species turnover has been interpreted
differently by different authors. In this contribution, I present
mathematical arguments and numerical experiments that should (i)
establish that the decomposition of networks as it is currently done is
indeed fit for purpose, and (ii) provide guidelines to interpret the
values of the components tied to turnover and rewiring.

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

Ecological networks are variable both in time and space (Poisot \emph{et
al.} 2015; Trøjelsgaard \& Olesen 2016) - this variability motivated the
emergence of methodology to compare ecological networks, in a way that
meshes with the usual approaches of comparison of ecological
communities, \emph{i.e.} \(\beta\)-diversity; although the definiton of
\(\beta\)-diversity is a contentious topic amongst community ecologists
(see \emph{e.g.} Tuomisto 2010), the need to understand network
variability is motivated by the fact that species that make up the
networks do not react to their environment in the same way, and
therefore the \(\beta\)-diversity of networks may behave in complex
ways.

Poisot \emph{et al.} (2012) and Canard \emph{et al.} (2014) have
suggested an approach to \(\beta\)-diversity for ecological networks
which is based on the comparison of shared and unique links among
species, and differentiate this sharing of links between common and
unique species. This framework can be summarized as
\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that overall
network dissimilarity (\(\beta_{wn}\)) has a component that can be
calculated directly from the dissimilarity of interactions between
shared species (\(\beta_{os}\)), and a component that cannot, the later
originating in unique species introducing their unique interactions
(\(\beta_{st}\)). This approach has been widely adopted since its
publication, with recent examples using it to understand the effect of
fire on pollination systems (Baronio \emph{et al.} 2021); the impact of
rewiring on spatio-temporal network dynamics (Campos-Moreno \emph{et
al.} 2021); the effects of farming on rural and urban landscapes on
species interactions (Olsson \emph{et al.} 2021); and as a tool to
estimate the sampling completeness of networks (Souza \emph{et al.}
2021). It has, similarly, received a number of extensions, including the
ability to account for interaction strength (Magrach \emph{et al.}
2017), the ability to handle probabilistic ecological networks (Poisot
\emph{et al.} 2016), and the integration into the Local Contribution to
Beta Diversity (Legendre \& De Cáceres 2013) approach to understand how
environment changes drive network dissimilarity (Poisot \emph{et al.}
2017).

In a recent contribution, Fründ (2021) argues that the calculation of
network dissimilarity terms as outlined by Poisot \emph{et al.} (2012)
is incorrect, as it can lead to over-estimating the role of interactions
between shared species in a network (``rewiring''), and therefore
underestimate the importance of species turnover across networks. Here,
I present a more thorough justification of the methodological choices
for the Poisot \emph{et al.} (2012) method, explain how information
about species turnover can be extracted from its decomposition, and
conduct numerical experiments to guide the interpretation of the
\(\beta\)-diversity values thus obtained. These numerical experiments
establish three core facts. First, the decomposition responds to the
correct sources of network variation; second, the decomposition
adequately captures the relative roles of species turnover and
interaction rewiring; finally, the decomposition adequately captures the
role of turnover vs.~non-turnover (like changes in connectance)
processes. Although the alternative normalization suggested by Fründ
(2021) is not without its uses, which I discuss in conclusion, it is
inadequate as a network \(\beta\)-diversity measurement, as it
introduces many confounding elements that make the interpretation of the
results more difficult, and should likely not be used as a default.

\hypertarget{partitioning-network-dissimilarity}{%
\subsection{Partitioning network
dissimilarity}\label{partitioning-network-dissimilarity}}

The approach to quantifying the difference between pairs of networks
established in Poisot \emph{et al.} (2012) is a simple extension of the
overall method by Koleff \emph{et al.} (2003) for species dissimilarity
baed on presence-absence data. The objects to compare, \(X_1\) and
\(X_2\), are partitioned into three values, \(a = |X_1 \cup X_2|\),
\(b = |X_2 \setminus X_1|\), and \(c = |X_1 \setminus X_2|\), where
\(|x|\) is the cardinality of set \(x\), and \(\setminus\) is the set
substraction operation. In the perspective of species composition
comparison, \(X_1\) and \(X_2\) are the sets of species in either
community, so that if \(X_1 = \{x, y, z\}\) and
\(X_2 = \{v, w, x, y\}\), we have \(X_1 \cup X_2 = \{v, w, x, y, z\}\),
\(X_1 \cap X_2 = \{x, y\}\), \(X_2 \setminus X_1 = \{v, w\}\), and
\(X_1 \setminus X_2 = \{z\}\). The core message of Koleff \emph{et al.}
(2003) is that the overwheling majority of measures of
\(\beta\)-diversity can be re-expressed as functions that operate on the
cardinality (number of elements) of these sets.

\hypertarget{re-expressing-networks-as-sets}{%
\subsubsection{Re-expressing networks as
sets}\label{re-expressing-networks-as-sets}}

Applying this framework to networks requires a few additional
definitions. Although ecologists tend to think of networks as their
adjacency matrix, this representation is far from optimal to get a solid
understanding of which elements should be counted as part of which set
when measuring network dissimilarity. For this reason, we need fall back
on the definition of a graph as a pair of sets, wherein
\(\mathcal{G} = (V, E)\). These two components \(V\) and \(E\) represent
vertices (nodes, species) and edges (interactions), where \(V\) is
specifically a set containing the vertices \(\mathcal{G}\), and \(E\) is
a set of ordered pairs, in which every pair is composed of two elements
of \(V\); an element \(\{i,j\}\) in \(E\) indicates that there is an
interaction \emph{from} species \(i\) to species \(j\) in the network
\(\mathcal{G}\).

In the context of networks comparison (assuming the networks to compare
are \(\mathcal{M}\) and \(\mathcal{N}\)), we can further decompose the
contents of these sets as

\[\mathcal{M} = (V_c \cup V_m, E_c \cup E_{sm} \cup E_{um}) \,,\]

and

\[\mathcal{M} = (V_c \cup V_n, E_c \cup E_{sn} \cup E_{un}) \,,\]

where \(V_c\) is the set of shared species, \(V_k\) are the species
belonging only to network \(k\), \(E_c\) are the shared edges, and
\(E_{sk}\) and \(E_{uk}\) are the interactions unique to \(k\)
involving, respectively, only species in \(V_c\), and at least one
species from \(V_k\).

\hypertarget{defining-the-partitions-from-networks-as-sets}{%
\subsubsection{Defining the partitions from networks as
sets}\label{defining-the-partitions-from-networks-as-sets}}

The metaweb (Dunne 2006), which is to say the entire regional species
pool and their interaction, can be defined as
\(\mathcal{M} \cup \mathcal{N}\) (this operation is commutative), which
is to say

\[\mathcal{M} \cup \mathcal{N} = (V_c \cup V_m \cup V_n, E_c \cup E_{sm} \cup E_{um} \cup E_{sn} \cup E_{un}) \,.\]

This operation gives us an equivalent to \(\gamma\)-diversity for
networks, in that the set of vertices contains \emph{all} species from
the two networks, and the set of edges contains \emph{all} the
interactions between these species. If, further, we make the usual
assumption that only species with at least one interaction are present
in the set of vertices, then all elements of the set of vertices are
present at least once in the set of edges, and the set of vertices can
be entire reconstructed from the set of edges. Although measures of
network \(\beta\)-diversity operate on interactions (not species), this
property is maintained at every decomposition we will describe next.

We can similarly define the intersection (similarly commutative) of two
networks:

\[\mathcal{M} \cap \mathcal{N} = (V_c, E_c)\,.\]

The decomposition of \(\beta\)-diversity from Poisot \emph{et al.}
(2012) uses these components to measure \(\beta_{os}\) (the interaction
dissimilarity between shared species, which Fründ (2021) terms
``rewiring''), and \(\beta_{wn}\) (the overall dissimilarity including
non-shared species). We can express the components \(a\), \(b\), and
\(c\) of Koleff \emph{et al.} (2003) as the cardinality of the following
sets:

\begin{longtable}[]{@{}llll@{}}
\toprule
Component & \(a\) & \(b\) & \(c\)\tabularnewline
\midrule
\endhead
\(\beta_{os}\) & \(E_c\) & \(E_{sn}\) & \(E_{sm}\)\tabularnewline
\(\beta_{wn}\) & \(E_c\) & \(E_{sn} \cup E_{un}\) &
\(E_{sm} \cup E_{um}\)\tabularnewline
\bottomrule
\end{longtable}

These decompositions are used to perform the calculations of
\(\beta\)-diversity in the \texttt{EcologicalNetworks.jl} package
(Banville \emph{et al.} 2021) for Julia, which I use for the following
numerical experiments.

\hypertarget{quantifying-the-importance-of-species-turnover}{%
\subsection{Quantifying the importance of species
turnover}\label{quantifying-the-importance-of-species-turnover}}

The difference between \(\beta_{os}\) and \(\beta_{wn}\) stems from the
species dissimilarity between \(\mathcal{M}\) and \(\mathcal{N}\), and
it is easier to understand the effect of turnover by picking a
dissimilarity measure to work as an exemplar. At this point, Fründ
(2021) introduce a confusin terminology in their work, stating that
S\o rensen's and Whittaker's measures of dissimilarity are the same in
the Koleff \emph{et al.} (2003) framework (they are not; in practice,
\(\beta_{Sor}=1-\beta_{w}\)), and (ii) noting Whittaker's measure as
\((b+c)/(2a+b+c)\), which in the Koleff \emph{et al.} (2003) framework
is, in fact, \(\beta_t\) (Wilson \& Shmida 1984). This does not change
the overall conclusions as these measures can be re-expressed to
converge to the same value. For the sake of consistency, I will use
\(\beta_t\) moving forward; it returns values in \([0,1]\), with \(0\)
meaning complete similarity, and \(1\) meaning complete dissimilarity.

\hypertarget{establishing-that-beta_wn-ge-beta_os}{%
\subsubsection{\texorpdfstring{Establishing that
\(\beta_{wn} \ge \beta_{os}\)}{Establishing that \textbackslash beta\_\{wn\} \textbackslash ge \textbackslash beta\_\{os\}}}\label{establishing-that-beta_wn-ge-beta_os}}

Based on a partition between three sets of cardinality \(a\), \(b\), and
\(c\),

\[\beta_t = \frac{b+c}{2a+b+c}\,.\]

So as to simplify the notation of the following section, I will
introduce a series of new variables. Let \(A = |E_c|\) be the number of
links that are identical between networks; \(S = |E_{sn} \cup E_{sm}|\)
be the number of links that are not shared, but only involve shared
species (\emph{i.e.} links from \(\mathcal{M}\cup\mathcal{N}\)
established between species from \(\mathcal{M}\cap\mathcal{N}\)); and
\(U = |E_{un} \cup E_{um}|\) the number of links that are not shared,
and involve at least one unique species. Adopting the perspective
developed in the previous section, wherein networks are sets and the
measures of \(\beta\)-diversity operates on these sets, highlights the
conceptual issue in the Fründ (2021) alternative normalization: they are
using components of the networks that are \emph{not} part of the
networks being compared.

There are two important points to note here. First, the number or
proportion of species that are shared is not involved in the
calculation. Second, the connectance of either network is not involved
in the calculation. That all links counted in \emph{e.g.} \(U\) come
from \(\mathcal{M}\), or that they are evenly distributed between
\(\mathcal{M}\) and \(\mathcal{N}\), has no impact on the result. This
is a desirable property of the approach: whatever quantitative value of
the components of dissimilarity can be interpreted in the light of the
connectance and species turnover \emph{without} any risk of circularity.
Therefore the argument of Fründ (2021), whereby the \(\beta_{os}\)
component should decrease with turnover, and be invariant to
connectance, does not hold: the very point of the approach is to provide
measures that can be interpreted in the light of connectance and species
turnover.

The final component of network dissimilarity in Poisot \emph{et al.}
(2012) is \(\beta_{st}\), \emph{i.e.} the part of \(\beta_{wn}\) that is
not explained by changes in interactions between shared species
(\(\beta_{os}\)), and therefore stems from species turnover. This
fraction is defined as \(\beta_{st} = \beta_{wn}-\beta_{os}\).

The expression of \(\beta_{st}\) does not involve a partition into sets
that can be plugged into the framework of Koleff \emph{et al.} (2003),
because the part of \(\mathcal{M}\) and \(\mathcal{N}\) that are
composed of their unique species cannot, by definition, share
interactions. One could, theoretically, express these as
\(\mathcal{M} \setminus \mathcal{N} = (V_m, E_{um})\) and
\(\mathcal{N} \setminus \mathcal{M} = (V_v, E_{un})\) (note the
non-commutativity here), but the dissimilarity between these networks is
trivially maximal for the measures considered.

Using the \(\beta_t\) measure of dissimilarity, we can re-write (using
the notation with \(A\), \(S\), and \(U\))

\[\beta_{os} = \frac{S}{2A+S}\,,\]

and

\[\beta_{wn} = \frac{S+U}{2A+S+U}\,.\]

Note that \(\beta_{os}\) has the form \(x/y\) with \(x = S\) and
\(y = 2A+S\), and \(\beta_{wn}\) has the form \((x+k)/(y+k)\), with
\(k = U\). As long as \(k \ge 0\), it is guaranteed that
\(\beta_{wn} \ge \beta_{os}\), and therefore that
\(0 \ge \beta_{st} \ge 1\); as \(A\), \(S\), and \(U\) are cardinalities
of sets, they are necessarily satisfying this condition.

We can get an expression for \(\beta_{st}\), by bringing \(\beta_{os}\)
and \(\beta_{wn}\) to a common denominator and simplifying the
numerator:

\[\beta_{st} = \frac{2AU}{(2A+S)(2A+S+U)}\,.\]

Note that this value varies in a non-monotonic way with regards to the
number of interactions that are part of the common set of species --
this is obvious when developing the denominator into

\[4A^2 + S^2 + 4AS + 2AU + SU\,,\]

As such, we expect that the value of \(\beta_{st}\) will vary in a
hump-shaped way with the proportion of shared interactions. For this
reason, Poisot \emph{et al.} (2012) suggest that
\(\beta_{st}/\beta{wn}\) (alt. \(1-\beta_{os}/\beta_{wn}\)) is a better
indicator of the \emph{relative} importance of turnover processes on
network dissimilarity. This can be calculated as

\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(2A+S+U)}\times\frac{S+U}{2A+S+U}\,,\]

which reduces to

\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(S+U)}\,.\]

The roots of this expression are \(A=0\) (the turnover of species has no
contribution to the difference between \(\beta_{wn}\) and \(\beta_{os}\)
if there are no shared species, and therefore no rewiring), and for
\(U = 0\) (the turnover of species has no contribution if all species
are shared).

\hypertarget{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}{%
\subsubsection{Numerical experiment: response of the components to
different sources of network
variation}\label{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}}

To illustrate the behavior of \(\beta_{st}\), I conducted a simple
numerical experiment in which two networks have the same number of
interactions \(L\) (recall from the previous section that we do not need
to set a number of species yet), and these interactions are partitionned
according to proportions \(p_s\) and \(p_r\) into shared (\(A\)),
rewired (\(S\)), and unique (\(U\)) links, with \(A = p_s\times L\),
\(S = (1-p_s)\times p_r\times L\), and
\(U = (1-p_s)\times (1-p_r)\times L\). The results are represented in
fig.~\ref{fig:numexp1}.

\begin{figure}
\hypertarget{fig:numexp1}{%
\centering
\includegraphics{figures/numexp1.png}
\caption{Values of \(\beta_{os}\), \(\beta_{wn}\), \(\beta_{st}\), and
\(\beta_{st}/\beta_{wn}\) as a function of the proportion of rewired
links and the proportion of shared links.}\label{fig:numexp1}
}
\end{figure}

The rewiring component \(\beta_{os}\) varies as a function of the
proportion of shared links that are rewired; by contrast, \(\beta_{wn}\)
varies \emph{only} as a function of the proportion of links that are
shared: that the unshared links are established between common or unique
species has no effect on overall network dissimilarity. The quadratic
nature of the denominator for \(\beta_{st}\) is clear here, with a
maximum reach when there is no re-wiring, and a small number of shared
links (\emph{i.e.} the networks are almost entirely dissimilar except
for the links between shared species). Althought the \emph{raw} values
of \(\beta_{st}\) may seem low, the normalization using
\(\beta_{st}/\beta_{wn}\) magnifies this effect: its values are indeed
maximized when the rewiring is lower, \emph{i.e.} all of the network
variation stems from turnover processes.

\hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
\subsection{Is this decomposition over-estimating the effect of
``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}

One of the arguments put forth by Fründ (2021) is that the decomposition
outlined above will overestimate the effect of rewiring; I argue that
this is based on a misunderstanding of what \(\beta_{st}\) achieves. It
is paramount to clarify that \(\beta_{st}\) is not a direct measure of
the importance of turnover: it is a quantification of the relative
impact of rewiring to overall dissimilarity, which, all non-turnover
mechanisms being accounted for in the decomposition, can be explained by
turnover mechanisms. In this section, I present two numerical
experiments showing (i) that the \(\beta_{os}\) component is in fact an
accurate measure of rewiring, and (ii) that \(\beta_{st}\) captures the
consequences of species turnover, and of the interactions brought by
unique species.

\hypertarget{illustrations-on-arbitrarily-small-networks-are-biased}{%
\subsubsection{Illustrations on arbitrarily small networks are
biased}\label{illustrations-on-arbitrarily-small-networks-are-biased}}

We can re-calculate the illustration of Fründ (2021), wherein a pair of
networks with two shared interactions (\(A = 2\)) receive either an
interaction in \(S\), in \(U\), or in both:

\begin{longtable}[]{@{}lllllll@{}}
\toprule
\(A\) & \(S\) & \(U\) & \(\beta_{os}\) & \(\beta_{wn}\) & \(\beta_{st}\)
& \(\beta_{st}/\beta_{wn}\)\tabularnewline
\midrule
\endhead
2 & 0 & 0 & 0 & 0 & 0 &\tabularnewline
2 & 1 & 0 & \(1/5\) & \(1/5\) & 0 & 0\tabularnewline
2 & 0 & 1 & 0 & \(1/5\) & \(1/5\) & 0\tabularnewline
2 & 1 & 1 & \(1/5\) & \(1/3\) & \(2/15\) & \(2/5\)\tabularnewline
\bottomrule
\end{longtable}

The over-estimation argument hinges on the fact that
\(\beta_{st} < \beta_{os}\) in the last situation (one interaction as
rewiring, one as turnover). Reaching the conclusion of an overestimation
from this is based on a mis-interpretation of what \(\beta_{st}\) means.
The correct interpretation is that, out of the entire network
dissimilarity, only three-fifths are explained by re-wiring. The fact
that this fraction is not exactly one-half comes from the fact that the
Wilson \& Shmida (1984) measure counts shared interactions \emph{twice}
(\emph{i.e.} it has a \(2A\) term), which over-amplifies the effect of
shared interactions as the network is really small. Running the same
calculations with \(A = 10\) gives a relative importance of the turnover
processes of 47\%, and \(\beta_{st}\) goes to \(1/2\) as \(A/(S+U)\)
increases. As an additional caveat, the value of \(\beta_{st}\) will
depend on the measure of beta-diversity used. Measures that do not count
the shared interaction twice are not going to amplify the effect of
rewiring.

\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}{%
\subsubsection{Numerical experiment: the decomposition captures the
roles of rewiring and turnover
accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}}

Consider two bipartite networks, each with \(R\) species on either side,
and each with the same connectance \(\rho\). We will assume that these
networks \emph{share} a proportion \(p\) of their species from one side
(and share all species from the other), and that the interactions
between these species are undergo rewiring with at a rate \(q\). This is
sufficient information to calculate the values of \(A\), \(S\), and
\(U\) required to get the values of \(\beta_{os}\) and \(\beta_{wn}\).
Note that the simplification of assuming that only species from one side
can vary is merely for the sake of simplicity, but does not decrease the
generality of the argument.

Each network will have \(\rho(1-p)R^2\) interactions that are unique due
to species turnover, and so

\[U = 2\rho(1-p)R^2\,.\]

The part of both networks composed of overlapping species has
\(\rho pR^2\) interactions, of which \(\rho (1-q) p R^2\) are shared,
and \(\rho qp R^2\) underwent rewiring. This leads to

\[A = \rho (1-q) p R^2\,,\]

and

\[S = \rho pq R^2\,.\]

Note that we can drop the multiplicative constant \(R^2\), making the
result independent of the size of the network. Based on these
components, we can get the values of \(\beta_{os}\) and \(\beta_{wn}\),
as presented in fig.~\ref{fig:numexp2}.

\begin{figure}
\hypertarget{fig:numexp2}{%
\centering
\includegraphics{figures/numexp2.png}
\caption{Response of \(\beta_{os}\) and \(\beta_{wn}\), and the
consequences on \(\beta_{st}\), to changes in rewiring probability
(\(q\)) and probability of species sharing (\(p\)). As expected,
\(\beta_{os}\) is not affected by species turnover, but increases with
the rewiring probability. By contrast, \(\beta_{wn}\) increases when the
rewiring probability is higher \emph{and} when fewer species are shared.
This has important consequences for \(\beta_{st}\): its value is
maximized for low species sharing, and decreases for high rewiring
probability.}\label{fig:numexp2}
}
\end{figure}

The value of \(\beta_{os}\) is entirely unchanged by variations in \(p\)
(species sharing), and responds \emph{only} to changes in \(q\) (the
probability of rewiring), whereas as expected, \(\beta_{wn}\) responded
to changes in both of these parameters: the most dissimilar networks
have low species sharing (interactions are dissimilar because brought by
unique species), and high rewiring (shared species do not share
interactions). The relative changes in \(\beta_{os}\) and \(\beta_{wn}\)
lead to predictable changes in \(\beta_{st}\): its value is maximized
when both rewiring \emph{and} species sharing are low. Increasing
rewiring decreases the impact of species turnover (because, for an equal
number of interactions, the dissimilarity of interactins in shared
species contributes more to \(\beta_{wn}\)); increasing the chance of
sharing species also does decrease \(\beta_{st}\), trivially because
there is no species turnover anymore. Note that when using the
correction of \(\beta_{st}/\beta_{wn}\), the effect of species turnover
is magnified for low probabilities of re-wiring.

In conclusion, this numerical experiment shows that the decomposition as
initially presented by Poisot \emph{et al.} (2012), \emph{i.e.} using
denominators that make sense from a network composition point of view,
succeeds at capturing the relative effect of turnover and rewiring.
\#\#\# Numerical experiment: the decomposition captures the roles of
species turnover and connectance accurately

Consider now two bipartite networks, which still have \(R\) species on
either side, but differ in their connectance (\(\rho_1\) and \(\rho_2\))
-- by maintaining the assumption that species on one side are shared
with probability \(p\), and that interactions between shared species are
rewired at probability \(q\), we can examine the effect of varying both
connectance and turnover on the value of the \(\beta\)-diversity
components. Note that, although not presented, we will drop the
multiplicative constant \(R^2\) from all calculations, as it is a common
factor for all values; again, this implies that the results presented
here are independant of network richness.

The number of unique links due to species turnover is

\[U = (1-p)(\rho_1 + \rho_2)\,,\]

which decreases with the proportion of shared species, but increases
with connectance. The number of links between shared species takes a
little more steps to calculate. First, amongst the \(pR^2\) species in
both sub-graphs, network 1 will have \(\rho_1 pR^2\), and network 2 will
have \(\rho_2 pR^2\). Because \(\rho_1 \neq \rho_2\), there are only
\(\text{min}(\rho_1, \rho_2)pR^2\) links that can be shared, a
proportion \(q\) of which will undergo re-wiring, and a proportion
\((1-q)\) of which will be shared. This leads to the expression (after
dropping \(R^2\)) for the number of shared links:

\[A = p (1-q) \text{min}(\rho_1, \rho_2)\,.\]

The number of unique links due to shared species is the sum of all links
in network 1 (\(\rho_1 R^2\)), minus the sum of the shared links
(\(AR^2\)) and the unique links due to species turnover
(\((1-p)\rho_1R^2\)); this same quantity is calculated in the same way
for the second networks, leading to (after dropping the multiplicative
constant \(R^2\) and some simplifications)

\[S = p (\rho_1 + \rho_2) - 2A\,.\]

Note that as expected, this last quantity scales with the proportion of
shared species (\(p\)) \emph{and} with connectance (as shared species
bring more of their interactions), but decreases with the size of the
shared links components. The consequences of varying \(\rho_2\) and
\(p\) are presented in fig.~\ref{fig:numexp3}.

\begin{figure}
\hypertarget{fig:numexp3}{%
\centering
\includegraphics{figures/numexp3.png}
\caption{Effects of varying the connectance of the second network
(\(\rho_2\)) and the proportion of shared species (\(p\)) on the values
of the \(\beta\)-diversity components. As expected, \(\beta_{os}\) is
still independent of species turnover, and \(\beta_{wn}\) increases when
species turnover increases, or when the connectances become more
dissimilar. These figures have been generated with \(\rho_1 = 0.25\) and
\(q = 0.15\), and the results are qualitatively robust to changes in
these parameters.}\label{fig:numexp3}
}
\end{figure}

Although \(\beta_{os}\) is only responding to changes in connectance (as
is expected, seeing that the relative connectances of both networks
appear in the expression for \(S\) and \(A\)), \(\beta_{wn}\) changes in
response to both parameters. Specifically, increasing the difference in
connectance between the two networks, especially when also increasing
the species dissimilarity, results in more dissimilar networks -- this
is because unique species from both networks bring their own
interactions (at rate \(\rho_1\) and \(\rho_2\)), and therefore
contribute to dissimilarity. It is particularly noteworthy that
\(\beta_{st}\), regardless of the differences in connectance, increases
with the proportion of unique species. At an equal proportion of shared
species, \(\beta_{st}\) \emph{decreases} with differences in
connectance: this is an equally expected result, which indicates that
the difference between \(\beta_{os}\) and \(\beta_{wn}\) is in part
explained by non-turnover mechanisms (here, changes in connectance).
Relying on the \(\beta_{st}/\beta_{wn}\) correction again magnifies this
effect, without changing their interpretation.

\hypertarget{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}{%
\subsection{Does the partition of network dissimilarity needs a new
normalization?}\label{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}}

Based on the arguments presented above, I do not think the suggestion of
Fründ (2021) to change the denominator of \(\beta_{os}\) makes sense as
a default; the strength of the original approach by Poisot \emph{et al.}
(2012) is indeed that the effect of turnover is based on a rigorous
definition of networks as graphs (as opposed to networks as matrices),
in which the induction of vertices from the edgelist being compared
gives rise to biologically meaningful denominators. The advantage of
this approach is that at no time does the turnover of species itself (or
indeed, as shown in many places in this manuscript, the network
richness), or the connectance of the network, enter into the
calculation. As such, it is possible to use \(\beta_{os}\) and
\(\beta_{wn}\) in relationship to these terms, calculated externally (as
was recently done by \emph{e.g.} Higino \& Poisot 2021), without
creating circularities.

The choice of changing the denominator hinges on what one admits as a
definition for \(\beta_{st}\). If the point of \(\beta_{st}\) is to be a
component of overall \(\beta\)-diversity as advocated by Fründ (2021)
and Novotny (2009), a change of numerator \emph{might} be acceptable.
Nevertheless, this change of numerator contributes to blurring the
frontier between a measure of interaction dissimilarity and a measure of
community dissimilarity which starts to add the effect of relative
richness; this later case warrants a thorough methodological assessment.
Conversely, if as we argue in Poisot \emph{et al.} (2012),
\(\beta_{st}\) is to be meant as a \emph{guide} to the interpretation of
\(\beta_{wn}\) and \(\beta_{os}\), and related to actual measures of
species turnover and network connectance, one must not change the
denominator.

It is essential to recognize that there are multiple reasons to
calculate network dissimilarity, and it is our opinion that the
arguments levied by Fründ (2021) against the original partition stem
from a misunderstanding of what it intends to do (and does, indeed, do
well), not from intrinsic methodological issues in the partition itself.
Based on the results presented in this contribution, I argue that the
original partition of network \(\beta\)-diversity from Poisot \emph{et
al.} (2012) should remain the default.

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\singlespacing

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
EcologicalNetworks.jl: Two complementary packages for analyzing
ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
2721.

\leavevmode\hypertarget{ref-Baronio2021NatFir}{}%
Baronio, G.J., Souza, C.S., Maruyama, P.K., Raizer, J., Sigrist, M.R. \&
Aoki, C. (2021). Natural fire does not affect the structure and beta
diversity of plant-pollinator networks, but diminishes floral-visitor
specialization in Cerrado. \emph{Flora}, 281, 151869.

\leavevmode\hypertarget{ref-Campos-Moreno2021ImpInt}{}%
Campos-Moreno, D.F., Dyer, L.A., Salcido, D., Massad, T.J.,
Pérez-Lachaud, G., Tepe, E.J., \emph{et al.} (2021). Importance of
interaction rewiring in determining spatial and temporal turnover of
tritrophic (Piper-caterpillar-parasitoid) metanetworks in the Yucatán
Península, México. \emph{Biotropica}, 53, 1071--1081.

\leavevmode\hypertarget{ref-Canard2014EmpEva}{}%
Canard, E.F., Mouquet, N., Mouillot, D., Stanko, M., Miklisova, D. \&
Gravel, D. (2014). Empirical evaluation of neutral interactions in
host-parasite networks. \emph{The American Naturalist}, 183, 468--479.

\leavevmode\hypertarget{ref-Dunne2006NetStr}{}%
Dunne, J.A. (2006). The Network Structure of Food Webs. In:
\emph{Ecological networks: Linking structure and dynamics} (eds. Dunne,
J.A. \& Pascual, M.). Oxford University Press, pp. 27--86.

\leavevmode\hypertarget{ref-Frund2021DisSpe}{}%
Fründ, J. (2021). Dissimilarity of species interaction networks: How to
partition rewiring and species turnover components. \emph{Ecosphere},
12, e03653.

\leavevmode\hypertarget{ref-Higino2021BetPhy}{}%
Higino, G.T. \& Poisot, T. (2021). Beta and phylogenetic diversities
tell complementary stories about ecological networks biogeography.
\emph{Parasitology}, 1--23.

\leavevmode\hypertarget{ref-Koleff2003MeaBet}{}%
Koleff, P., Gaston, K.J. \& Lennon, J.J. (2003). Measuring beta
diversity for presence--absence data. \emph{Journal of Animal Ecology},
72, 367--382.

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
Legendre, P. \& De Cáceres, M. (2013). Beta diversity as the variance of
community data: Dissimilarity coefficients and partitioning.
\emph{Ecology Letters}, 16, 951--963.

\leavevmode\hypertarget{ref-Magrach2017PlaNet}{}%
Magrach, A., Holzschuh, A., Bartomeus, I., Riedinger, V., Roberts,
S.P.M., Rundlöf, M., \emph{et al.} (2017). Plant-pollinator networks in
semi-natural grasslands are resistant to the loss of pollinators during
blooming of mass-flowering crops. \emph{Ecography}, n/a--n/a.

\leavevmode\hypertarget{ref-Novotny2009BetDiv}{}%
Novotny, V. (2009). Beta diversity of plant--insect food webs in
tropical forests: A conceptual framework. \emph{Insect Conservation and
Diversity}, 2, 5--9.

\leavevmode\hypertarget{ref-Olsson2021IntPla}{}%
Olsson, R.L., Brousil, M.R., Clark, R.E., Baine, Q. \& Crowder, D.W.
(2021). Interactions between plants and pollinators across urban and
rural farming landscapes. \emph{Food Webs}, 27, e00194.

\leavevmode\hypertarget{ref-Poisot2012DisSpe}{}%
Poisot, T., Canard, E., Mouillot, D., Mouquet, N. \& Gravel, D. (2012).
The dissimilarity of species interaction networks. \emph{Ecology
Letters}, 15, 1353--1361.

\leavevmode\hypertarget{ref-Poisot2016StrPro}{}%
Poisot, T., Cirtwill, A.R., Cazelles, K., Gravel, D., Fortin, M.-J. \&
Stouffer, D.B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, 7, 303--312.

\leavevmode\hypertarget{ref-Poisot2017HosPar}{}%
Poisot, T., Gueveneux-Julien, C., Fortin, M.-J., Gravel, D. \& Legendre,
P. (2017). Hosts, parasites and their interactions respond to different
climatic variables. \emph{Global Ecology and Biogeography}, n/a--n/a.

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D.B. \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, 124, 243--251.

\leavevmode\hypertarget{ref-Souza2021PlaSam}{}%
Souza, C.S., Maruyama, P.K., Santos, K.C.B.S., Varassin, I.G., Gross,
C.L. \& Araujo, A.C. (2021). Plant-centred sampling estimates higher
beta diversity of interactions than pollinator-based sampling across
habitats. \emph{New Phytologist}, 230, 2501--2512.

\leavevmode\hypertarget{ref-Trojelsgaard2016EcoNet}{}%
Trøjelsgaard, K. \& Olesen, J.M. (2016). Ecological networks in motion:
Micro- and macroscopic variability across scales. \emph{Functional
Ecology}, 30, 1926--1935.

\leavevmode\hypertarget{ref-Tuomisto2010DivBet}{}%
Tuomisto, H. (2010). A diversity of beta diversities: Straightening up a
concept gone awry. Part 1. Defining beta diversity as a function of
alpha and gamma diversity. \emph{Ecography}, 33, 2--22.

\leavevmode\hypertarget{ref-Wilson1984MeaBet}{}%
Wilson, M.V. \& Shmida, A. (1984). Measuring Beta Diversity with
Presence-Absence Data. \emph{Journal of Ecology}, 72, 1055--1064.

\end{CSLReferences}

\end{document}

### 2021-09-24 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@0bc3b6c27c90321866495cce141c996ca59614b5 🚀

--- 2021-09-24 draft.tex
@@
 %!TEX TS-program = xelatex
-\documentclass[11pt]{article}
+\documentclass[10pt,oneside]{article}
+\usepackage[fontsize=9pt]{scrextend}
 
@@
 \newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\usepackage[table,dvipsnames]{xcolor}
 
-\usepackage{lineno}
-\usepackage[nolists,noheads]{endfloat}
+\geometry{includemp,
+            letterpaper,
+            top=2.4cm,
+            bottom=2.4cm,
+            left=1.0cm,
+            right=1.0cm,
+            marginparwidth=5cm,
+            marginparsep=1.0cm}
+
+\usepackage[singlelinecheck=off]{caption}
+
+\captionsetup{
+  font={small},
+  labelfont={bf},
+  format=plain,
+  labelsep=quad
+}
+
+\usepackage{floatrow}
+
+\floatsetup[figure]{margins=hangright,
+              facing=no,
+              capposition=beside,
+              capbesideposition={center,outside},
+              floatwidth=\textwidth}
+
+% \floatsetup[table]{margins=hangright,
+%              facing=no,
+%              capposition=beside,
+%              capbesideposition={center,outside},
+%              floatwidth=\textwidth}
 
@@
 
-\tolerance=1
-\emergencystretch=\maxdimen
-\hyphenpenalty=10000
-\hbadness=10000
+\setcounter{secnumdepth}{5}
 
-\doublespacing
+\usepackage{titlesec}
+
+\titleformat{\section}[block]
+{\normalfont\large\sffamily}
+{\thesection}{.5em}{\titlerule\\[.8ex]\bfseries}
+
+\titleformat{\subsection}[runin]
+{\normalfont\fontseries{b}\selectfont\filright\sffamily}
+{\thesubsection.}{.5em}{}
+
+\titleformat{\subsubsection}[runin]
+{\normalfont\itshape\rmfamily\bfseries}{\thesubsubsection}{1em}{}
+
+\fancypagestyle{firstpage}
+{
+   \fancyhf{}
+   \renewcommand{\headrulewidth}{0pt}
+   \fancyfoot[R]{\footnotesize\ccby}
+   \fancyfoot[L]{\footnotesize\sffamily\today}
+}
 
@@
 }
+
+\usepackage{tikz}
 \begin{document}
-\raggedright
-\thispagestyle{empty}
-{\Large\bfseries\sffamily Dissimilarity of species interaction networks:
-quantifying the effect of turnover and rewiring}
-\vskip 5em
+\tikz [remember picture, overlay] %
+\node [shift={(-0.6in,1.1cm)},scale=0.2,opacity=0.4] at (current page.south east)[anchor=south east]{\includegraphics{logo}};%
+\pagestyle{normal}
+\thispagestyle{firstpage}
 
+\newcommand{\colorRule}[3][black]{\textcolor[HTML]{#1}{\rule{#2}{#3}}}
+
+\noindent {\LARGE \textbf{\textsf{Dissimilarity of species interaction
+networks: quantifying the effect of turnover and rewiring}}}
+
+\medskip
+\begin{flushleft}
+{\small
 %
@@
 \,\textsuperscript{1,2,‡}
-
-\textsuperscript{1}\,Université de
-Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
-Sciences
-
+\vskip 1em
+\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
+Centre for Biodiversity Sciences\\
 \textsuperscript{‡}\,These authors contributed equally to the work\\
-
+\vskip 1em
 \textbf{Correspondance to:}\\
 Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
+}
+\end{flushleft}
 
-\vfill
-This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
-Last revision: \emph{\today}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
+\noindent
 
-\clearpage
-\thispagestyle{empty}
+\marginpar{\vskip 1em\flushright
+{\small{\bfseries Keywords}:\par
+ecological network dissimilarity\\turnover partitioning\\species
+interaction networks\\}
+}
 
-\vfill
 Despite having established its usefulness in the last ten years, the
@@
 
-\vfill
-
-\clearpage
-\linenumbers
-\pagestyle{normal}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
 

### 2021-09-24 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@9eca1242e2389cdc5ed5ec693adae6b6a0d1ce29 🚀

@@
 decomposition of ecological networks in components allowing to measure
-their \(eta\)-diversity retains some methodological ambiguities.
+their \(\beta\)-diversity retains some methodological ambiguities.
 Notably, how to quantify the relative effect of mechanisms tied to
@@
 succeeds at capturing the relative effect of turnover and rewiring.
-\#\#\# Numerical experiment: the decomposition captures the roles of
-species turnover and connectance accurately
+
+\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}{%
+\subsubsection{Numerical experiment: the decomposition captures the
+roles of species turnover and connectance
+accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}}
 
@@
 
-\newpage
-
 \hypertarget{references}{%
-\subsection{References}\label{references}}
-
-\singlespacing
+\subsection*{References}\label{references}}
+\addcontentsline{toc}{subsection}{References}
 

### 2021-09-24 code/experiments/01_sources_of_variation.jl
work on numexp code

@@ -1,0 +1,68 @@
+
+# Numerical experiment 1: L1 = L2, we vary the number of shared, unique, and
+# non-unique links, and examine the consequences on the values of the OS, ST,
+# and WN components of beta diversity. The goal of this experiment is to show
+# that, without considering connectance or species turnover, we can tie the
+# values of these components to the sources of network dissimilarity.
+
+# The values of L1 and L2 are entirely arbitrary -- indeed, the manuscript shows
+# how the values are independant from these values, as we can collect the
+# non-shared links in those that are rewired, and those that are connecting
+# non-shared species. Using different values for L1 and L2 complexifies the
+# problem a little, by introducing some combinatorics questions that would just
+# stand in the way.
+L1, L2 = 200, 200
+
+# These are proportions - first, we pick a proportion of links (out of L1 or L2)
+# that are shared, and then we divide the rest of the links in those that are
+# not-shared and rewired (S in the manuscript), and those that are not-shared
+# and involve non-overlapping species (U in the paper). Note that the species
+# richness does not matter here.
+shared_links = LinRange(0.0, 1.0, L1)
+links_rewired = LinRange(0.0, 1.0, L2)
+
+# We prepare matrices to fill the calculated values
+OS = zeros(Float64, length(shared_links), length(links_rewired))
+WN = similar(OS)
+
+# Finally, we measure OS and WN - we will get the rest by substraction
+for (i, sl) in enumerate(shared_links)
+    for (j, lr) in enumerate(links_rewired)
+        common = L1 * sl
+        rewired = (L1 - common) * lr
+        turned = L1 - (common + rewired)
+        # We can assume that all links are from the same network, since the
+        # measure is symetrical
+        OS[i,j] = βₜ(common, 0.0, rewired)
+        WN[i,j] = βₜ(common, turned, rewired)
+    end
+end
+
+# We get the values of ST
+ST = WN .- OS
+
+# The following lines are plots of the results of numerical experiment 1
+
+p1 = contour(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white, levels=4)
+xaxis!(p1, "Proportion of rewired links", (0, 1))
[…]

### 2021-09-24 code/main.jl

@@ -1,5 +1,4 @@
 # Dependencies for the simulations
-using EcologicalNetworks
 using StatsPlots
 using PlotThemes
 using Random
@@ -10,77 +9,16 @@ using Statistics
 # This ensures the reproducibility of the results
 Random.seed!(90210)
 theme(:mute)
-default(frame=:box, dpi=600, size=(500, 500))
+default(; frame=:box, dpi=600, size=(500, 500), c=:lapaz)
 
-# Shortcut notation for the betadiv measure we use
-βₜ = EcologicalNetworks.KGL08
+# β-diversity function we use
+βₜ = (a,b,c) -> (b+c)/(2a+b+c)
 
-# Numerical experiment 1: L1 = L2, we vary the number of shared, unique, and
-# non-unique links, and examine the consequences on the values of the OS, ST,
-# and WN components of beta diversity. The goal of this experiment is to show
-# that, without considering connectance or species turnover, we can tie the
-# values of these components to the sources of network dissimilarity.
+_FIGPATH = joinpath("figures")
+ispath(_FIGPATH) || mkpath(_FIGPATH)
 
-# The values of L1 and L2 are entirely arbitrary -- indeed, the manuscript shows
-# how the values are independant from these values, as we can collect the
-# non-shared links in those that are rewired, and those that are connecting
-# non-shared species. Using different values for L1 and L2 complexifies the
-# problem a little, by introducing some combinatorics questions that would just
-# stand in the way.
-L1, L2 = 200, 200
-
-# These are proportions - first, we pick a proportion of links (out of L1 or L2)
-# that are shared, and then we divide the rest of the links in those that are
-# not-shared and rewired (S in the manuscript), and those that are not-shared
-# and involve non-overlapping species (U in the paper). Note that the species
-# richness does not matter here.
-shared_links = LinRange(0.0, 1.0, 200)
-links_rewired = LinRange(0.0, 1.0, 202)
-
-# We prepare matrices to fill the calculated values
-OS = zeros(Float64, length(shared_links), length(links_rewired))
-WN = similar(OS)
-
-# Finally, we measure OS and WN - we will get the rest by substraction
-for (i, sl) in enumerate(shared_links)
-    for (j, lr) in enumerate(links_rewired)
-        common = L1 * sl
[…]

### 2021-09-24 code/experiments/01_sources_of_variation.jl
up figures

@@ -43,22 +43,22 @@ ST = WN .- OS
 
 # The following lines are plots of the results of numerical experiment 1
 
-p1 = contour(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white, levels=4)
+p1 = contour(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p1, "Proportion of rewired links", (0, 1))
 yaxis!(p1, "Proportion of shared links", (0, 1))
 title!(p1, "βos")
 
-p2 = contour(links_rewired, shared_links, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white, levels=4)
+p2 = contour(links_rewired, shared_links, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p2, "Proportion of rewired links", (0, 1))
 yaxis!(p2, "Proportion of shared links", (0, 1))
 title!(p2, "βst")
 
-p3 = contour(links_rewired, shared_links, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white, levels=4)
+p3 = contour(links_rewired, shared_links, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p3, "Proportion of rewired links", (0, 1))
 yaxis!(p3, "Proportion of shared links", (0, 1))
 title!(p3, "βwn")
 
-p4 = contour(links_rewired, shared_links, ST ./ WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white, levels=4)
+p4 = contour(links_rewired, shared_links, ST ./ WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p4, "Proportion of rewired links", (0, 1))
 yaxis!(p4, "Proportion of shared links", (0, 1))
 title!(p4, "βst / βwn")

### 2021-09-24 code/experiments/02_constant_connectance.jl

@@ -1,0 +1,51 @@
+# Numerical experiment 2 - at some constant connectance, we examine the effect
+# of changing the species overlap between two networks, as well as the
+# proportion that overlapping species keep their interactions. This will provide
+# evidence that the components of the beta diversity partition are responding to
+# the correct ecological processes. Note that, as explained in the manuscript,
+# this calculation is independent of network richness, and does not require any
+# simulations.
+
+Co = 0.4
+p_values = LinRange(0.0, 1.0, 200) # Fraction of shared species
+q_values = LinRange(0.0, 1.0, 202) # Fraction of rewired links
+
+# Stores
+OS = zeros(Float64, length(p_values), length(q_values))
+WN = similar(OS)
+
+for (i, p) in enumerate(p_values)
+    for (j, q) in enumerate(q_values)
+        A = p * (1 - q) * Co
+        S = p * q * Co
+        U = Co * (1 - p) * 2
+        OS[i,j] = S / (2A + S)
+        WN[i,j] = (S + U) / (2A + S + U)
+    end
+end
+
+ST = WN .- OS
+
+p1 = contour(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+xaxis!(p1, "Rewiring probability", (0, 1))
+yaxis!(p1, "Species sharing probability", (0, 1))
+title!(p1, "βos")
+
+p2 = contour(q_values, p_values, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+xaxis!(p2, "Rewiring probability", (0, 1))
+yaxis!(p2, "Species sharing probability", (0, 1))
+title!(p2, "βst")
+
+p3 = contour(q_values, p_values, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+xaxis!(p3, "Rewiring probability", (0, 1))
+yaxis!(p3, "Species sharing probability", (0, 1))
+title!(p3, "βwn")
+
+p4 = contour(q_values, p_values, ST./WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+xaxis!(p4, "Rewiring probability", (0, 1))
+yaxis!(p4, "Species sharing probability", (0, 1))
+title!(p4, "βst / βwn")
+
+plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
+
+savefig(joinpath(_FIGPATH, "numexp2.png"))

### 2021-09-24 code/experiments/03_variable_connectance.jl

@@ -1,0 +1,45 @@
+# Numerical experiment 3 - we vary the connectance of one of the networks
+# relative to that of the other
+
+Co1 = 0.25
+Co2 = LinRange(0.0, 1.0, 200)
+
+p = LinRange(0.0, 1.0, 204)'
+q = 0.15
+
+# The expressions are all calculated in main text
+A = p.*(1.0.-q).*min.(Co1,Co2)
+U = (1.0.-p).*Co1 .+ (1.0.-p).*Co2
+S = Co1 .- (A .+ (1.0.-p).*Co1) .+ Co2 .- (A .+ (1.0.-p).*Co2)
+
+OS = S ./ (2A .+ S)
+WN = (S.+U) ./ (2A .+ S .+ U)
+ST = WN .- OS
+
+p1 = contour(Co2, p', OS', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
+xaxis!(p1, "Connectance of N₂", (0, 1))
+yaxis!(p1, "Species sharing probability", (0, 1))
+title!(p1, "βos")
+
+p2 = contour(Co2, p', ST', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
+xaxis!(p2, "Connectance of N₂", (0, 1))
+yaxis!(p2, "Species sharing probability", (0, 1))
+title!(p2, "βst")
+
+p3 = contour(Co2, p', WN', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
+xaxis!(p3, "Connectance of N₂", (0, 1))
+yaxis!(p3, "Species sharing probability", (0, 1))
+title!(p3, "βwn")
+
+p4 = contour(Co2, p', (ST./WN)', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
+xaxis!(p4, "Connectance of N₂", (0, 1))
+yaxis!(p4, "Species sharing probability", (0, 1))
+title!(p4, "βst / βwn")
+
+plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
+
+savefig(joinpath(_FIGPATH, "numexp3.png"))

### 2021-09-24 code/main.jl

@@ -14,104 +14,16 @@ default(; frame=:box, dpi=600, size=(500, 500), c=:lapaz)
 # β-diversity function we use
 βₜ = (a,b,c) -> (b+c)/(2a+b+c)
 
+# Path to store the figures
 _FIGPATH = joinpath("figures")
 ispath(_FIGPATH) || mkpath(_FIGPATH)
 
-# Numerical experiment 1
-include(joinpath("code", "experiments", "01_sources_of_variation.jl"))
+# Path to store the experiments
+_EXPPATH = joinpath("code", "experiments")
 
-# Numerical experiment 2 - at some constant connectance, we examine the effect
-# of changing the species overlap between two networks, as well as the
-# proportion that overlapping species keep their interactions. This will provide
-# evidence that the components of the beta diversity partition are responding to
-# the correct ecological processes. Note that, as explained in the manuscript,
-# this calculation is independent of network richness, and does not require any
-# simulations.
-
-Co = 0.4
-p_values = LinRange(0.0, 1.0, 200) # Fraction of shared species
-q_values = LinRange(0.0, 1.0, 202) # Fraction of rewired links
-
-# Stores
-OS = zeros(Float64, length(p_values), length(q_values))
-WN = similar(OS)
-
-for (i, p) in enumerate(p_values)
-    for (j, q) in enumerate(q_values)
-        A = p * (1 - q) * Co
-        S = p * q * Co
-        U = Co * (1 - p) * 2
-        OS[i,j] = S / (2A + S)
-        WN[i,j] = (S + U) / (2A + S + U)
-    end
+# Run every experiment in the folder
+for f in filter(f -> endswith(f, ".jl"), readdir(_EXPPATH))
+    @info f
+    include(joinpath(_EXPPATH, f))
 end
 
-ST = WN .- OS
-
-p1 = contour(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white, levels=4)
-xaxis!(p1, "Rewiring probability", (0, 1))
-yaxis!(p1, "Species sharing probability", (0, 1))
-title!(p1, "βos")
-
[…]

### 2021-09-24 code/experiments/01_sources_of_variation.jl
replace contours by heatmaps

@@ -43,22 +43,22 @@ ST = WN .- OS
 
 # The following lines are plots of the results of numerical experiment 1
 
-p1 = contour(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p1 = heatmap(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p1, "Proportion of rewired links", (0, 1))
 yaxis!(p1, "Proportion of shared links", (0, 1))
 title!(p1, "βos")
 
-p2 = contour(links_rewired, shared_links, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p2 = heatmap(links_rewired, shared_links, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p2, "Proportion of rewired links", (0, 1))
 yaxis!(p2, "Proportion of shared links", (0, 1))
 title!(p2, "βst")
 
-p3 = contour(links_rewired, shared_links, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p3 = heatmap(links_rewired, shared_links, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p3, "Proportion of rewired links", (0, 1))
 yaxis!(p3, "Proportion of shared links", (0, 1))
 title!(p3, "βwn")
 
-p4 = contour(links_rewired, shared_links, ST ./ WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p4 = heatmap(links_rewired, shared_links, ST ./ WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p4, "Proportion of rewired links", (0, 1))
 yaxis!(p4, "Proportion of shared links", (0, 1))
 title!(p4, "βst / βwn")

### 2021-09-24 code/experiments/02_constant_connectance.jl

@@ -26,22 +26,22 @@ end
 
 ST = WN .- OS
 
-p1 = contour(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p1 = heatmap(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p1, "Rewiring probability", (0, 1))
 yaxis!(p1, "Species sharing probability", (0, 1))
 title!(p1, "βos")
 
-p2 = contour(q_values, p_values, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p2 = heatmap(q_values, p_values, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p2, "Rewiring probability", (0, 1))
 yaxis!(p2, "Species sharing probability", (0, 1))
 title!(p2, "βst")
 
-p3 = contour(q_values, p_values, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p3 = heatmap(q_values, p_values, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p3, "Rewiring probability", (0, 1))
 yaxis!(p3, "Species sharing probability", (0, 1))
 title!(p3, "βwn")
 
-p4 = contour(q_values, p_values, ST./WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p4 = heatmap(q_values, p_values, ST./WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
 xaxis!(p4, "Rewiring probability", (0, 1))
 yaxis!(p4, "Species sharing probability", (0, 1))
 title!(p4, "βst / βwn")

### 2021-09-24 code/experiments/03_variable_connectance.jl

@@ -16,25 +16,25 @@ OS = S ./ (2A .+ S)
 WN = (S.+U) ./ (2A .+ S .+ U)
 ST = WN .- OS
 
-p1 = contour(Co2, p', OS', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p1 = heatmap(Co2, p', OS', aspectratio=1, fill=true, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p1, "Connectance of N₂", (0, 1))
 yaxis!(p1, "Species sharing probability", (0, 1))
 title!(p1, "βos")
 
-p2 = contour(Co2, p', ST', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p2 = heatmap(Co2, p', ST', aspectratio=1, fill=true, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p2, "Connectance of N₂", (0, 1))
 yaxis!(p2, "Species sharing probability", (0, 1))
 title!(p2, "βst")
 
-p3 = contour(Co2, p', WN', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p3 = heatmap(Co2, p', WN', aspectratio=1, fill=true, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p3, "Connectance of N₂", (0, 1))
 yaxis!(p3, "Species sharing probability", (0, 1))
 title!(p3, "βwn")
 
-p4 = contour(Co2, p', (ST./WN)', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p4 = heatmap(Co2, p', (ST./WN)', aspectratio=1, fill=true, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p4, "Connectance of N₂", (0, 1))
 yaxis!(p4, "Species sharing probability", (0, 1))

### 2021-09-24 code/experiments/01_sources_of_variation.jl
rename beta_t

@@ -33,8 +33,8 @@ for (i, sl) in enumerate(shared_links)
         turned = L1 - (common + rewired)
         # We can assume that all links are from the same network, since the
         # measure is symetrical
-        OS[i,j] = βₜ(common, 0.0, rewired)
-        WN[i,j] = βₜ(common, turned, rewired)
+        OS[i,j] = beta_t(common, 0.0, rewired)
+        WN[i,j] = beta_t(common, turned, rewired)
     end
 end
 

### 2021-09-24 code/main.jl

@@ -12,7 +12,7 @@ theme(:mute)
 default(; frame=:box, dpi=600, size=(500, 500), c=:lapaz)
 
 # β-diversity function we use
-βₜ = (a,b,c) -> (b+c)/(2a+b+c)
+beta_t = (a,b,c) -> (b+c)/(2a+b+c)
 
 # Path to store the figures
 _FIGPATH = joinpath("figures")

### 2021-09-24 code/experiments/01_sources_of_variation.jl
change resolution of numexp 1

@@ -5,32 +5,27 @@
 # that, without considering connectance or species turnover, we can tie the
 # values of these components to the sources of network dissimilarity.
 
-# The values of L1 and L2 are entirely arbitrary -- indeed, the manuscript shows
-# how the values are independant from these values, as we can collect the
-# non-shared links in those that are rewired, and those that are connecting
-# non-shared species. Using different values for L1 and L2 complexifies the
-# problem a little, by introducing some combinatorics questions that would just
-# stand in the way.
-L1, L2 = 200, 200
+steps = 100
 
 # These are proportions - first, we pick a proportion of links (out of L1 or L2)
 # that are shared, and then we divide the rest of the links in those that are
 # not-shared and rewired (S in the manuscript), and those that are not-shared
 # and involve non-overlapping species (U in the paper). Note that the species
 # richness does not matter here.
-shared_links = LinRange(0.0, 1.0, L1)
-links_rewired = LinRange(0.0, 1.0, L2)
+
+shared_links = LinRange(0.0, 1.0, steps)
+links_rewired = LinRange(0.0, 1.0, steps)
 
 # We prepare matrices to fill the calculated values
-OS = zeros(Float64, length(shared_links), length(links_rewired))
+OS = zeros(Float64, steps, steps)
 WN = similar(OS)
 
 # Finally, we measure OS and WN - we will get the rest by substraction
 for (i, sl) in enumerate(shared_links)
     for (j, lr) in enumerate(links_rewired)
-        common = L1 * sl
-        rewired = (L1 - common) * lr
-        turned = L1 - (common + rewired)
+        common = sl
+        rewired = (1.0 - common) * lr
+        turned = 1.0 - (common + rewired)
         # We can assume that all links are from the same network, since the
         # measure is symetrical
         OS[i,j] = beta_t(common, 0.0, rewired)

### 2021-09-24 code/main.jl
use batlow for measures

@@ -9,7 +9,7 @@ using Statistics
 # This ensures the reproducibility of the results
 Random.seed!(90210)
 theme(:mute)
-default(; frame=:box, dpi=600, size=(500, 500), c=:lapaz)
+default(; frame=:box, dpi=600, size=(500, 500), c=:batlow)
 
 # β-diversity function we use
 beta_t = (a,b,c) -> (b+c)/(2a+b+c)

### 2021-09-24 code/experiments/04_difference_denominator.jl
add a numerical experiment to show the difference between methods

@@ -1,0 +1,52 @@
+# Numerical experiment 4 - difference in rewiring value
+
+steps = 150
+
+Co = 0.99
+p_values = LinRange(0.0, 1.0, steps) # Fraction of shared species
+q_values = LinRange(0.0, 1.0, steps) # Fraction of rewired links
+
+# Stores
+OS = zeros(Float64, length(p_values), length(q_values))
+FR = similar(OS)
+WN = similar(OS)
+
+for (i, p) in enumerate(p_values)
+    for (j, q) in enumerate(q_values)
+        A = p * (1 - q) * Co
+        S = p * q * Co
+        U = Co * (1 - p) * 2
+        OS[i,j] = S / (2A + S)
+        FR[i,j] = S / (2A + S + U)
+        WN[i,j] = (S+U) / (2A + S + U)
+    end
+end
+
+ST = WN .- OS
+FT = WN .- FR
+
+Δ = OS .- FR
+
+p1 = heatmap(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+xaxis!(p1, "Rewiring probability", (0, 1))
+yaxis!(p1, "Species sharing probability", (0, 1))
+title!(p1, "βos")
+
+p2 = heatmap(q_values, p_values, FR, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+xaxis!(p2, "Rewiring probability", (0, 1))
+yaxis!(p2, "Species sharing probability", (0, 1))
+title!(p2, "βos (common denominator)")
+
+p3 = heatmap(q_values, p_values, FT, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+xaxis!(p3, "Rewiring probability", (0, 1))
+yaxis!(p3, "Species sharing probability", (0, 1))
+title!(p3, "βst (common denominator)")
+
+p4 = heatmap(q_values, p_values, OS .- FR, clim=(0, 1), aspectratio=1, fill=true, lc=:white, c=:grayC)
+xaxis!(p4, "Rewiring probability", (0, 1))
+yaxis!(p4, "Species sharing probability", (0, 1))
+title!(p4, "Difference in rewiring estimate")
+
+plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
+
+savefig(joinpath(_FIGPATH, "numexp4.png"))

### 2021-09-24 code/experiments/04_difference_denominator.jl
update the code in numexp 4 to remove connectance

@@ -2,7 +2,6 @@
 
 steps = 150
 
-Co = 0.99
 p_values = LinRange(0.0, 1.0, steps) # Fraction of shared species
 q_values = LinRange(0.0, 1.0, steps) # Fraction of rewired links
 
@@ -13,9 +12,9 @@ WN = similar(OS)
 
 for (i, p) in enumerate(p_values)
     for (j, q) in enumerate(q_values)
-        A = p * (1 - q) * Co
-        S = p * q * Co
-        U = Co * (1 - p) * 2
+        A = p * (1 - q)
+        S = p * q
+        U = (1 - p) * 2
         OS[i,j] = S / (2A + S)
         FR[i,j] = S / (2A + S + U)
         WN[i,j] = (S+U) / (2A + S + U)

### 2021-09-25 code/experiments/01_sources_of_variation.jl
add correct solutions for next numexp

@@ -38,22 +38,22 @@ ST = WN .- OS
 
 # The following lines are plots of the results of numerical experiment 1
 
-p1 = heatmap(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p1 = contour(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p1, "Proportion of rewired links", (0, 1))
 yaxis!(p1, "Proportion of shared links", (0, 1))
 title!(p1, "βos")
 
-p2 = heatmap(links_rewired, shared_links, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p2 = contour(links_rewired, shared_links, ST, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p2, "Proportion of rewired links", (0, 1))
 yaxis!(p2, "Proportion of shared links", (0, 1))
 title!(p2, "βst")
 
-p3 = heatmap(links_rewired, shared_links, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p3 = contour(links_rewired, shared_links, WN, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p3, "Proportion of rewired links", (0, 1))
 yaxis!(p3, "Proportion of shared links", (0, 1))
 title!(p3, "βwn")
 
-p4 = heatmap(links_rewired, shared_links, ST ./ WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p4 = contour(links_rewired, shared_links, ST ./ WN, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p4, "Proportion of rewired links", (0, 1))
 yaxis!(p4, "Proportion of shared links", (0, 1))
 title!(p4, "βst / βwn")

### 2021-09-25 code/experiments/02_constant_connectance.jl

@@ -26,22 +26,22 @@ end
 
 ST = WN .- OS
 
-p1 = heatmap(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p1 = contour(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p1, "Rewiring probability", (0, 1))
 yaxis!(p1, "Species sharing probability", (0, 1))
 title!(p1, "βos")
 
-p2 = heatmap(q_values, p_values, ST, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p2 = contour(q_values, p_values, ST, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p2, "Rewiring probability", (0, 1))
 yaxis!(p2, "Species sharing probability", (0, 1))
 title!(p2, "βst")
 
-p3 = heatmap(q_values, p_values, WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p3 = contour(q_values, p_values, WN, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p3, "Rewiring probability", (0, 1))
 yaxis!(p3, "Species sharing probability", (0, 1))
 title!(p3, "βwn")
 
-p4 = heatmap(q_values, p_values, ST./WN, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p4 = contour(q_values, p_values, ST./WN, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p4, "Rewiring probability", (0, 1))
 yaxis!(p4, "Species sharing probability", (0, 1))
 title!(p4, "βst / βwn")

### 2021-09-25 code/experiments/03_variable_connectance.jl

@@ -16,25 +16,25 @@ OS = S ./ (2A .+ S)
 WN = (S.+U) ./ (2A .+ S .+ U)
 ST = WN .- OS
 
-p1 = heatmap(Co2, p', OS', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p1 = contour(Co2, p', OS', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p1, "Connectance of N₂", (0, 1))
 yaxis!(p1, "Species sharing probability", (0, 1))
 title!(p1, "βos")
 
-p2 = heatmap(Co2, p', ST', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p2 = contour(Co2, p', ST', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p2, "Connectance of N₂", (0, 1))
 yaxis!(p2, "Species sharing probability", (0, 1))
 title!(p2, "βst")
 
-p3 = heatmap(Co2, p', WN', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p3 = contour(Co2, p', WN', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p3, "Connectance of N₂", (0, 1))
 yaxis!(p3, "Species sharing probability", (0, 1))
 title!(p3, "βwn")
 
-p4 = heatmap(Co2, p', (ST./WN)', aspectratio=1, fill=true, lc=:white, clim=(0,1))
+p4 = contour(Co2, p', (ST./WN)', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
 vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
 xaxis!(p4, "Connectance of N₂", (0, 1))
 yaxis!(p4, "Species sharing probability", (0, 1))

### 2021-09-25 code/experiments/04_difference_denominator.jl

@@ -26,22 +26,22 @@ FT = WN .- FR
 
 Δ = OS .- FR
 
-p1 = heatmap(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p1 = contour(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p1, "Rewiring probability", (0, 1))
 yaxis!(p1, "Species sharing probability", (0, 1))
 title!(p1, "βos")
 
-p2 = heatmap(q_values, p_values, FR, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p2 = contour(q_values, p_values, FR, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p2, "Rewiring probability", (0, 1))
 yaxis!(p2, "Species sharing probability", (0, 1))
 title!(p2, "βos (common denominator)")
 
-p3 = heatmap(q_values, p_values, FT, clim=(0, 1), aspectratio=1, fill=true, lc=:white)
+p3 = contour(q_values, p_values, FT, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
 xaxis!(p3, "Rewiring probability", (0, 1))
 yaxis!(p3, "Species sharing probability", (0, 1))
 title!(p3, "βst (common denominator)")
 
-p4 = heatmap(q_values, p_values, OS .- FR, clim=(0, 1), aspectratio=1, fill=true, lc=:white, c=:grayC)
+p4 = contour(q_values, p_values, OS .- FR, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white, c=:grayC)
 xaxis!(p4, "Rewiring probability", (0, 1))
 yaxis!(p4, "Species sharing probability", (0, 1))
 title!(p4, "Difference in rewiring estimate")

### 2021-09-25 code/experiments/05_connectance_ratio.jl

@@ -1,0 +1,87 @@
+# Numerical experiment 5 - we vary the connectance of one of the networks
+# relative to that of the other
+
+q = LinRange(0.0, 1.0, 120)
+p = 0.75
+a = 10.0.^LinRange(0.0, 1.0, 97)'
+
+# The expressions are all calculated in main text
+A = p.^2.0 .* (1.0.-q)
+U = (1.0.+a) .- p.^2.0 .- a.*p.^2.0
+S = (1.0.+a).*(p.^2.0) .- 2.0.*(p.^2.0).*(1.0.-q)
+
+OS = S ./ (2A .+ S)
+WN = (S.+U) ./ (2A .+ S .+ U)
+
+ST = WN .- OS
+
+p1 = contour(log10.(vec(a)), q, OS, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
+xaxis!(p1, "Connectance ratio (log10)", (0, 1))
+yaxis!(p1, "Rewiring probability", (0, 1))
+title!(p1, "βos")
+
+p2 = contour(log10.(vec(a)), q, ST, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
+xaxis!(p2, "Connectance ratio (log10)", (0, 1))
+yaxis!(p2, "Rewiring probability", (0, 1))
+title!(p2, "βst")
+
+p3 = contour(log10.(vec(a)), q, WN, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
+xaxis!(p3, "Connectance ratio (log10)", (0, 1))
+yaxis!(p3, "Rewiring probability", (0, 1))
+title!(p3, "βwn")
+
+p4 = contour(log10.(vec(a)), q, ST./WN, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
+xaxis!(p4, "Connectance ratio (log10)", (0, 1))
+yaxis!(p4, "Rewiring probability", (0, 1))
+title!(p4, "βst / βwn ")
+
+plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
+
+savefig(joinpath(_FIGPATH, "numexp5.png"))
+
+
+
+
+
+
+
+
+
+
+q = 0.15
+p = LinRange(0.0, 1.0, 120)
+a = 10.0.^LinRange(0.0, 1.0, 97)'
+
+# The expressions are all calculated in main text
[…]

### 2021-09-25 code/experiments/XX_connectance_symbolics.jl

@@ -1,0 +1,33 @@
+using Symbolics
+
+# Probability of sharing species
+# Probability of rewiring
+@variables p q
+
+# Connectance of the two networks, c2 = a*c, 0 <= c <= c2 <= 1, a >= 1
+@variables a c
+
+# Links unique to N1 due to turnover
+turnover1 = c - p^2*c
+
+# Links unique to N2 due to turnover
+turnover2 = a*c - p^2*c*a
+
+# Total unique links due to turnover
+turnover = turnover1 + turnover2
+
+# Shareable links
+shareable = p^2*c
+
+# Common shared links
+common = (1-q) * c * p^2
+
+# Rewired shared links
+rewired1 = p^2*c-common
+rewired2 = p^2*a*c-common
+rewired = rewired1 + rewired2
+
+# Partitions as expressions
+A = simplify(common)
+U = simplify(turnover)
+S = simplify(rewired)

### 2021-09-25 code/experiment_01.jl
much simplified numexp1

@@ -1,0 +1,28 @@
+using StatsPlots
+default(; frame=:box, dpi=500, size=(500,500), aspectratio=1, c=:lapaz)
+
+# C - common links
+# R - rewired links
+# T - turnover links
+
+# Number of steps at which we evaluate r and d
+steps = 200
+
+# Values of d (proportion of different links)
+d = LinRange(0.0, 1.0, steps)
+
+# Values of r (proportion of unshared links due to rewiring)
+r = LinRange(0.0, 1.0, steps)
+
+# Components as a function of r and d
+C = 1.0.-d
+R = d .* r'
+T = d .* (1.0 .- r')
+
+OS = R ./ (2.0.*C .+ R)
+WN = (R .+ T) ./ (2.0.*C .+R .+T)
+ST = WN .- OS
+
+contour(r, d, OS, levels=9, fill=true)
+xaxis!((0,1), "Proportion of rewired links")
+yaxis!((0,1), "Proportion of unshared links")

### 2021-09-25 code/main.jl

@@ -6,13 +6,8 @@ using StatsBase
 using Base.Threads
 using Statistics
 
-# This ensures the reproducibility of the results
-Random.seed!(90210)
 theme(:mute)
 default(; frame=:box, dpi=600, size=(500, 500), c=:batlow)
-
-# β-diversity function we use
-beta_t = (a,b,c) -> (b+c)/(2a+b+c)
 
 # Path to store the figures
 _FIGPATH = joinpath("figures")

### 2021-09-26 code/experiment_01.jl
move experiments to Pluto notebooks

@@ -1,28 +1,0 @@
-using StatsPlots
-default(; frame=:box, dpi=500, size=(500,500), aspectratio=1, c=:lapaz)
-
-# C - common links
-# R - rewired links
-# T - turnover links
-
-# Number of steps at which we evaluate r and d
-steps = 200
-
-# Values of d (proportion of different links)
-d = LinRange(0.0, 1.0, steps)
-
-# Values of r (proportion of unshared links due to rewiring)
-r = LinRange(0.0, 1.0, steps)
-
-# Components as a function of r and d
-C = 1.0.-d
-R = d .* r'
-T = d .* (1.0 .- r')
-
-OS = R ./ (2.0.*C .+ R)
-WN = (R .+ T) ./ (2.0.*C .+R .+T)
-ST = WN .- OS
-
-contour(r, d, OS, levels=9, fill=true)
-xaxis!((0,1), "Proportion of rewired links")
-yaxis!((0,1), "Proportion of unshared links")

### 2021-09-26 code/experiments/01_sources_of_variation.jl

@@ -1,63 +1,0 @@
-
-# Numerical experiment 1: L1 = L2, we vary the number of shared, unique, and
-# non-unique links, and examine the consequences on the values of the OS, ST,
-# and WN components of beta diversity. The goal of this experiment is to show
-# that, without considering connectance or species turnover, we can tie the
-# values of these components to the sources of network dissimilarity.
-
-steps = 100
-
-# These are proportions - first, we pick a proportion of links (out of L1 or L2)
-# that are shared, and then we divide the rest of the links in those that are
-# not-shared and rewired (S in the manuscript), and those that are not-shared
-# and involve non-overlapping species (U in the paper). Note that the species
-# richness does not matter here.
-
-shared_links = LinRange(0.0, 1.0, steps)
-links_rewired = LinRange(0.0, 1.0, steps)
-
-# We prepare matrices to fill the calculated values
-OS = zeros(Float64, steps, steps)
-WN = similar(OS)
-
-# Finally, we measure OS and WN - we will get the rest by substraction
-for (i, sl) in enumerate(shared_links)
-    for (j, lr) in enumerate(links_rewired)
-        common = sl
-        rewired = (1.0 - common) * lr
-        turned = 1.0 - (common + rewired)
-        # We can assume that all links are from the same network, since the
-        # measure is symetrical
-        OS[i,j] = beta_t(common, 0.0, rewired)
-        WN[i,j] = beta_t(common, turned, rewired)
-    end
-end
-
-# We get the values of ST
-ST = WN .- OS
-
-# The following lines are plots of the results of numerical experiment 1
-
-p1 = contour(links_rewired, shared_links, OS, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p1, "Proportion of rewired links", (0, 1))
-yaxis!(p1, "Proportion of shared links", (0, 1))
-title!(p1, "βos")
-
-p2 = contour(links_rewired, shared_links, ST, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p2, "Proportion of rewired links", (0, 1))
-yaxis!(p2, "Proportion of shared links", (0, 1))
-title!(p2, "βst")
-
[…]

### 2021-09-26 notebooks/01_components.jl

@@ -1,0 +1,1371 @@
+### A Pluto.jl notebook ###
+# v0.16.1
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ b585acc6-050c-4c80-8ba4-7220c2af5224
+using StatsPlots, Symbolics
+
+# ╔═╡ ed4e6594-1ed9-11ec-3f18-d9c695099d51
+md"""
+# Numerical experiment 1 - importance of rewiring on total link dissimilarity
+
+In this experiment, we will split the origins of network dissimilarity into two components: the total dissimilarity, expressed as the proportion of links that are different between the two networks, and the *relative* importance of rewiring, which is the proportion of different links that are established between common species.
+"""
+
+# ╔═╡ db35b6f1-d487-475f-9199-da5c22420a48
+default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1))
+
+# ╔═╡ 1cc2c524-a46a-447d-8c77-921c71a74422
+md"""
+## Getting the expressions for the partitions from the proportions of different and rewired links
+"""
+
+# ╔═╡ ff0e3e93-f6bd-43bf-9532-387c0f29467e
+# d: proportion of total links that are different
+# r: proportion of different links that are rewired
+@variables d, r
+
+# ╔═╡ cd501c75-500b-44ac-9da2-156407bd3495
+common_links = 1 - d
+
+# ╔═╡ 261e7989-8cac-49ed-ad4d-337e5e15a8d9
+rewired_links = d * r
+
+# ╔═╡ 66baac1a-c473-4912-9ee4-6541f021c147
+turnover_links = d * (1-r)
+
+# ╔═╡ be60bea6-4723-47ff-afd2-4a10436a610d
+md"""
+We can express the values of OS, WN, and ST as a function of these three quantities:
+"""
+
+# ╔═╡ c0af19a3-7d1e-4565-9614-33b5591c2756
+beta_os = rewired_links / (2*common_links + rewired_links)
+
[…]

### 2021-09-26 notebooks/02_speciesoverlap.jl

@@ -1,0 +1,1436 @@
+### A Pluto.jl notebook ###
+# v0.16.1
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ a553ce49-c274-4be5-910c-b3cef97dc2d5
+using StatsPlots, Symbolics
+
+# ╔═╡ 7cfca8be-1ee0-11ec-30be-a58d5339cfb9
+md"""
+# Numerical experiment 2 - effect of rewiring and connectance
+"""
+
+# ╔═╡ 4b071b9e-c85e-445e-a730-b832c1d9b918
+default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1))
+
+# ╔═╡ c5776ce9-9c52-4529-b7be-35a2122570f7
+# p: probability of sharing species
+# q: probability of sharing a link between overlapping species
+# c: connectance of network 1
+# a: multiplicative constant so that the connectance of network 2 is a*c, a >= 1
+@variables p q a c
+
+# ╔═╡ c7696c55-c260-49aa-b964-f303d9502f48
+n1_unique_turnover = 1 - p*p*1
+
+# ╔═╡ b4146a9b-c391-466a-bad9-8545e16997d1
+n2_unique_turnover = a*1 - p*p*a*1
+
+# ╔═╡ d65f7927-13f9-43ea-8bdd-9eeba6923c59
+unique_turnover = n1_unique_turnover + n2_unique_turnover |> simplify
+
+# ╔═╡ 1c65ee48-d1bc-43b5-8f21-9cf7a3ac2e40
+shareable_links = p*p
+
+# ╔═╡ 1395ffe7-7eaa-4aa7-a3e3-0d311ce10749
+shared_shareable_links = q*shareable_links
+
+# ╔═╡ 9d897ef9-f48d-437d-900e-5776d8b48a3c
+n1_rewired_links = 1 - n1_unique_turnover - shared_shareable_links
+
+# ╔═╡ 57a6e5ee-c0b8-4e6c-8a18-c5a0e8350d8c
+n2_rewired_links = a - n2_unique_turnover - shared_shareable_links
+
+# ╔═╡ 65ebea02-113e-4c8d-93b8-91c9faa9c658
+rewired_links = n1_rewired_links + n2_rewired_links |> simplify
+
[…]

### 2021-09-26 notebooks/02_speciesoverlap.jl
bleh

@@ -185,6 +185,9 @@ begin
 	savefig(joinpath(@__DIR__, "..", "figures", "sm_1_fig_2.png"))
 end
 
+# ╔═╡ 074433e6-d5fe-479c-8a3f-f1917a9bcdfa
+(p*(1+a)-p*(1-a))/p |> e -> simplify(e; expand=true)
+
 # ╔═╡ 00000000-0000-0000-0000-000000000001
 PLUTO_PROJECT_TOML_CONTENTS = """
 [deps]
@@ -1432,5 +1435,6 @@ version = "0.9.1+5"
 # ╠═8e38ed48-480a-4851-ae11-b02187d68b30
 # ╠═446b882c-52e8-407d-8f66-69d97d933936
 # ╠═65a84c4c-a746-4ce6-97fa-34707fccb03d
+# ╠═074433e6-d5fe-479c-8a3f-f1917a9bcdfa
 # ╟─00000000-0000-0000-0000-000000000001
 # ╟─00000000-0000-0000-0000-000000000002

### 2021-09-26 code/experiments/02_constant_connectance.jl
🔬 move all code to pluto

@@ -1,51 +1,0 @@
-# Numerical experiment 2 - at some constant connectance, we examine the effect
-# of changing the species overlap between two networks, as well as the
-# proportion that overlapping species keep their interactions. This will provide
-# evidence that the components of the beta diversity partition are responding to
-# the correct ecological processes. Note that, as explained in the manuscript,
-# this calculation is independent of network richness, and does not require any
-# simulations.
-
-Co = 0.4
-p_values = LinRange(0.0, 1.0, 200) # Fraction of shared species
-q_values = LinRange(0.0, 1.0, 202) # Fraction of rewired links
-
-# Stores
-OS = zeros(Float64, length(p_values), length(q_values))
-WN = similar(OS)
-
-for (i, p) in enumerate(p_values)
-    for (j, q) in enumerate(q_values)
-        A = p * (1 - q) * Co
-        S = p * q * Co
-        U = Co * (1 - p) * 2
-        OS[i,j] = S / (2A + S)
-        WN[i,j] = (S + U) / (2A + S + U)
-    end
-end
-
-ST = WN .- OS
-
-p1 = contour(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p1, "Rewiring probability", (0, 1))
-yaxis!(p1, "Species sharing probability", (0, 1))
-title!(p1, "βos")
-
-p2 = contour(q_values, p_values, ST, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p2, "Rewiring probability", (0, 1))
-yaxis!(p2, "Species sharing probability", (0, 1))
-title!(p2, "βst")
-
-p3 = contour(q_values, p_values, WN, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p3, "Rewiring probability", (0, 1))
-yaxis!(p3, "Species sharing probability", (0, 1))
-title!(p3, "βwn")
-
-p4 = contour(q_values, p_values, ST./WN, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p4, "Rewiring probability", (0, 1))
-yaxis!(p4, "Species sharing probability", (0, 1))
-title!(p4, "βst / βwn")
-
-plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
-
-savefig(joinpath(_FIGPATH, "numexp2.png"))

### 2021-09-26 code/experiments/03_variable_connectance.jl

@@ -1,45 +1,0 @@
-# Numerical experiment 3 - we vary the connectance of one of the networks
-# relative to that of the other
-
-Co1 = 0.25
-Co2 = LinRange(0.0, 1.0, 200)
-
-p = LinRange(0.0, 1.0, 204)'
-q = 0.15
-
-# The expressions are all calculated in main text
-A = p.*(1.0.-q).*min.(Co1,Co2)
-U = (1.0.-p).*Co1 .+ (1.0.-p).*Co2
-S = Co1 .- (A .+ (1.0.-p).*Co1) .+ Co2 .- (A .+ (1.0.-p).*Co2)
-
-OS = S ./ (2A .+ S)
-WN = (S.+U) ./ (2A .+ S .+ U)
-ST = WN .- OS
-
-p1 = contour(Co2, p', OS', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
-xaxis!(p1, "Connectance of N₂", (0, 1))
-yaxis!(p1, "Species sharing probability", (0, 1))
-title!(p1, "βos")
-
-p2 = contour(Co2, p', ST', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
-xaxis!(p2, "Connectance of N₂", (0, 1))
-yaxis!(p2, "Species sharing probability", (0, 1))
-title!(p2, "βst")
-
-p3 = contour(Co2, p', WN', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
-xaxis!(p3, "Connectance of N₂", (0, 1))
-yaxis!(p3, "Species sharing probability", (0, 1))
-title!(p3, "βwn")
-
-p4 = contour(Co2, p', (ST./WN)', aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-vline!([Co1], c=:black, lab="", lw=3, ls=:dash)
-xaxis!(p4, "Connectance of N₂", (0, 1))
-yaxis!(p4, "Species sharing probability", (0, 1))
-title!(p4, "βst / βwn")
-
-plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
-
-savefig(joinpath(_FIGPATH, "numexp3.png"))

### 2021-09-26 code/experiments/04_difference_denominator.jl

@@ -1,51 +1,0 @@
-# Numerical experiment 4 - difference in rewiring value
-
-steps = 150
-
-p_values = LinRange(0.0, 1.0, steps) # Fraction of shared species
-q_values = LinRange(0.0, 1.0, steps) # Fraction of rewired links
-
-# Stores
-OS = zeros(Float64, length(p_values), length(q_values))
-FR = similar(OS)
-WN = similar(OS)
-
-for (i, p) in enumerate(p_values)
-    for (j, q) in enumerate(q_values)
-        A = p * (1 - q)
-        S = p * q
-        U = (1 - p) * 2
-        OS[i,j] = S / (2A + S)
-        FR[i,j] = S / (2A + S + U)
-        WN[i,j] = (S+U) / (2A + S + U)
-    end
-end
-
-ST = WN .- OS
-FT = WN .- FR
-
-Δ = OS .- FR
-
-p1 = contour(q_values, p_values, OS, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p1, "Rewiring probability", (0, 1))
-yaxis!(p1, "Species sharing probability", (0, 1))
-title!(p1, "βos")
-
-p2 = contour(q_values, p_values, FR, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p2, "Rewiring probability", (0, 1))
-yaxis!(p2, "Species sharing probability", (0, 1))
-title!(p2, "βos (common denominator)")
-
-p3 = contour(q_values, p_values, FT, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white)
-xaxis!(p3, "Rewiring probability", (0, 1))
-yaxis!(p3, "Species sharing probability", (0, 1))
-title!(p3, "βst (common denominator)")
-
-p4 = contour(q_values, p_values, OS .- FR, clim=(0, 1), aspectratio=1, fill=true, levels=9, lc=:white, c=:grayC)
-xaxis!(p4, "Rewiring probability", (0, 1))
-yaxis!(p4, "Species sharing probability", (0, 1))
-title!(p4, "Difference in rewiring estimate")
-
-plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
-
-savefig(joinpath(_FIGPATH, "numexp4.png"))

### 2021-09-26 code/experiments/05_connectance_ratio.jl

@@ -1,87 +1,0 @@
-# Numerical experiment 5 - we vary the connectance of one of the networks
-# relative to that of the other
-
-q = LinRange(0.0, 1.0, 120)
-p = 0.75
-a = 10.0.^LinRange(0.0, 1.0, 97)'
-
-# The expressions are all calculated in main text
-A = p.^2.0 .* (1.0.-q)
-U = (1.0.+a) .- p.^2.0 .- a.*p.^2.0
-S = (1.0.+a).*(p.^2.0) .- 2.0.*(p.^2.0).*(1.0.-q)
-
-OS = S ./ (2A .+ S)
-WN = (S.+U) ./ (2A .+ S .+ U)
-
-ST = WN .- OS
-
-p1 = contour(log10.(vec(a)), q, OS, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-xaxis!(p1, "Connectance ratio (log10)", (0, 1))
-yaxis!(p1, "Rewiring probability", (0, 1))
-title!(p1, "βos")
-
-p2 = contour(log10.(vec(a)), q, ST, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-xaxis!(p2, "Connectance ratio (log10)", (0, 1))
-yaxis!(p2, "Rewiring probability", (0, 1))
-title!(p2, "βst")
-
-p3 = contour(log10.(vec(a)), q, WN, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-xaxis!(p3, "Connectance ratio (log10)", (0, 1))
-yaxis!(p3, "Rewiring probability", (0, 1))
-title!(p3, "βwn")
-
-p4 = contour(log10.(vec(a)), q, ST./WN, aspectratio=1, fill=true, levels=9, lc=:white, clim=(0,1))
-xaxis!(p4, "Connectance ratio (log10)", (0, 1))
-yaxis!(p4, "Rewiring probability", (0, 1))
-title!(p4, "βst / βwn ")
-
-plot(p1, p2, p3, p4, size=(800, 800), dpi=600)
-
-savefig(joinpath(_FIGPATH, "numexp5.png"))
-
-
-
-
-
-
-
-
-
-
-q = 0.15
-p = LinRange(0.0, 1.0, 120)
-a = 10.0.^LinRange(0.0, 1.0, 97)'
-
-# The expressions are all calculated in main text
[…]

### 2021-09-26 code/experiments/XX_connectance_symbolics.jl

@@ -1,33 +1,0 @@
-using Symbolics
-
-# Probability of sharing species
-# Probability of rewiring
-@variables p q
-
-# Connectance of the two networks, c2 = a*c, 0 <= c <= c2 <= 1, a >= 1
-@variables a c
-
-# Links unique to N1 due to turnover
-turnover1 = c - p^2*c
-
-# Links unique to N2 due to turnover
-turnover2 = a*c - p^2*c*a
-
-# Total unique links due to turnover
-turnover = turnover1 + turnover2
-
-# Shareable links
-shareable = p^2*c
-
-# Common shared links
-common = (1-q) * c * p^2
-
-# Rewired shared links
-rewired1 = p^2*c-common
-rewired2 = p^2*a*c-common
-rewired = rewired1 + rewired2
-
-# Partitions as expressions
-A = simplify(common)
-U = simplify(turnover)
-S = simplify(rewired)

### 2021-09-26 code/main.jl

@@ -1,24 +1,0 @@
-# Dependencies for the simulations
-using StatsPlots
-using PlotThemes
-using Random
-using StatsBase
-using Base.Threads
-using Statistics
-
-theme(:mute)
-default(; frame=:box, dpi=600, size=(500, 500), c=:batlow)
-
-# Path to store the figures
-_FIGPATH = joinpath("figures")
-ispath(_FIGPATH) || mkpath(_FIGPATH)
-
-# Path to store the experiments
-_EXPPATH = joinpath("code", "experiments")
-
-# Run every experiment in the folder
-for f in filter(f -> endswith(f, ".jl"), readdir(_EXPPATH))
-    @info f
-    include(joinpath(_EXPPATH, f))
-end
-

### 2021-09-26 notebooks/03_denominators.jl

@@ -4,96 +4,86 @@
 using Markdown
 using InteractiveUtils
 
-# ╔═╡ a553ce49-c274-4be5-910c-b3cef97dc2d5
+# ╔═╡ 42b6f8c2-1f11-11ec-2372-4134895e6fe2
 using StatsPlots, Symbolics
 
-# ╔═╡ 7cfca8be-1ee0-11ec-30be-a58d5339cfb9
-md"""
-# Numerical experiment 2 - effect of rewiring and connectance
-"""
+# ╔═╡ 67c1480e-6cae-4963-bfc4-6437f81d7953
+# p: probability of sharing a species
+# q: probability of sharing an interaction
+@variables p q
 
-# ╔═╡ 4b071b9e-c85e-445e-a730-b832c1d9b918
-default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1))
+# ╔═╡ d47db712-18a5-4b97-b487-7290c7f94200
+each_turnover = 1-p*p
 
-# ╔═╡ c5776ce9-9c52-4529-b7be-35a2122570f7
-# p: probability of sharing species
-# q: probability of sharing a link between overlapping species
-# c: connectance of network 1
-# a: multiplicative constant so that the connectance of network 2 is a*c, a >= 1
-@variables p q a c
+# ╔═╡ 65ba1d3a-879d-4bba-ad5b-6154a70bf38a
+each_rewired = p*p*(1-q)
 
-# ╔═╡ c7696c55-c260-49aa-b964-f303d9502f48
-n1_unique_turnover = 1 - p*p*1
+# ╔═╡ 992195f0-7629-43de-8151-20b57a878a4b
+each_common = p*p*q
 
-# ╔═╡ b4146a9b-c391-466a-bad9-8545e16997d1
-n2_unique_turnover = a*1 - p*p*a*1
+# ╔═╡ 60fd4b47-7228-4687-8d57-6ff505b930f7
+common_links = each_common
 
-# ╔═╡ d65f7927-13f9-43ea-8bdd-9eeba6923c59
-unique_turnover = n1_unique_turnover + n2_unique_turnover |> simplify
+# ╔═╡ b89551fa-6e7d-4489-866e-06d792218f91
[…]

### 2021-09-26 notebooks/02_speciesoverlap.jl
🔬 common denom analysis

@@ -130,7 +130,7 @@ md"""
 
 # ╔═╡ 7a09953a-eb9e-4e40-9862-8ea2d4e1da0f
 begin
-	_cvalues = LinRange(1.0, 2.0, _pixels)
+	_cvalues = LinRange(1.0, 5.0, _pixels)
 	_spsharing = 0.7
 	cos = broadcast(eval(Bos), _spsharing, _values, _cvalues')
 	cwn = broadcast(eval(Bwn), _spsharing, _values, _cvalues')
@@ -149,33 +149,33 @@ end
 
 # ╔═╡ 1606ef32-64a0-4267-aa33-9b80b63c62d4
 begin
-	cp_os = contour(_values, _cvalues, cos', levels=9, fill=true)
+	cp_os = contour(_values, _cvalues, cos', levels=9, fill=true, aspectratio=1/4)
 	xaxis!((0, 1), "Probability of sharing a link")
-	yaxis!((1, 2), "Difference in connectance")
+	yaxis!((1, 5), "Difference in connectance")
 	title!("\\beta os")
 end
 
 # ╔═╡ 003bf76d-70ab-4cf2-8d1c-5d307058a93f
 begin
-	cp_wn = contour(_values, _cvalues, cwn', levels=9, fill=true)
+	cp_wn = contour(_values, _cvalues, cwn', levels=9, fill=true, aspectratio=1/4)
 	xaxis!((0, 1), "Probability of sharing a link")
-	yaxis!((1, 2), "Difference in connectance")
+	yaxis!((1, 5), "Difference in connectance")
 	title!("\\beta wn")
 end
 
 # ╔═╡ 8e38ed48-480a-4851-ae11-b02187d68b30
 begin
-	cp_st = contour(_values, _cvalues, cst', levels=9, fill=true)
+	cp_st = contour(_values, _cvalues, cst', levels=9, fill=true, aspectratio=1/4)
 	xaxis!((0, 1), "Probability of sharing a link")
-	yaxis!((1, 2), "Difference in connectance")
+	yaxis!((1, 5), "Difference in connectance")
 	title!("\\beta st")
 end
 
 # ╔═╡ 446b882c-52e8-407d-8f66-69d97d933936
 begin
-	cp_ratio = contour(_values, _cvalues, cratio', levels=9, fill=true)
+	cp_ratio = contour(_values, _cvalues, cratio', levels=9, fill=true, aspectratio=1/4)
[…]

### 2021-09-26 notebooks/03_denominators.jl

@@ -7,6 +7,9 @@ using InteractiveUtils
 # ╔═╡ 42b6f8c2-1f11-11ec-2372-4134895e6fe2
 using StatsPlots, Symbolics
 
+# ╔═╡ cfe5378b-02b0-4141-bdcd-35b0c622dc9b
+default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1))
+
 # ╔═╡ 67c1480e-6cae-4963-bfc4-6437f81d7953
 # p: probability of sharing a species
 # q: probability of sharing an interaction
@@ -77,10 +80,10 @@ end
 
 # ╔═╡ 877a65f5-5da3-4e44-93fe-58de51693fab
 begin
-	p_os = contour(_values, _values, osf .- stf, levels=9, fill=true)
+	p_os = contour(_values, _values, osf, levels=9, fill=true)
 	xaxis!((0, 1), "Probability of sharing a link")
 	yaxis!((0, 1), "Probability of sharing a species")
-	title!("\\beta os")
+	title!("\\beta os (common denom.)")
 end
 
 # ╔═╡ 27247554-cc9e-4d62-b6aa-04058dabb856
@@ -96,7 +99,7 @@ begin
 	p_st = contour(_values, _values, stf, levels=9, fill=true)
 	xaxis!((0, 1), "Probability of sharing a link")
 	yaxis!((0, 1), "Probability of sharing a species")
-	title!("\\beta st")
+	title!("\\beta st (common denom.)")
 end
 
 # ╔═╡ 8301f419-ab60-424e-a279-7715d883dae0
@@ -104,7 +107,7 @@ begin
 	p_ratio = contour(_values, _values, stf ./ wn, levels=9, fill=true)
 	xaxis!((0, 1), "Probability of sharing a link")
 	yaxis!((0, 1), "Probability of sharing a species")
-	title!("\\beta st / \\beta wn")
+	title!("\\beta st / \\beta wn (common denom.)")
 end
 
 # ╔═╡ 2824cc03-4d64-4b16-b71c-5ceed924ca62
@@ -1328,6 +1331,7 @@ version = "0.9.1+5"
 
 # ╔═╡ Cell order:
 # ╠═42b6f8c2-1f11-11ec-2372-4134895e6fe2
+# ╠═cfe5378b-02b0-4141-bdcd-35b0c622dc9b
[…]

### 2021-09-27 notebooks/01_components.jl
simplify some code

@@ -12,6 +12,8 @@ md"""
 # Numerical experiment 1 - importance of rewiring on total link dissimilarity
 
 In this experiment, we will split the origins of network dissimilarity into two components: the total dissimilarity, expressed as the proportion of links that are different between the two networks, and the *relative* importance of rewiring, which is the proportion of different links that are established between common species.
+
+The rationale for this decomposition is to illustrate how the measures react to (i) how dissimilar the networks are, in terms of proportion of shared links, and (ii) how the measures react to increased rewiring for a given level of dissimilarity. In the next case studies, we will rely on more straighforward decompositions in species / link sharing.
 """
 
 # ╔═╡ db35b6f1-d487-475f-9199-da5c22420a48
@@ -20,6 +22,8 @@ default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1
 # ╔═╡ 1cc2c524-a46a-447d-8c77-921c71a74422
 md"""
 ## Getting the expressions for the partitions from the proportions of different and rewired links
+
+In this section, we rely on the `Symbolics.jl` package to get solutions from the number of links in the different categories, and from there express the values of $\beta_{os}$ and $\beta_{wn}$:
 """
 
 # ╔═╡ ff0e3e93-f6bd-43bf-9532-387c0f29467e
@@ -45,7 +49,7 @@ We can express the values of OS, WN, and ST as a function of these three quantit
 beta_os = rewired_links / (2*common_links + rewired_links)
 
 # ╔═╡ a4904916-c2be-42e0-923a-1bbf9aca0ac4
-beta_wn = (rewired_links + turnover_links) / (2*common_links + rewired_links + turnover_links) |> simplify
+beta_wn = (rewired_links + turnover_links) / (2*common_links + rewired_links + turnover_links) |> simplify_fractions
 
 # ╔═╡ fc31c00a-ba80-4bab-81a5-5a0f0b7bb97d
 beta_st = beta_wn - beta_os
@@ -1347,7 +1351,7 @@ version = "0.9.1+5"
 # ╟─ed4e6594-1ed9-11ec-3f18-d9c695099d51
 # ╠═b585acc6-050c-4c80-8ba4-7220c2af5224
[…]

### 2021-09-27 notebooks/02_speciesoverlap.jl
update simplification code

@@ -52,13 +52,13 @@ Value of the components
 """
 
 # ╔═╡ 70918f82-d207-430a-a036-3e9d09098a40
-beta_os = rewired_links / (2*shared_shareable_links + rewired_links) |> simplify
+beta_os = rewired_links / (2*shared_shareable_links + rewired_links) |> expand |> simplify_fractions
 
 # ╔═╡ 329c5b67-ecc5-4ae4-bcdb-8728f2bb8d7a
 beta_wn = (rewired_links + unique_turnover) / (2*shared_shareable_links + rewired_links + unique_turnover) |> simplify
 
 # ╔═╡ f3f6ed36-e7a9-477c-aa2c-49491076debf
-beta_st = beta_wn - beta_os
+beta_st = beta_wn - beta_os |> expand
 
 # ╔═╡ 0745c28f-f8d3-4355-9180-88c0e28e3dfc
 beta_ratio = beta_st / beta_wn

### 2021-09-27 notebooks/03_denominators.jl
plots update

@@ -37,7 +37,7 @@ turnover_links = 2*each_turnover
 WN = (rewired_links + turnover_links) / (2*common_links + rewired_links + turnover_links) |> expand
 
 # ╔═╡ 905419a8-9f6d-4e3c-a6bc-eef900bdec16
-OSp = rewired_links / (2*common_links + rewired_links) |> expand
+OSp = rewired_links / (2*common_links + rewired_links) |> expand |> simplify
 
 # ╔═╡ a8276021-01ca-4036-8b1b-fc0c45166ae8
 OSf = rewired_links / (2*common_links + rewired_links + turnover_links) |> expand

### 2021-09-28 notebooks/02_speciesoverlap.jl
update figure paths

@@ -1,1436 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.16.1
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ a553ce49-c274-4be5-910c-b3cef97dc2d5
-using StatsPlots, Symbolics
-
-# ╔═╡ 7cfca8be-1ee0-11ec-30be-a58d5339cfb9
-md"""
-# Numerical experiment 2 - effect of rewiring and connectance
-"""
-
-# ╔═╡ 4b071b9e-c85e-445e-a730-b832c1d9b918
-default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1))
-
-# ╔═╡ c5776ce9-9c52-4529-b7be-35a2122570f7
-# p: probability of sharing species
-# q: probability of sharing a link between overlapping species
-# c: connectance of network 1
-# a: multiplicative constant so that the connectance of network 2 is a*c, a >= 1
-@variables p q a c
-
-# ╔═╡ c7696c55-c260-49aa-b964-f303d9502f48
-n1_unique_turnover = 1 - p*p*1
-
-# ╔═╡ b4146a9b-c391-466a-bad9-8545e16997d1
-n2_unique_turnover = a*1 - p*p*a*1
-
-# ╔═╡ d65f7927-13f9-43ea-8bdd-9eeba6923c59
-unique_turnover = n1_unique_turnover + n2_unique_turnover |> simplify
-
-# ╔═╡ 1c65ee48-d1bc-43b5-8f21-9cf7a3ac2e40
-shareable_links = p*p
-
-# ╔═╡ 1395ffe7-7eaa-4aa7-a3e3-0d311ce10749
-shared_shareable_links = q*shareable_links
-
-# ╔═╡ 9d897ef9-f48d-437d-900e-5776d8b48a3c
-n1_rewired_links = 1 - n1_unique_turnover - shared_shareable_links
-
-# ╔═╡ 57a6e5ee-c0b8-4e6c-8a18-c5a0e8350d8c
-n2_rewired_links = a - n2_unique_turnover - shared_shareable_links
-
-# ╔═╡ 65ebea02-113e-4c8d-93b8-91c9faa9c658
-rewired_links = n1_rewired_links + n2_rewired_links |> simplify
-
[…]

### 2021-09-28 notebooks/sharing_v_rewiring.jl

@@ -4,135 +4,203 @@
 using Markdown
 using InteractiveUtils
 
-# ╔═╡ b585acc6-050c-4c80-8ba4-7220c2af5224
-using StatsPlots, Symbolics
+# ╔═╡ 64c8e293-dc9b-4a47-9017-cbd75794e341
+begin
+	using StatsPlots, Symbolics
+	_pixels = 200
+	_values = LinRange(0.0, 1.0, _pixels)
+	default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1))
+	_EXPNAME = split(split(@__FILE__, '/')[end], ".jl")[begin]
+	_FIGPATH = joinpath(@__DIR__, "..", "figures",  _EXPNAME)
+	ispath(_FIGPATH) || mkpath(_FIGPATH)
+end
 
-# ╔═╡ ed4e6594-1ed9-11ec-3f18-d9c695099d51
+# ╔═╡ 7cfca8be-1ee0-11ec-30be-a58d5339cfb9
 md"""
-# Numerical experiment 1 - importance of rewiring on total link dissimilarity
+# Numerical experiment - effect of species sharing and rewiring
 
-In this experiment, we will split the origins of network dissimilarity into two components: the total dissimilarity, expressed as the proportion of links that are different between the two networks, and the *relative* importance of rewiring, which is the proportion of different links that are established between common species.
+The purpose of this experiment is to show how the decomposition behaves with regards to two core mechanisms: species sharing, and link rewiring. We will assume that the networks are unipartite, have the same number of species $S$, the same connectance $\rho$, and therefore have the same number of links $L = \rho\times S^2$.
 
-The rationale for this decomposition is to illustrate how the measures react to (i) how dissimilar the networks are, in terms of proportion of shared links, and (ii) how the measures react to increased rewiring for a given level of dissimilarity. In the next case studies, we will rely on more straighforward decompositions in species / link sharing.
[…]

### 2021-09-28 notebooks/connectance.jl
results for step 2 to check

@@ -4,140 +4,150 @@
 using Markdown
 using InteractiveUtils
 
-# ╔═╡ 64c8e293-dc9b-4a47-9017-cbd75794e341
+# ╔═╡ 663343c5-0487-4b56-8a74-7471bb3bfec1
 begin
 	using StatsPlots, Symbolics
 	_pixels = 200
 	_values = LinRange(0.0, 1.0, _pixels)
+	_species_sharing = 0.7
 	default(; c=:oslo, lw=0.0, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1), frame=:box)
 	_EXPNAME = split(split(@__FILE__, '/')[end], ".jl")[begin]
 	_FIGPATH = joinpath(@__DIR__, "..", "figures",  _EXPNAME)
 	ispath(_FIGPATH) || mkpath(_FIGPATH)
 end
 
-# ╔═╡ 7cfca8be-1ee0-11ec-30be-a58d5339cfb9
+# ╔═╡ 13a78b04-206b-11ec-3967-97c11148de03
 md"""
-# Numerical experiment - effect of species sharing and rewiring
+# Numerical experiment - connectance gradient
 
-The purpose of this experiment is to show how the decomposition behaves with regards to two core mechanisms: species sharing, and link rewiring. We will assume that the networks are unipartite, have the same number of species $S$, the same connectance $\rho$, and therefore have the same number of links $L = \rho\times S^2$.
+In this numerical experiment, we will reproduce the results of the examination of rewiring/sharing, by relaxing the constraint that the connectance of the two networks is the same, but by maintaining the constraint that they have equal richess. The sole variation in this numerical experiment is therefore that one network has $L_1 = \rho\times a\times S^2$, and the other network has $L_2 = \rho\times S^2$; in other words, $L_1 = a\times L$ and $L_2 = L$.
 
[…]

### 2021-09-28 notebooks/common_denominator.jl
figures updates

@@ -116,6 +116,9 @@ begin
 	savefig(joinpath(@__DIR__, "..", "figures", "sm_3_fig_1.png"))
 end
 
+# ╔═╡ 07218335-8c59-4690-b908-a3bf2eb8d3a3
+
+
 # ╔═╡ 00000000-0000-0000-0000-000000000001
 PLUTO_PROJECT_TOML_CONTENTS = """
 [deps]
@@ -1354,5 +1357,6 @@ version = "0.9.1+5"
 # ╠═7ebced26-071c-41f0-982c-32bbbc7a86e5
 # ╠═8301f419-ab60-424e-a279-7715d883dae0
 # ╠═2824cc03-4d64-4b16-b71c-5ceed924ca62
+# ╠═07218335-8c59-4690-b908-a3bf2eb8d3a3
 # ╟─00000000-0000-0000-0000-000000000001
 # ╟─00000000-0000-0000-0000-000000000002

### 2021-09-28 notebooks/connectance.jl

@@ -9,8 +9,8 @@ begin
 	using StatsPlots, Symbolics
 	_pixels = 200
 	_values = LinRange(0.0, 1.0, _pixels)
-	_species_sharing = 0.7
-	default(; c=:oslo, lw=0.0, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1), frame=:box)
+	_species_sharing = 0.8
+	default(; c=:oslo, lw=0.0, levels=9, aspectratio=1, size=(400, 400), dpi=500, frame=:box, clim=(0,1))
 	_EXPNAME = split(split(@__FILE__, '/')[end], ".jl")[begin]
 	_FIGPATH = joinpath(@__DIR__, "..", "figures",  _EXPNAME)
 	ispath(_FIGPATH) || mkpath(_FIGPATH)
@@ -20,9 +20,9 @@ end
 md"""
 # Numerical experiment - connectance gradient
 
-In this numerical experiment, we will reproduce the results of the examination of rewiring/sharing, by relaxing the constraint that the connectance of the two networks is the same, but by maintaining the constraint that they have equal richess. The sole variation in this numerical experiment is therefore that one network has $L_1 = \rho\times a\times S^2$, and the other network has $L_2 = \rho\times S^2$; in other words, $L_1 = a\times L$ and $L_2 = L$.
+In this numerical experiment, we will reproduce the results of the examination of rewiring/sharing, by relaxing the constraint that the connectance of the two networks must be the same, but by maintaining the constraint that they have equal richness.
 
-As one step of the components calculations involves a $\text{min}$ operation, we will force the constraint that $L_1 \le L_2$, which is to say $0 < a \le 1$. The value of $a$ is the *ratio* of connectances of the two networks.
[…]

### 2021-09-28 notebooks/common_denominator.jl
third numerical experiment

@@ -4,11 +4,24 @@
 using Markdown
 using InteractiveUtils
 
-# ╔═╡ 42b6f8c2-1f11-11ec-2372-4134895e6fe2
-using StatsPlots, Symbolics
+# ╔═╡ 07218335-8c59-4690-b908-a3bf2eb8d3a3
+begin
+	using StatsPlots, Symbolics
+	_pixels = 200
+	_values = LinRange(0.0, 1.0, _pixels)
+	_species_sharing = 0.8
+	default(; c=:oslo, lw=0.0, levels=9, aspectratio=1, size=(400, 400), dpi=500, frame=:box, clim=(0,1))
+	_EXPNAME = split(split(@__FILE__, '/')[end], ".jl")[begin]
+	_FIGPATH = joinpath(@__DIR__, "..", "figures",  _EXPNAME)
+	ispath(_FIGPATH) || mkpath(_FIGPATH)
+end
 
 # ╔═╡ cfe5378b-02b0-4141-bdcd-35b0c622dc9b
-default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1))
+md"""
+# Numerical experiment - consequences of the common denominator
+
+We keep the same constraints as the first numerical experiment, this time focusing on methodological differences as to the numerator.
+"""
 
 # ╔═╡ 67c1480e-6cae-4963-bfc4-6437f81d7953
 # p: probability of sharing a species
@@ -16,71 +29,44 @@ default(; c=:lapaz, levels=9, aspectratio=1, size=(400, 400), dpi=500, clim=(0,1
 @variables p q
 
 # ╔═╡ d47db712-18a5-4b97-b487-7290c7f94200
-each_turnover = 1-p*p
+begin
+	C = q*p^2
+	r = (1-q)*p^2
+	R = 2r
+	t = simplify(1 - C - r; expand=true)
+	T = 2t
+end
+	
 
 # ╔═╡ 65ba1d3a-879d-4bba-ad5b-6154a70bf38a
-each_rewired = p*p*(1-q)
+beta_os = R/(2C+R+T) |> expand |> simplify_fractions
 
 # ╔═╡ 992195f0-7629-43de-8151-20b57a878a4b
-each_common = p*p*q
[…]

### 2021-09-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@4e8ee43f38230d0f7b71ec0d6cdee36478cd10d1 🚀

@@
 al.} 2015; Trøjelsgaard \& Olesen 2016) - this variability motivated the
-emergence of methodology to compare ecological networks, in a way that
-meshes with the usual approaches of comparison of ecological
-communities, \emph{i.e.} \(\beta\)-diversity; although the definiton of
-\(\beta\)-diversity is a contentious topic amongst community ecologists
-(see \emph{e.g.} Tuomisto 2010), the need to understand network
-variability is motivated by the fact that species that make up the
-networks do not react to their environment in the same way, and
-therefore the \(\beta\)-diversity of networks may behave in complex
-ways.
+emergence of methodology to compare ecological networks, including in a
+way that meshes with the core concept for the comparison of ecological
+communities, namely \(\beta\)-diversity (Poisot \emph{et al.} 2012). The
+need to understand network variability through partitioning in
+components equivalent to \(\alpha\), \(\beta\), and \(\gamma\)
+diversities is motivated by the prospect to further integrate the
+analysis of species interactions to the analysis of species
+compositions. Because species that make up the networks do not react to
+their environment in the same way, and because interactions are only
+expressed in subsets of the environments in which species co-occurr, the
+\(\beta\)-diversity of networks may behave in complex ways, and its
+quantification is likely to be ecologically informative.
 
@@
 suggested an approach to \(\beta\)-diversity for ecological networks
-which is based on the comparison of shared and unique links among
-species, and differentiate this sharing of links between common and
-unique species. This framework can be summarized as
-\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that overall
-network dissimilarity (\(\beta_{wn}\)) has a component that can be
-calculated directly from the dissimilarity of interactions between
-shared species (\(\beta_{os}\)), and a component that cannot, the later
-originating in unique species introducing their unique interactions
-(\(\beta_{st}\)). This approach has been widely adopted since its
-publication, with recent examples using it to understand the effect of
-fire on pollination systems (Baronio \emph{et al.} 2021); the impact of
-rewiring on spatio-temporal network dynamics (Campos-Moreno \emph{et
-al.} 2021); the effects of farming on rural and urban landscapes on
-species interactions (Olsson \emph{et al.} 2021); and as a tool to
-estimate the sampling completeness of networks (Souza \emph{et al.}
-2021). It has, similarly, received a number of extensions, including the
-ability to account for interaction strength (Magrach \emph{et al.}
-2017), the ability to handle probabilistic ecological networks (Poisot
-\emph{et al.} 2016), and the integration into the Local Contribution to
-Beta Diversity (Legendre \& De Cáceres 2013) approach to understand how
-environment changes drive network dissimilarity (Poisot \emph{et al.}
-2017).
+which is based on the comparison of the number of shared and unique
+links among species within a pair of networks. Their approach
+differentiates this sharing of links between those established between
+species occurring in both networks, and those established with at least
+one unique species. This framework is expressed as the decomposition
+\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that network
+dissimilarity (\(\beta_{wn}\)) has a component that can be calculated
+directly from the dissimilarity of interactions between shared species
+(\(\beta_{os}\)), and a component that cannot (\(\beta_{st}\)).
+Presumably, the value of these components for a pair of networks can
+generate insights about the mechanisms involved in dissimilarity.
 
-In a recent contribution, Fründ (2021) argues that the calculation of
-network dissimilarity terms as outlined by Poisot \emph{et al.} (2012)
-is incorrect, as it can lead to over-estimating the role of interactions
-between shared species in a network (``rewiring''), and therefore
-underestimate the importance of species turnover across networks. Here,
-I present a more thorough justification of the methodological choices
-for the Poisot \emph{et al.} (2012) method, explain how information
-about species turnover can be extracted from its decomposition, and
-conduct numerical experiments to guide the interpretation of the
-\(\beta\)-diversity values thus obtained. These numerical experiments
-establish three core facts. First, the decomposition responds to the
-correct sources of network variation; second, the decomposition
-adequately captures the relative roles of species turnover and
-interaction rewiring; finally, the decomposition adequately captures the
-role of turnover vs.~non-turnover (like changes in connectance)
-processes. Although the alternative normalization suggested by Fründ
-(2021) is not without its uses, which I discuss in conclusion, it is
-inadequate as a network \(\beta\)-diversity measurement, as it
-introduces many confounding elements that make the interpretation of the
-results more difficult, and should likely not be used as a default.
+This approach has been widely adopted since its publication, with recent
+examples using it to understand the effect of fire on pollination
+systems (Baronio \emph{et al.} 2021); the impact of rewiring on
+spatio-temporal network dynamics (Campos-Moreno \emph{et al.} 2021); the
+effects of farming on rural and urban landscapes on species interactions
+(Olsson \emph{et al.} 2021); the impact of environment gradients on
+multi-trophic metacommunities (\textbf{Ohlmann2018MapImp?}); and as a
+tool to estimate the sampling completeness of networks (Souza \emph{et
+al.} 2021). It has, similarly, received a number of extensions,
+including the ability to account for interaction strength (Magrach
+\emph{et al.} 2017), the ability to handle probabilistic ecological
+networks (Poisot \emph{et al.} 2016), and the integration into the Local
+Contribution to Beta Diversity (Legendre \& De Cáceres 2013) approach to
+understand how environment changes drive network dissimilarity (Poisot
+\emph{et al.} 2017).
+
+This later component originates in unique species introducing their
+unique interactions both between themselves, and with species that are
+common to both networks. CONTINUE WITH BETA ST
+
+Furthermore, much like the definition of \(\beta\)-diversity in all its
+forms is a contentious topic amongst community ecologists (see
+\emph{e.g.} Tuomisto 2010), the \(\beta\)-diversity of networks has been
+submitted to methodological scrutiny over the years. A synthesis of some
+criticisms, related to the correct denominator to use to express the
+proportion of different links, has recently been published (Fründ 2021).
+It argues that the calculation of network dissimilarity terms as
+originally outlined by Poisot \emph{et al.} (2012) is incorrect, as it
+can lead to over-estimating the role of interactions between shared
+species in a network (``rewiring''), and therefore underestimate the
+importance of species turnover across networks. As mist-understanding
+either of these quantities can lead to biased inferences about the
+mechanisms generating network dissimilarity, it is important to assess
+how the values (notably of \(\beta_{os}\), and therefore of
+\(\beta_{st}\)) react to methodological choices.
+
+Here, I present a more thorough justification of the methodological
+choices for the Poisot \emph{et al.} (2012) method, explain how
+information about species turnover can be extracted from its
+decomposition, and conduct numerical experiments to guide the
+interpretation of the \(\beta\)-diversity values thus obtained. These
+numerical experiments establish three core facts. First, the
+decomposition responds to the correct sources of network variation;
+second, the decomposition adequately captures the relative roles of
+species turnover and interaction rewiring; finally, the decomposition
+adequately captures the role of turnover vs.~non-turnover (like changes
+in connectance) processes. Although the alternative normalization
+suggested by Fründ (2021) is not without its uses, which I discuss in
+conclusion, it is inadequate as a network \(\beta\)-diversity
+measurement, as it introduces many confounding elements that make the
+interpretation of the results more difficult, and should likely not be
+used as a default.
 

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@7470a928f270c48fb57fc0b7dd67471648c4edc9 🚀

@@
 
-This later component originates in unique species introducing their
-unique interactions both between themselves, and with species that are
-common to both networks. CONTINUE WITH BETA ST
+Yet, the precise meaning of \(\beta_{st}\), namely the importance of
+species turnover in the overall dissimilarity, has been difficult to
+capture, and a source of confusion for some practitioners. This is not
+particularly surprising, as this component of the decomposition responds
+to unique species introducing their unique interactions both between
+themselves, and with species that are common to both networks. For this
+reason, it is important to come up with guidelines for the
+interpretation of this measure, and how to use it to extract ecological
+insights.
 
@@
 
-Here, I present a more thorough justification of the methodological
-choices for the Poisot \emph{et al.} (2012) method, explain how
-information about species turnover can be extracted from its
-decomposition, and conduct numerical experiments to guide the
-interpretation of the \(\beta\)-diversity values thus obtained. These
+Here, I present a mathematical analysis of the Poisot \emph{et al.}
+(2012) method, explain how information about species turnover and link
+rewiring can be extracted from its decomposition, and conduct numerical
+experiments to guide the interpretation of the \(\beta\)-diversity
+values thus obtained (with a specific focus on \(\beta_{st}\)). These
 numerical experiments establish three core facts. First, the
-decomposition responds to the correct sources of network variation;
-second, the decomposition adequately captures the relative roles of
-species turnover and interaction rewiring; finally, the decomposition
-adequately captures the role of turnover vs.~non-turnover (like changes
-in connectance) processes. Although the alternative normalization
-suggested by Fründ (2021) is not without its uses, which I discuss in
-conclusion, it is inadequate as a network \(\beta\)-diversity
-measurement, as it introduces many confounding elements that make the
-interpretation of the results more difficult, and should likely not be
-used as a default.
+decomposition adequately captures the relative roles of species turnover
+and interaction rewiring; second, the decomposition responds to
+differences in network structure (like connectance) as expected;
+finally, the decomposition more accurately captures rewiring than the
+proposed alternative using a different denominator put forth by Fründ
+(2021).
 

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@3f468889bcd3806cf5f54efed99708a934173124 🚀

@@
 
+\begin{figure}
+\hypertarget{fig:conceptual}{%
+\centering
+\includegraphics{figures/betadiv_response_figure.png}
+\caption{TK}\label{fig:conceptual}
+}
+\end{figure}
+
 Yet, the precise meaning of \(\beta_{st}\), namely the importance of
@@
 to unique species introducing their unique interactions both between
-themselves, and with species that are common to both networks. For this
-reason, it is important to come up with guidelines for the
-interpretation of this measure, and how to use it to extract ecological
-insights.
+themselves, and with species that are common to both networks
+fig.~\ref{fig:concptual}. For this reason, it is important to come up
+with guidelines for the interpretation of this measure, and how to use
+it to extract ecological insights.
 

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@47163354c060412ccb1b62ee807bba3ef6e0340f 🚀

@@
 themselves, and with species that are common to both networks
-fig.~\ref{fig:concptual}. For this reason, it is important to come up
+fig.~\ref{fig:conceptual}. For this reason, it is important to come up
 with guidelines for the interpretation of this measure, and how to use

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@36af1046e0f58287231a230fcb53eec08dcd4896 🚀

@@
 \(b = |X_2 \setminus X_1|\), and \(c = |X_1 \setminus X_2|\), where
-\(|x|\) is the cardinality of set \(x\), and \(\setminus\) is the set
-substraction operation. In the perspective of species composition
-comparison, \(X_1\) and \(X_2\) are the sets of species in either
-community, so that if \(X_1 = \{x, y, z\}\) and
-\(X_2 = \{v, w, x, y\}\), we have \(X_1 \cup X_2 = \{v, w, x, y, z\}\),
-\(X_1 \cap X_2 = \{x, y\}\), \(X_2 \setminus X_1 = \{v, w\}\), and
-\(X_1 \setminus X_2 = \{z\}\). The core message of Koleff \emph{et al.}
-(2003) is that the overwheling majority of measures of
-\(\beta\)-diversity can be re-expressed as functions that operate on the
-cardinality (number of elements) of these sets.
+\(|\dot|\) is the cardinality of set \(\dot\) (the number of elements it
+contains), and \(\setminus\) is the set substraction operation. In the
+perspective of species composition comparison, \(X_1\) and \(X_2\) are
+the sets of species in either community, so that if
+\(X_1 = \{x, y, z\}\) and \(X_2 = \{v, w, x, y\}\), we have
+\(X_1 \cup X_2 = \{v, w, x, y, z\}\), \(X_1 \cap X_2 = \{x, y\}\),
+\(X_2 \setminus X_1 = \{v, w\}\), and \(X_1 \setminus X_2 = \{z\}\). The
+core message of Koleff \emph{et al.} (2003) is that the overwheling
+majority of measures of \(\beta\)-diversity can be re-expressed as
+functions that operate on the cardinality of these sets -- this allows
+to focus on the number of unique and common elements, as outlined in
+fig.~\ref{fig:conceptual}.
 

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@1922cf9aeb188f3736264694e3931b1cc0a7deba 🚀

@@
 overall method by Koleff \emph{et al.} (2003) for species dissimilarity
-baed on presence-absence data. The objects to compare, \(X_1\) and
+based on presence-absence data. The objects to compare, \(X_1\) and
 \(X_2\), are partitioned into three values, \(a = |X_1 \cup X_2|\),

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@3767f8dbef7c29829f310e04a5edeb5fb7ccffb5 🚀

@@
 \includegraphics{figures/betadiv_response_figure.png}
-\caption{TK}\label{fig:conceptual}
+\caption{The dissimilarity of two networks (green and orange) of equal
+richness \(S\) (this also holds for unequal richness) depends on three
+families of interactions: those that are unique because of species
+turnover (in a pale color), those that are unique because of rewiring
+(in a saturated color), and those that are shared (in black). Assuming
+that the chance of sharing a species between the two networks is \(p\),
+then there can be at most \(p^2\times S^2\) shared links -- for this
+reason, overall network dissimilarity (\(\beta_{wn}\)) will have a
+component tied to species turnover, which is
+\(\beta_{st}\).}\label{fig:conceptual}
 }
@@
 definitions. Although ecologists tend to think of networks as their
-adjacency matrix, this representation is far from optimal to get a solid
-understanding of which elements should be counted as part of which set
-when measuring network dissimilarity. For this reason, we need fall back
-on the definition of a graph as a pair of sets, wherein
-\(\mathcal{G} = (V, E)\). These two components \(V\) and \(E\) represent
-vertices (nodes, species) and edges (interactions), where \(V\) is
-specifically a set containing the vertices \(\mathcal{G}\), and \(E\) is
-a set of ordered pairs, in which every pair is composed of two elements
-of \(V\); an element \(\{i,j\}\) in \(E\) indicates that there is an
-interaction \emph{from} species \(i\) to species \(j\) in the network
-\(\mathcal{G}\).
+adjacency matrix (as is presented in fig.~\ref{fig:conceptual}), this
+representation is not optimal to reach a robust understanding of which
+elements should be counted as part of which set when measuring network
+dissimilarity. For this reason, we need fall back on the definition of a
+graph as a pair of sets, wherein \(\mathcal{G} = (V, E)\). These two
+components \(V\) and \(E\) represent vertices (nodes, species) and edges
+(interactions), where \(V\) is specifically a set containing the
+vertices \(\mathcal{G}\), and \(E\) is a set of ordered pairs, in which
+every pair is composed of two elements of \(V\); an element \(\{i,j\}\)
+in \(E\) indicates that there is an interaction \emph{from} species
+\(i\) to species \(j\) in the network \(\mathcal{G}\).
 

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@d5c2f7fb602e23a83bab8f09321a9a8fc7286211 🚀

@@
 \(b = |X_2 \setminus X_1|\), and \(c = |X_1 \setminus X_2|\), where
-\(|\dot|\) is the cardinality of set \(\dot\) (the number of elements it
-contains), and \(\setminus\) is the set substraction operation. In the
-perspective of species composition comparison, \(X_1\) and \(X_2\) are
-the sets of species in either community, so that if
+\(|\cdot|\) is the cardinality of set \(\cdot\) (the number of elements
+it contains), and \(\setminus\) is the set substraction operation. In
+the perspective of species composition comparison, \(X_1\) and \(X_2\)
+are the sets of species in either community, so that if
 \(X_1 = \{x, y, z\}\) and \(X_2 = \{v, w, x, y\}\), we have
@@
 (interactions), where \(V\) is specifically a set containing the
-vertices \(\mathcal{G}\), and \(E\) is a set of ordered pairs, in which
-every pair is composed of two elements of \(V\); an element \(\{i,j\}\)
-in \(E\) indicates that there is an interaction \emph{from} species
-\(i\) to species \(j\) in the network \(\mathcal{G}\).
+vertices of \(\mathcal{G}\), and \(E\) is a set of ordered pairs, in
+which every pair is composed of two elements of \(V\); an element
+\(\{i,j\}\) in \(E\) indicates that there is an interaction \emph{from}
+species \(i\) to species \(j\) in the network \(\mathcal{G}\). The
+adjancency matrix \(\mathbf{A}\) of this network would therefore have a
+non-zero entry at \(A_{ij}\).
 
@@
 
-\[\mathcal{M} = (V_c \cup V_n, E_c \cup E_{sn} \cup E_{un}) \,,\]
+\[\mathcal{N} = (V_c \cup V_n, E_c \cup E_{sn} \cup E_{un}) \,,\]
 
-where \(V_c\) is the set of shared species, \(V_k\) are the species
-belonging only to network \(k\), \(E_c\) are the shared edges, and
-\(E_{sk}\) and \(E_{uk}\) are the interactions unique to \(k\)
-involving, respectively, only species in \(V_c\), and at least one
-species from \(V_k\).
+where \(V_c\) is the set of common species, \(V_m\) and \(V_n\) are the
+species belonging only to network \(m\) and \(n\) (respectivly), \(E_c\)
+are the common edges, and \(E_{sm}\) and \(E_{um}\) are the interactions
+unique to \(k\) involving, respectively, only species in \(V_c\), and at
+least one species from \(V_m\) (the same notation applies for the
+subscript \(_{n}\)).
 
@@
 The decomposition of \(\beta\)-diversity from Poisot \emph{et al.}
-(2012) uses these components to measure \(\beta_{os}\) (the interaction
-dissimilarity between shared species, which Fründ (2021) terms
-``rewiring''), and \(\beta_{wn}\) (the overall dissimilarity including
-non-shared species). We can express the components \(a\), \(b\), and
-\(c\) of Koleff \emph{et al.} (2003) as the cardinality of the following
-sets:
+(2012) uses these components to measure \(\beta_{os}\) (``rewiring''),
+and \(\beta_{wn}\) (the overall dissimilarity including non-shared
+species). We can express the components \(a\), \(b\), and \(c\) of
+Koleff \emph{et al.} (2003) as the cardinality of the following sets:
 
@@
 
-These decompositions are used to perform the calculations of
-\(\beta\)-diversity in the \texttt{EcologicalNetworks.jl} package
-(Banville \emph{et al.} 2021) for Julia, which I use for the following
-numerical experiments.
+It is fundamental to note that these components can be measured entirely
+from the interactions, and that the number of species in either network
+are never directly involved.
+
+In the following sections, I present a series of calculations aimed at
+expressing the values of \(\beta_{os}\), \(\beta_{wn}\), and therefore
+\(\beta_{st}\) as a function of species sharing probability (as a proxy
+for mechanisms generating turnover), and link rewiring probability (as a
+proxy for mechanisms generating differences in interactions among shared
+species). These calculations are done using \texttt{Symbolics.jl}
+(\textbf{Gowda2021HigSym?}), and subsequently transformed in executable
+code for \emph{Julia} (\textbf{Bezanson2017JulFre?}), used to produce
+the figures.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
-Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
-EcologicalNetworks.jl: Two complementary packages for analyzing
-ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
-2721.
-
 \leavevmode\hypertarget{ref-Baronio2021NatFir}{}%

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@89457d2a04a32340d0f87bb6edb050002296d5a5 🚀

@@
 it is easier to understand the effect of turnover by picking a
-dissimilarity measure to work as an exemplar. At this point, Fründ
-(2021) introduce a confusin terminology in their work, stating that
-S\o rensen's and Whittaker's measures of dissimilarity are the same in
-the Koleff \emph{et al.} (2003) framework (they are not; in practice,
-\(\beta_{Sor}=1-\beta_{w}\)), and (ii) noting Whittaker's measure as
-\((b+c)/(2a+b+c)\), which in the Koleff \emph{et al.} (2003) framework
-is, in fact, \(\beta_t\) (Wilson \& Shmida 1984). This does not change
-the overall conclusions as these measures can be re-expressed to
-converge to the same value. For the sake of consistency, I will use
-\(\beta_t\) moving forward; it returns values in \([0,1]\), with \(0\)
-meaning complete similarity, and \(1\) meaning complete dissimilarity.
+dissimilarity measure to work as an exemplar. We will use
+\(\beta = (b+c)/(2a+b+c)\), which in the Koleff \emph{et al.} (2003)
+framework is (Wilson \& Shmida 1984). This measure returns values in
+\([0,1]\), with \(0\) meaning complete similarity, and \(1\) meaning
+complete dissimilarity.
 

### 2021-09-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@defa89250558fde8b6679c4f29cd9ff592e3768b 🚀

@@
 So as to simplify the notation of the following section, I will
-introduce a series of new variables. Let \(A = |E_c|\) be the number of
-links that are identical between networks; \(S = |E_{sn} \cup E_{sm}|\)
-be the number of links that are not shared, but only involve shared
-species (\emph{i.e.} links from \(\mathcal{M}\cup\mathcal{N}\)
-established between species from \(\mathcal{M}\cap\mathcal{N}\)); and
-\(U = |E_{un} \cup E_{um}|\) the number of links that are not shared,
-and involve at least one unique species. Adopting the perspective
-developed in the previous section, wherein networks are sets and the
-measures of \(\beta\)-diversity operates on these sets, highlights the
-conceptual issue in the Fründ (2021) alternative normalization: they are
-using components of the networks that are \emph{not} part of the
-networks being compared.
+introduce a series of new variables. Let \(C = |E_c|\) be the number of
+links that are identical between networks (as a mnemonic, \(C\) stands
+for ``common''); \(R = |E_{sn} \cup E_{sm}|\) be the number of links
+that are not shared, but only involve shared species (\emph{i.e.} links
+from \(\mathcal{M}\cup\mathcal{N}\) established between species from
+\(\mathcal{M}\cap\mathcal{N}\); as a mnemonic, \(R\) stands for
+``rewired''); and \(T = |E_{un} \cup E_{um}|\) the number of links that
+are not shared, and involve at least one unique species (as a mnemonic,
+\(T\) stands for ``turnover'').
 
-There are two important points to note here. First, the number or
-proportion of species that are shared is not involved in the
-calculation. Second, the connectance of either network is not involved
-in the calculation. That all links counted in \emph{e.g.} \(U\) come
-from \(\mathcal{M}\), or that they are evenly distributed between
-\(\mathcal{M}\) and \(\mathcal{N}\), has no impact on the result. This
-is a desirable property of the approach: whatever quantitative value of
-the components of dissimilarity can be interpreted in the light of the
-connectance and species turnover \emph{without} any risk of circularity.
-Therefore the argument of Fründ (2021), whereby the \(\beta_{os}\)
-component should decrease with turnover, and be invariant to
-connectance, does not hold: the very point of the approach is to provide
-measures that can be interpreted in the light of connectance and species
-turnover.
+There are two important points to note here. First, as mentionned
+earlier, the number or proportion of species that are shared is not
+involved in the calculation. Second, the connectance of either network
+is not involved in the calculation. That all links counted in
+\emph{e.g.} \(U\) come from \(\mathcal{M}\), or that they are evenly
+distributed between \(\mathcal{M}\) and \(\mathcal{N}\), has no impact
+on the result. This is a desirable property of the approach: whatever
+quantitative value of the components of dissimilarity can be interpreted
+in the light of the connectance and species turnover \emph{without} any
+risk of circularity; indeed, I present a numerical experiment where
+connectance varies independently later in this manuscript.
 
@@
 
+\textbf{TK} Therefore the argument of Fründ (2021), whereby the
+\(\beta_{os}\) component should decrease with turnover, and be invariant
+to connectance, does not hold: the very point of the approach is to
+provide measures that can be interpreted in the light of connectance and
+species turnover.
+
+\textbf{TK} Adopting the perspective developed in the previous section,
+wherein networks are sets and the measures of \(\beta\)-diversity
+operates on these sets, highlights the conceptual issue in the Fründ
+(2021) alternative normalization: they are using components of the
+networks that are \emph{not} part of the networks being compared.
+
 The choice of changing the denominator hinges on what one admits as a

### 2021-10-04 notebooks/sharing_v_rewiring.jl
🎓 new numexp

@@ -224,9 +224,9 @@ version = "1.16.1+0"
 
 [[ChainRulesCore]]
 deps = ["Compat", "LinearAlgebra", "SparseArrays"]
-git-tree-sha1 = "e8a30e8019a512e4b6c56ccebc065026624660e8"
+git-tree-sha1 = "a325370b9dd0e6bf5656a6f1a7ae80755f8ccc46"
 uuid = "d360d2e6-b24c-11e9-a2a3-2a2ae2dbcce4"
-version = "1.7.0"
+version = "1.7.2"
 
 [[Clustering]]
 deps = ["Distances", "LinearAlgebra", "NearestNeighbors", "Printf", "SparseArrays", "Statistics", "StatsBase"]
@@ -337,9 +337,9 @@ uuid = "8ba89e20-285c-5b6f-9357-94700520ee1b"
 
 [[Distributions]]
 deps = ["ChainRulesCore", "FillArrays", "LinearAlgebra", "PDMats", "Printf", "QuadGK", "Random", "SparseArrays", "SpecialFunctions", "Statistics", "StatsBase", "StatsFuns"]
-git-tree-sha1 = "f4efaa4b5157e0cdb8283ae0b5428bc9208436ed"
+git-tree-sha1 = "a9b99024b57d12fb19892d3f2230856f6d9671a4"
 uuid = "31c24e10-a181-5473-b8eb-7969acd0382f"
-version = "0.25.16"
+version = "0.25.17"
 
 [[DocStringExtensions]]
 deps = ["LibGit2"]
@@ -359,9 +359,9 @@ uuid = "f43a241f-c20a-4ad4-852c-f6b1247861c6"
 
 [[DynamicPolynomials]]
 deps = ["DataStructures", "Future", "LinearAlgebra", "MultivariatePolynomials", "MutableArithmetics", "Pkg", "Reexport", "Test"]
-git-tree-sha1 = "05b68e727a192783be0b34bd8fee8f678505c0bf"
+git-tree-sha1 = "1b4665a7e303eaa7e03542cfaef0730cb056cb00"
 uuid = "7c1d4256-1411-5781-91ec-d7bc3513ac07"
-version = "0.3.20"
+version = "0.3.21"
 
 [[EarCut_jll]]
 deps = ["Artifacts", "JLLWrappers", "Libdl", "Pkg"]
@@ -499,9 +499,9 @@ version = "1.0.2"
 
 [[HTTP]]
[…]

### 2021-10-04 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@300fdc716e0ff11462d9af2047c23b97c09e771d 🚀

@@
 where \(V_c\) is the set of common species, \(V_m\) and \(V_n\) are the
-species belonging only to network \(m\) and \(n\) (respectivly), \(E_c\)
-are the common edges, and \(E_{sm}\) and \(E_{um}\) are the interactions
-unique to \(k\) involving, respectively, only species in \(V_c\), and at
-least one species from \(V_m\) (the same notation applies for the
-subscript \(_{n}\)).
+species belonging only to network \(m\) and \(n\) (respectively),
+\(E_c\) are the common edges, and \(E_{sm}\) and \(E_{um}\) are the
+interactions unique to \(k\) involving, respectively, only species in
+\(V_c\), and at least one species from \(V_m\) (the same notation
+applies for the subscript \(_{n}\)).
 
@@
 \hypertarget{quantifying-the-importance-of-species-turnover}{%
-\subsection{Quantifying the importance of species
+\subsubsection{Quantifying the importance of species
 turnover}\label{quantifying-the-importance-of-species-turnover}}
@@
 complete dissimilarity.
-
-\hypertarget{establishing-that-beta_wn-ge-beta_os}{%
-\subsubsection{\texorpdfstring{Establishing that
-\(\beta_{wn} \ge \beta_{os}\)}{Establishing that \textbackslash beta\_\{wn\} \textbackslash ge \textbackslash beta\_\{os\}}}\label{establishing-that-beta_wn-ge-beta_os}}
 
@@
 risk of circularity; indeed, I present a numerical experiment where
-connectance varies independently later in this manuscript.
+connectance varies independently later in this manuscript, reinforcing
+this point.
 
@@
 (\(\beta_{os}\)), and therefore stems from species turnover. This
-fraction is defined as \(\beta_{st} = \beta_{wn}-\beta_{os}\).
-
-The expression of \(\beta_{st}\) does not involve a partition into sets
-that can be plugged into the framework of Koleff \emph{et al.} (2003),
+fraction is defined as \(\beta_{st} = \beta_{wn}-\beta_{os}\). The
+expression of \(\beta_{st}\) does not involve a partition into sets that
+can be plugged into the framework of Koleff \emph{et al.} (2003),
 because the part of \(\mathcal{M}\) and \(\mathcal{N}\) that are
@@
 
-\[\beta_{os} = \frac{S}{2A+S}\,,\]
+\[\beta_{os} = \frac{R}{2C+R}\,,\]
 
@@
 
-\[\beta_{wn} = \frac{S+U}{2A+S+U}\,.\]
+\[\beta_{wn} = \frac{R+T}{2C+R+T}\,.\]
 
@@
 \(\beta_{wn} \ge \beta_{os}\), and therefore that
-\(0 \ge \beta_{st} \ge 1\); as \(A\), \(S\), and \(U\) are cardinalities
+\(0 \ge \beta_{st} \ge 1\); as \(C\), \(T\), and \(R\) are cardinalities
 of sets, they are necessarily satisfying this condition.
@@
 
-\[\beta_{st} = \frac{2AU}{(2A+S)(2A+S+U)}\,.\]
+\[\beta_{st} = \frac{2CT}{(2C+R)(2C+R+T)}\,.\]
 
@@
 this is obvious when developing the denominator into
+\(4C^2 + R^2 + 4CR + 2CT + RT\). As such, we expect that the value of
+\(\beta_{st}\) will vary in a hump-shaped way with the proportion of
+shared interactions. For this reason, Poisot \emph{et al.} (2012)
+suggest that \(\beta_{st}/\beta{wn}\) (alt. \(1-\beta_{os}/\beta_{wn}\))
+is a better indicator of the \emph{relative} importance of turnover
+processes on network dissimilarity. This can be calculated as
 
-\[4A^2 + S^2 + 4AS + 2AU + SU\,,\]
-
-As such, we expect that the value of \(\beta_{st}\) will vary in a
-hump-shaped way with the proportion of shared interactions. For this
-reason, Poisot \emph{et al.} (2012) suggest that
-\(\beta_{st}/\beta{wn}\) (alt. \(1-\beta_{os}/\beta_{wn}\)) is a better
-indicator of the \emph{relative} importance of turnover processes on
-network dissimilarity. This can be calculated as
-
-\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(2A+S+U)}\times\frac{S+U}{2A+S+U}\,,\]
+\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2CT}{(2C+S)(2C+R+T)}\times\frac{R+T}{2C+R+T}\,,\]
 
@@
 
-\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(S+U)}\,.\]
+\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2CT}{(2C+R)(R+T)}\,.\]
 
-The roots of this expression are \(A=0\) (the turnover of species has no
+The roots of this expression are \(C=0\) (the turnover of species has no
 contribution to the difference between \(\beta_{wn}\) and \(\beta_{os}\)
 if there are no shared species, and therefore no rewiring), and for
-\(U = 0\) (the turnover of species has no contribution if all species
+\(T = 0\) (the turnover of species has no contribution if all species
 are shared).
 
-\hypertarget{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}{%
-\subsubsection{Numerical experiment: response of the components to
-different sources of network
-variation}\label{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}}
+\hypertarget{quantifying-the-response-of-network-beta-diversity-to-souces-of-variation}{%
+\subsection{Quantifying the response of network beta-diversity to souces
+of
+variation}\label{quantifying-the-response-of-network-beta-diversity-to-souces-of-variation}}
 
-To illustrate the behavior of \(\beta_{st}\), I conducted a simple
-numerical experiment in which two networks have the same number of
-interactions \(L\) (recall from the previous section that we do not need
-to set a number of species yet), and these interactions are partitionned
-according to proportions \(p_s\) and \(p_r\) into shared (\(A\)),
-rewired (\(S\)), and unique (\(U\)) links, with \(A = p_s\times L\),
-\(S = (1-p_s)\times p_r\times L\), and
-\(U = (1-p_s)\times (1-p_r)\times L\). The results are represented in
-fig.~\ref{fig:numexp1}.
+\hypertarget{the-relative-effect-of-species-turnover-and-link-rewiring}{%
+\subsubsection{The relative effect of species turnover and link
+rewiring}\label{the-relative-effect-of-species-turnover-and-link-rewiring}}
+
+As the decomposition of beta diversity into sets presented above
+reveals, the value of the components \(\beta_{os}\) and \(\beta_{st}\)
+will respond to two family of mechanisms: the probability of sharing a
+species between the two networks, noted \(p\), which will impose bounds
+on the value of \(T\); and the probability of an interactions between
+shared species \emph{not} being rewired, noted \(q\), which will impose
+bounds on the value of \(C\). These two probabilities represent,
+respectively, mechanisms involved in species turnover and link turnover,
+as per Poisot \emph{et al.} (2015), and the aim of this numerical
+experiment is to describe how these families of processes drive network
+dissimilarity.
+
+In order to simplify the calculations, I make the assumptions that the
+networks have equal species richness (noted \(S\)), so that
+\(S_1 = S_2 = S\), and the same connectance (noted \(\rho\)), so that
+\(\rho_1 = \rho_2 = \rho\). As a consequence, the two networks have the
+same number of links \(L = \rho\times S_1 = \rho\times S_2\). The
+assumption of equal connectance will be relaxed in a subsequent
+numerical experiment. These simplifications allow to express the size of
+\(C\), \(R\), and \(T\) only as functions of \(p\) and \(q\), as they
+would all be multiplied by \(L\), which can therefore be dropped from
+the calculation.
 
@@
 \centering
-\includegraphics{figures/numexp1.png}
+\includegraphics{figures/sharing_v_rewiring/components.png}
 \caption{Values of \(\beta_{os}\), \(\beta_{wn}\), \(\beta_{st}\), and
-\(\beta_{st}/\beta_{wn}\) as a function of the proportion of rewired
-links and the proportion of shared links.}\label{fig:numexp1}
+\(\beta_{st}/\beta_{wn}\) as a function of the probability \(q\) or
+sharing a link (\(x\)-axis), and the probability \(p\) of sharing a
+species (\(y\)-axis). Larger values indicate \emph{more} dissimilarity,
+such that for \(p=q=1\) the dissimilarity as measured by
+\(\beta_{wn}=0\), and for \(p=q=0\) the dissimilarity as measured by
+\(\beta_{wn}=1\).}\label{fig:numexp1}
 }
@@
 
-\hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
-\subsection{Is this decomposition over-estimating the effect of
-``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}
-
-One of the arguments put forth by Fründ (2021) is that the decomposition
-outlined above will overestimate the effect of rewiring; I argue that
-this is based on a misunderstanding of what \(\beta_{st}\) achieves. It
-is paramount to clarify that \(\beta_{st}\) is not a direct measure of
-the importance of turnover: it is a quantification of the relative
-impact of rewiring to overall dissimilarity, which, all non-turnover
-mechanisms being accounted for in the decomposition, can be explained by
-turnover mechanisms. In this section, I present two numerical
-experiments showing (i) that the \(\beta_{os}\) component is in fact an
-accurate measure of rewiring, and (ii) that \(\beta_{st}\) captures the
-consequences of species turnover, and of the interactions brought by
-unique species.
-
-\hypertarget{illustrations-on-arbitrarily-small-networks-are-biased}{%
-\subsubsection{Illustrations on arbitrarily small networks are
-biased}\label{illustrations-on-arbitrarily-small-networks-are-biased}}
-
-We can re-calculate the illustration of Fründ (2021), wherein a pair of
-networks with two shared interactions (\(A = 2\)) receive either an
-interaction in \(S\), in \(U\), or in both:
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-\(A\) & \(S\) & \(U\) & \(\beta_{os}\) & \(\beta_{wn}\) & \(\beta_{st}\)
-& \(\beta_{st}/\beta_{wn}\)\tabularnewline
-\midrule
-\endhead
-2 & 0 & 0 & 0 & 0 & 0 &\tabularnewline
-2 & 1 & 0 & \(1/5\) & \(1/5\) & 0 & 0\tabularnewline
-2 & 0 & 1 & 0 & \(1/5\) & \(1/5\) & 0\tabularnewline
-2 & 1 & 1 & \(1/5\) & \(1/3\) & \(2/15\) & \(2/5\)\tabularnewline
-\bottomrule
-\end{longtable}
-
-The over-estimation argument hinges on the fact that
-\(\beta_{st} < \beta_{os}\) in the last situation (one interaction as
-rewiring, one as turnover). Reaching the conclusion of an overestimation
-from this is based on a mis-interpretation of what \(\beta_{st}\) means.
-The correct interpretation is that, out of the entire network
-dissimilarity, only three-fifths are explained by re-wiring. The fact
-that this fraction is not exactly one-half comes from the fact that the
-Wilson \& Shmida (1984) measure counts shared interactions \emph{twice}
-(\emph{i.e.} it has a \(2A\) term), which over-amplifies the effect of
-shared interactions as the network is really small. Running the same
-calculations with \(A = 10\) gives a relative importance of the turnover
-processes of 47\%, and \(\beta_{st}\) goes to \(1/2\) as \(A/(S+U)\)
-increases. As an additional caveat, the value of \(\beta_{st}\) will
-depend on the measure of beta-diversity used. Measures that do not count
-the shared interaction twice are not going to amplify the effect of
-rewiring.
-
-\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}{%
-\subsubsection{Numerical experiment: the decomposition captures the
-roles of rewiring and turnover
-accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}}
+\hypertarget{sensibility-of-the-decomposition-to-differences-in-connectance}{%
+\subsubsection{Sensibility of the decomposition to differences in
+connectance}\label{sensibility-of-the-decomposition-to-differences-in-connectance}}
 
@@
 denominators that make sense from a network composition point of view,
-succeeds at capturing the relative effect of turnover and rewiring.
+succeeds at capturing the relative effect of turnover and rewiring. \#\#
+Does the partition of network dissimilarity needs a new normalization?
+
+\hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
+\subsubsection{Is this decomposition over-estimating the effect of
+``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}
+
+One of the arguments put forth by Fründ (2021) is that the decomposition
+outlined above will overestimate the effect of rewiring; I argue that
+this is based on a misunderstanding of what \(\beta_{st}\) achieves. It
+is paramount to clarify that \(\beta_{st}\) is not a direct measure of
+the importance of turnover: it is a quantification of the relative
+impact of rewiring to overall dissimilarity, which, all non-turnover
+mechanisms being accounted for in the decomposition, can be explained by
+turnover mechanisms. In this section, I present two numerical
+experiments showing (i) that the \(\beta_{os}\) component is in fact an
+accurate measure of rewiring, and (ii) that \(\beta_{st}\) captures the
+consequences of species turnover, and of the interactions brought by
+unique species.
+
+\hypertarget{illustrations-on-arbitrarily-small-networks-are-biased}{%
+\subsubsection{Illustrations on arbitrarily small networks are
+biased}\label{illustrations-on-arbitrarily-small-networks-are-biased}}
+
+We can re-calculate the illustration of Fründ (2021), wherein a pair of
+networks with two shared interactions (\(A = 2\)) receive either an
+interaction in \(S\), in \(U\), or in both:
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+\(A\) & \(S\) & \(U\) & \(\beta_{os}\) & \(\beta_{wn}\) & \(\beta_{st}\)
+& \(\beta_{st}/\beta_{wn}\)\tabularnewline
+\midrule
+\endhead
+2 & 0 & 0 & 0 & 0 & 0 &\tabularnewline
+2 & 1 & 0 & \(1/5\) & \(1/5\) & 0 & 0\tabularnewline
+2 & 0 & 1 & 0 & \(1/5\) & \(1/5\) & 0\tabularnewline
+2 & 1 & 1 & \(1/5\) & \(1/3\) & \(2/15\) & \(2/5\)\tabularnewline
+\bottomrule
+\end{longtable}
+
+The over-estimation argument hinges on the fact that
+\(\beta_{st} < \beta_{os}\) in the last situation (one interaction as
+rewiring, one as turnover). Reaching the conclusion of an overestimation
+from this is based on a mis-interpretation of what \(\beta_{st}\) means.
+The correct interpretation is that, out of the entire network
+dissimilarity, only three-fifths are explained by re-wiring. The fact
+that this fraction is not exactly one-half comes from the fact that the
+Wilson \& Shmida (1984) measure counts shared interactions \emph{twice}
+(\emph{i.e.} it has a \(2A\) term), which over-amplifies the effect of
+shared interactions as the network is really small. Running the same
+calculations with \(A = 10\) gives a relative importance of the turnover
+processes of 47\%, and \(\beta_{st}\) goes to \(1/2\) as \(A/(S+U)\)
+increases. As an additional caveat, the value of \(\beta_{st}\) will
+depend on the measure of beta-diversity used. Measures that do not count
+the shared interaction twice are not going to amplify the effect of
+rewiring.
+
+Based on the arguments presented above, I do not think the suggestion of
+Fründ (2021) to change the denominator of \(\beta_{os}\) makes sense as
+a default; the strength of the original approach by Poisot \emph{et al.}
+(2012) is indeed that the effect of turnover is based on a rigorous
+definition of networks as graphs (as opposed to networks as matrices),
+in which the induction of vertices from the edgelist being compared
+gives rise to biologically meaningful denominators. The advantage of
+this approach is that at no time does the turnover of species itself (or
+indeed, as shown in many places in this manuscript, the network
+richness), or the connectance of the network, enter into the
+calculation. As such, it is possible to use \(\beta_{os}\) and
+\(\beta_{wn}\) in relationship to these terms, calculated externally (as
+was recently done by \emph{e.g.} Higino \& Poisot 2021), without
+creating circularities.
+
+\textbf{TK} Therefore the argument of Fründ (2021), whereby the
+\(\beta_{os}\) component should decrease with turnover, and be invariant
+to connectance, does not hold: the very point of the approach is to
+provide measures that can be interpreted in the light of connectance and
+species turnover.
+
+\textbf{TK} Adopting the perspective developed in the previous section,
+wherein networks are sets and the measures of \(\beta\)-diversity
+operates on these sets, highlights the conceptual issue in the Fründ
+(2021) alternative normalization: they are using components of the
+networks that are \emph{not} part of the networks being compared.
 
@@
 
-\hypertarget{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}{%
-\subsection{Does the partition of network dissimilarity needs a new
-normalization?}\label{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}}
-
-Based on the arguments presented above, I do not think the suggestion of
-Fründ (2021) to change the denominator of \(\beta_{os}\) makes sense as
-a default; the strength of the original approach by Poisot \emph{et al.}
-(2012) is indeed that the effect of turnover is based on a rigorous
-definition of networks as graphs (as opposed to networks as matrices),
-in which the induction of vertices from the edgelist being compared
-gives rise to biologically meaningful denominators. The advantage of
-this approach is that at no time does the turnover of species itself (or
-indeed, as shown in many places in this manuscript, the network
-richness), or the connectance of the network, enter into the
-calculation. As such, it is possible to use \(\beta_{os}\) and
-\(\beta_{wn}\) in relationship to these terms, calculated externally (as
-was recently done by \emph{e.g.} Higino \& Poisot 2021), without
-creating circularities.
-
-\textbf{TK} Therefore the argument of Fründ (2021), whereby the
-\(\beta_{os}\) component should decrease with turnover, and be invariant
-to connectance, does not hold: the very point of the approach is to
-provide measures that can be interpreted in the light of connectance and
-species turnover.
-
-\textbf{TK} Adopting the perspective developed in the previous section,
-wherein networks are sets and the measures of \(\beta\)-diversity
-operates on these sets, highlights the conceptual issue in the Fründ
-(2021) alternative normalization: they are using components of the
-networks that are \emph{not} part of the networks being compared.
+\hypertarget{measuring-network-beta-diversity-recommendations}{%
+\subsection{Measuring network beta-diversity:
+recommendations}\label{measuring-network-beta-diversity-recommendations}}
 

### 2021-10-04 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@300fdc716e0ff11462d9af2047c23b97c09e771d 🚀

@@
 this is obvious when developing the denominator into
-
-\[4A^2 + S^2 + 4AS + 2AU + SU\,,\]
-
-As such, we expect that the value of \(\beta_{st}\) will vary in a
-hump-shaped way with the proportion of shared interactions. For this
-reason, Poisot \emph{et al.} (2012) suggest that
-\(\beta_{st}/\beta{wn}\) (alt. \(1-\beta_{os}/\beta_{wn}\)) is a better
-indicator of the \emph{relative} importance of turnover processes on
-network dissimilarity. This can be calculated as
-
-\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(2A+S+U)}\times\frac{S+U}{2A+S+U}\,,\]
+\(4C^2 + R^2 + 4CR + 2CT + RT\). As such, we expect that the value of
+\(\beta_{st}\) will vary in a hump-shaped way with the proportion of
+shared interactions. For this reason, Poisot \emph{et al.} (2012)
+suggest that \(\beta_{st}/\beta{wn}\) (alt. \(1-\beta_{os}/\beta_{wn}\))
+is a better indicator of the \emph{relative} importance of turnover
+processes on network dissimilarity. This can be calculated as
+
+\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2CT}{(2C+S)(2C+R+T)}\times\frac{R+T}{2C+R+T}\,,\]
 
@@
 
-\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(S+U)}\,.\]
-
-The roots of this expression are \(A=0\) (the turnover of species has no
+\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2CT}{(2C+R)(R+T)}\,.\]
+
+The roots of this expression are \(C=0\) (the turnover of species has no
 contribution to the difference between \(\beta_{wn}\) and \(\beta_{os}\)
 if there are no shared species, and therefore no rewiring), and for
-\(U = 0\) (the turnover of species has no contribution if all species
+\(T = 0\) (the turnover of species has no contribution if all species
 are shared).
 
-\hypertarget{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}{%
-\subsubsection{Numerical experiment: response of the components to
-different sources of network
-variation}\label{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}}
-
-To illustrate the behavior of \(\beta_{st}\), I conducted a simple
-numerical experiment in which two networks have the same number of
-interactions \(L\) (recall from the previous section that we do not need
-to set a number of species yet), and these interactions are partitionned
-according to proportions \(p_s\) and \(p_r\) into shared (\(A\)),
-rewired (\(S\)), and unique (\(U\)) links, with \(A = p_s\times L\),
-\(S = (1-p_s)\times p_r\times L\), and
-\(U = (1-p_s)\times (1-p_r)\times L\). The results are represented in
-fig.~\ref{fig:numexp1}.
+\hypertarget{quantifying-the-response-of-network-beta-diversity-to-souces-of-variation}{%
+\subsection{Quantifying the response of network beta-diversity to souces
+of
+variation}\label{quantifying-the-response-of-network-beta-diversity-to-souces-of-variation}}
+
+\hypertarget{the-relative-effect-of-species-turnover-and-link-rewiring}{%
+\subsubsection{The relative effect of species turnover and link
+rewiring}\label{the-relative-effect-of-species-turnover-and-link-rewiring}}
+
+As the decomposition of beta diversity into sets presented above
+reveals, the value of the components \(\beta_{os}\) and \(\beta_{st}\)
+will respond to two family of mechanisms: the probability of sharing a
+species between the two networks, noted \(p\), which will impose bounds
+on the value of \(T\); and the probability of an interactions between
+shared species \emph{not} being rewired, noted \(q\), which will impose
+bounds on the value of \(C\). These two probabilities represent,
+respectively, mechanisms involved in species turnover and link turnover,
+as per Poisot \emph{et al.} (2015), and the aim of this numerical
+experiment is to describe how these families of processes drive network
+dissimilarity.
+
+In order to simplify the calculations, I make the assumptions that the
+networks have equal species richness (noted \(S\)), so that
+\(S_1 = S_2 = S\), and the same connectance (noted \(\rho\)), so that
+\(\rho_1 = \rho_2 = \rho\). As a consequence, the two networks have the
+same number of links \(L = \rho\times S_1 = \rho\times S_2\). The
+assumption of equal connectance will be relaxed in a subsequent
+numerical experiment. These simplifications allow to express the size of
+\(C\), \(R\), and \(T\) only as functions of \(p\) and \(q\), as they
+would all be multiplied by \(L\), which can therefore be dropped from
+the calculation.
 
@@
 
+\hypertarget{sensibility-of-the-decomposition-to-differences-in-connectance}{%
+\subsubsection{Sensibility of the decomposition to differences in
+connectance}\label{sensibility-of-the-decomposition-to-differences-in-connectance}}
+
+Consider two bipartite networks, each with \(R\) species on either side,
+and each with the same connectance \(\rho\). We will assume that these
+networks \emph{share} a proportion \(p\) of their species from one side
+(and share all species from the other), and that the interactions
+between these species are undergo rewiring with at a rate \(q\). This is
+sufficient information to calculate the values of \(A\), \(S\), and
+\(U\) required to get the values of \(\beta_{os}\) and \(\beta_{wn}\).
+Note that the simplification of assuming that only species from one side
+can vary is merely for the sake of simplicity, but does not decrease the
+generality of the argument.
+
+Each network will have \(\rho(1-p)R^2\) interactions that are unique due
+to species turnover, and so
+
+\[U = 2\rho(1-p)R^2\,.\]
+
+The part of both networks composed of overlapping species has
+\(\rho pR^2\) interactions, of which \(\rho (1-q) p R^2\) are shared,
+and \(\rho qp R^2\) underwent rewiring. This leads to
+
+\[A = \rho (1-q) p R^2\,,\]
+
+and
+
+\[S = \rho pq R^2\,.\]
+
+Note that we can drop the multiplicative constant \(R^2\), making the
+result independent of the size of the network. Based on these
+components, we can get the values of \(\beta_{os}\) and \(\beta_{wn}\),
+as presented in fig.~\ref{fig:numexp2}.
+
+\begin{figure}
+\hypertarget{fig:numexp2}{%
+\centering
+\includegraphics{figures/numexp2.png}
+\caption{Response of \(\beta_{os}\) and \(\beta_{wn}\), and the
+consequences on \(\beta_{st}\), to changes in rewiring probability
+(\(q\)) and probability of species sharing (\(p\)). As expected,
+\(\beta_{os}\) is not affected by species turnover, but increases with
+the rewiring probability. By contrast, \(\beta_{wn}\) increases when the
+rewiring probability is higher \emph{and} when fewer species are shared.
+This has important consequences for \(\beta_{st}\): its value is
+maximized for low species sharing, and decreases for high rewiring
+probability.}\label{fig:numexp2}
+}
+\end{figure}
+
+The value of \(\beta_{os}\) is entirely unchanged by variations in \(p\)
+(species sharing), and responds \emph{only} to changes in \(q\) (the
+probability of rewiring), whereas as expected, \(\beta_{wn}\) responded
+to changes in both of these parameters: the most dissimilar networks
+have low species sharing (interactions are dissimilar because brought by
+unique species), and high rewiring (shared species do not share
+interactions). The relative changes in \(\beta_{os}\) and \(\beta_{wn}\)
+lead to predictable changes in \(\beta_{st}\): its value is maximized
+when both rewiring \emph{and} species sharing are low. Increasing
+rewiring decreases the impact of species turnover (because, for an equal
+number of interactions, the dissimilarity of interactins in shared
+species contributes more to \(\beta_{wn}\)); increasing the chance of
+sharing species also does decrease \(\beta_{st}\), trivially because
+there is no species turnover anymore. Note that when using the
+correction of \(\beta_{st}/\beta_{wn}\), the effect of species turnover
+is magnified for low probabilities of re-wiring.
+
+In conclusion, this numerical experiment shows that the decomposition as
+initially presented by Poisot \emph{et al.} (2012), \emph{i.e.} using
+denominators that make sense from a network composition point of view,
+succeeds at capturing the relative effect of turnover and rewiring. \#\#
+Does the partition of network dissimilarity needs a new normalization?
+
 \hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
-\subsection{Is this decomposition over-estimating the effect of
+\subsubsection{Is this decomposition over-estimating the effect of
 ``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}
@@
 
-\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}{%
-\subsubsection{Numerical experiment: the decomposition captures the
-roles of rewiring and turnover
-accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}}
-
-Consider two bipartite networks, each with \(R\) species on either side,
-and each with the same connectance \(\rho\). We will assume that these
-networks \emph{share} a proportion \(p\) of their species from one side
-(and share all species from the other), and that the interactions
-between these species are undergo rewiring with at a rate \(q\). This is
-sufficient information to calculate the values of \(A\), \(S\), and
-\(U\) required to get the values of \(\beta_{os}\) and \(\beta_{wn}\).
-Note that the simplification of assuming that only species from one side
-can vary is merely for the sake of simplicity, but does not decrease the
-generality of the argument.
-
-Each network will have \(\rho(1-p)R^2\) interactions that are unique due
-to species turnover, and so
-
-\[U = 2\rho(1-p)R^2\,.\]
-
-The part of both networks composed of overlapping species has
-\(\rho pR^2\) interactions, of which \(\rho (1-q) p R^2\) are shared,
-and \(\rho qp R^2\) underwent rewiring. This leads to
-
-\[A = \rho (1-q) p R^2\,,\]
-
-and
-
-\[S = \rho pq R^2\,.\]
-
-Note that we can drop the multiplicative constant \(R^2\), making the
-result independent of the size of the network. Based on these
-components, we can get the values of \(\beta_{os}\) and \(\beta_{wn}\),
-as presented in fig.~\ref{fig:numexp2}.
-
-\begin{figure}
-\hypertarget{fig:numexp2}{%
-\centering
-\includegraphics{figures/numexp2.png}
-\caption{Response of \(\beta_{os}\) and \(\beta_{wn}\), and the
-consequences on \(\beta_{st}\), to changes in rewiring probability
-(\(q\)) and probability of species sharing (\(p\)). As expected,
-\(\beta_{os}\) is not affected by species turnover, but increases with
-the rewiring probability. By contrast, \(\beta_{wn}\) increases when the
-rewiring probability is higher \emph{and} when fewer species are shared.
-This has important consequences for \(\beta_{st}\): its value is
-maximized for low species sharing, and decreases for high rewiring
-probability.}\label{fig:numexp2}
-}
-\end{figure}
-
-The value of \(\beta_{os}\) is entirely unchanged by variations in \(p\)
-(species sharing), and responds \emph{only} to changes in \(q\) (the
-probability of rewiring), whereas as expected, \(\beta_{wn}\) responded
-to changes in both of these parameters: the most dissimilar networks
-have low species sharing (interactions are dissimilar because brought by
-unique species), and high rewiring (shared species do not share
-interactions). The relative changes in \(\beta_{os}\) and \(\beta_{wn}\)
-lead to predictable changes in \(\beta_{st}\): its value is maximized
-when both rewiring \emph{and} species sharing are low. Increasing
-rewiring decreases the impact of species turnover (because, for an equal
-number of interactions, the dissimilarity of interactins in shared
-species contributes more to \(\beta_{wn}\)); increasing the chance of
-sharing species also does decrease \(\beta_{st}\), trivially because
-there is no species turnover anymore. Note that when using the
-correction of \(\beta_{st}/\beta_{wn}\), the effect of species turnover
-is magnified for low probabilities of re-wiring.
-
-In conclusion, this numerical experiment shows that the decomposition as
-initially presented by Poisot \emph{et al.} (2012), \emph{i.e.} using
-denominators that make sense from a network composition point of view,
-succeeds at capturing the relative effect of turnover and rewiring.
+Based on the arguments presented above, I do not think the suggestion of
+Fründ (2021) to change the denominator of \(\beta_{os}\) makes sense as
+a default; the strength of the original approach by Poisot \emph{et al.}
+(2012) is indeed that the effect of turnover is based on a rigorous
+definition of networks as graphs (as opposed to networks as matrices),
+in which the induction of vertices from the edgelist being compared
+gives rise to biologically meaningful denominators. The advantage of
+this approach is that at no time does the turnover of species itself (or
+indeed, as shown in many places in this manuscript, the network
+richness), or the connectance of the network, enter into the
+calculation. As such, it is possible to use \(\beta_{os}\) and
+\(\beta_{wn}\) in relationship to these terms, calculated externally (as
+was recently done by \emph{e.g.} Higino \& Poisot 2021), without
+creating circularities.
+
+\textbf{TK} Therefore the argument of Fründ (2021), whereby the
+\(\beta_{os}\) component should decrease with turnover, and be invariant
+to connectance, does not hold: the very point of the approach is to
+provide measures that can be interpreted in the light of connectance and
+species turnover.
+
+\textbf{TK} Adopting the perspective developed in the previous section,
+wherein networks are sets and the measures of \(\beta\)-diversity
+operates on these sets, highlights the conceptual issue in the Fründ
+(2021) alternative normalization: they are using components of the
+networks that are \emph{not} part of the networks being compared.
 

### 2021-10-04 notebooks/connectance.jl
🎓 numexp 2

@@ -234,9 +234,9 @@ version = "1.16.1+0"
 
 [[ChainRulesCore]]
 deps = ["Compat", "LinearAlgebra", "SparseArrays"]
-git-tree-sha1 = "e8a30e8019a512e4b6c56ccebc065026624660e8"
+git-tree-sha1 = "a325370b9dd0e6bf5656a6f1a7ae80755f8ccc46"
 uuid = "d360d2e6-b24c-11e9-a2a3-2a2ae2dbcce4"
-version = "1.7.0"
+version = "1.7.2"
 
 [[Clustering]]
 deps = ["Distances", "LinearAlgebra", "NearestNeighbors", "Printf", "SparseArrays", "Statistics", "StatsBase"]
@@ -347,9 +347,9 @@ uuid = "8ba89e20-285c-5b6f-9357-94700520ee1b"
 
 [[Distributions]]
 deps = ["ChainRulesCore", "FillArrays", "LinearAlgebra", "PDMats", "Printf", "QuadGK", "Random", "SparseArrays", "SpecialFunctions", "Statistics", "StatsBase", "StatsFuns"]
-git-tree-sha1 = "f4efaa4b5157e0cdb8283ae0b5428bc9208436ed"
+git-tree-sha1 = "a9b99024b57d12fb19892d3f2230856f6d9671a4"
 uuid = "31c24e10-a181-5473-b8eb-7969acd0382f"
-version = "0.25.16"
+version = "0.25.17"
 
 [[DocStringExtensions]]
 deps = ["LibGit2"]
@@ -369,9 +369,9 @@ uuid = "f43a241f-c20a-4ad4-852c-f6b1247861c6"
 
 [[DynamicPolynomials]]
 deps = ["DataStructures", "Future", "LinearAlgebra", "MultivariatePolynomials", "MutableArithmetics", "Pkg", "Reexport", "Test"]
-git-tree-sha1 = "05b68e727a192783be0b34bd8fee8f678505c0bf"
+git-tree-sha1 = "1b4665a7e303eaa7e03542cfaef0730cb056cb00"
 uuid = "7c1d4256-1411-5781-91ec-d7bc3513ac07"
-version = "0.3.20"
+version = "0.3.21"
 
 [[EarCut_jll]]
 deps = ["Artifacts", "JLLWrappers", "Libdl", "Pkg"]
@@ -509,9 +509,9 @@ version = "1.0.2"
 
 [[HTTP]]
[…]

### 2021-10-04 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@1fb29d938b9c265612a7294ca72ccb9420809de7 🚀

@@
 \(\rho_1 = \rho_2 = \rho\). As a consequence, the two networks have the
-same number of links \(L = \rho\times S_1 = \rho\times S_2\). The
+same number of links \(L = \rho\times S_1^2 = \rho\times S_2^2\). The
 assumption of equal connectance will be relaxed in a subsequent
@@
 \begin{figure}
-\hypertarget{fig:numexp1}{%
+\hypertarget{fig:turnrew}{%
 \centering
@@
 \(\beta_{wn}=0\), and for \(p=q=0\) the dissimilarity as measured by
-\(\beta_{wn}=1\).}\label{fig:numexp1}
+\(\beta_{wn}=1\). As expected, the relative importance of turnover
+(\(\beta_{st}\)) is maximal when there is no rewiring, and when turnover
+increases.}\label{fig:turnrew}
 }
@@
 
-The rewiring component \(\beta_{os}\) varies as a function of the
-proportion of shared links that are rewired; by contrast, \(\beta_{wn}\)
-varies \emph{only} as a function of the proportion of links that are
-shared: that the unshared links are established between common or unique
-species has no effect on overall network dissimilarity. The quadratic
-nature of the denominator for \(\beta_{st}\) is clear here, with a
-maximum reach when there is no re-wiring, and a small number of shared
-links (\emph{i.e.} the networks are almost entirely dissimilar except
-for the links between shared species). Althought the \emph{raw} values
-of \(\beta_{st}\) may seem low, the normalization using
-\(\beta_{st}/\beta_{wn}\) magnifies this effect: its values are indeed
-maximized when the rewiring is lower, \emph{i.e.} all of the network
-variation stems from turnover processes.
+The value of \(C\) is the proportion of shared species \(p^2\), as per
+fig.~\ref{fig:conceptual}, times the proportion of shared links, \(q\),
+giving \(C = qp^2\). Each network has \(r = p^2-(qp^2)\) rewired links,
+which leads to \(R = 2r = 2p^2(1-q)\). Finally, we can get the number of
+unique links in each network \(t\) by substracting \(C+r\) from the
+total number of links (which, since we scale everything by \(L\), is 1),
+yielding \(t = 1 - qp^2 - p^2 + qp^2\), which is \(t = 1-p^2\). The
+total number of unique links due to turnover is \(T = 2t = 2(1-p^2)\).
+It is important to note that \(C\) and \(R\), namely the number of links
+that are kept or rewired, depends on species sharing (\(p\)), as the
+possible size of the overlap between the two networks does, but the
+quantity of links that are different due to turnover does not depends on
+rewiring.
+
+With the values of \(C\), \(R\), and \(T\), we can write
+
+\[\beta_{os} = \frac{2p^2(1-q)}{2p^2q+2p^2(1-q)} = \frac{1-q}{q + 1 -q} =
+(1-q)\,.\]
+
+This is a first noteworthy result: the value of \(\beta_{os}\), in the
+ideal scenario of equal links and richness, is the probability of link
+re-wiring. Because this is true regardless of the value of \(p\)
+(species turnover), this makes \(\beta_{os}\) a strongly ecologically
+informative component.
+
+Similarly, we can write
+
+\[\beta_{wn} = \frac{2p^2(1-q)+2(1-p^2)}{2p^2q + 2p^2(1-q)+2(1-p^2)} = \frac{p^2(1-q)+(1-p^2)}{p^2q+p^2(1-q)+(1-p^2)} = 1-qp^2\,.\]
+
+The overall dissimilarity responds to \(q\) (rewiring) linerarly, and to
+\(p\) quadratically (which is expected assuming unipartite networks, in
+which species are present on both sides).
+
+Expressing \(\beta_{os}\) and \(\beta_{wn}\) as functions of \(p\) and
+\(q\) trivializes the search for the expression of \(\beta_{st}\), which
+is
+
+\[\beta_{st} = 1 - p^2q - 1 + q = q\times(1-p^2)\,.\]
+
+It is worth examining this solution in some detail. \(\beta_{st}\)
+scales linearly with the probability that a link will \emph{not} be
+rewired -- in other words, in a pair of networks for which rewiring is
+important (\(q\) goes to 0), species turnover is going to be a
+\emph{relatively} less important mechanism to dissimilarity.
+\(\beta_{st}\) increases when turnover is important (\(p\) goes to 0),
+and therefore \(\beta_{st}\) represents a \emph{balance} between species
+turnover and link rewiring. These three values, as well as
+\(\beta_{st}/\beta_{wn}\), are represented in fig.~\ref{fig:turnrew}.
 
@@
 
-Consider two bipartite networks, each with \(R\) species on either side,
-and each with the same connectance \(\rho\). We will assume that these
-networks \emph{share} a proportion \(p\) of their species from one side
-(and share all species from the other), and that the interactions
-between these species are undergo rewiring with at a rate \(q\). This is
-sufficient information to calculate the values of \(A\), \(S\), and
-\(U\) required to get the values of \(\beta_{os}\) and \(\beta_{wn}\).
-Note that the simplification of assuming that only species from one side
-can vary is merely for the sake of simplicity, but does not decrease the
-generality of the argument.
+The results presented in fig.~\ref{fig:turnrew} include the strong
+assumption that the two networks have equal connectance. Although the
+range of connectances in nature tends to be very strongly conserved
+within a system, we can relax this assumption, by letting one network
+have more interactions than the other. Note that for the sake of
+notation simplicity, I maintain the constraint that the two networks are
+equally species rich. Therefore, the sole variation in this numerical
+experiment is that one network has \(L_1 = \rho\times a\times S^2\), and
+the other network has \(L_2 = \rho\times S^2\); in other words,
+\(L_1 = a\times L\) and \(L_2 = L\). As one step of the components
+calculations involves a \(\text{min}\) operation, I will add the
+constraint that \(L_1 \le L_2\), which is to say \(0 < a \le 1\). The
+value of \(a\) is the \emph{ratio} of connectances of the two networks,
+and the terms \(S^2\) and \(\rho\) being shared across all factors, they
+will be dropped from the calculations.
 
-Each network will have \(\rho(1-p)R^2\) interactions that are unique due
-to species turnover, and so
+The maximal number of links that can be shared is \(ap^2\) (\emph{i.e.}
+\(\text{min}(p^2, ap^2)\)), as we cannot share more links than are in
+the sparsest of the two networks. Of these, \(q\) are not rewired,
+leading to \(C = aqp^2\). The number of links that are rewired in
+network 1 is the number of its links between shared species minus \(C\),
+\emph{i.e.} \(r_1 = ap^2 - aqp^2 = ap^2(1-q)\), and similarly
+\(r_2 = p^2 - aqp^2 = p^2(1-aq)\), leading to
+\(R = r_1 + r_2 = p^2 \left[a(1-q)+1\right]\). Using the same approach,
+we can get \(t_1 = a(1-p^2)\) and \(t_2 = (1-p^2)\), leading to
+\(T = t_1 + t_2 = (1-p^2)(1+a)\).
 

### 2021-10-04 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@1fb29d938b9c265612a7294ca72ccb9420809de7 🚀

@@
 
-Consider two bipartite networks, each with \(R\) species on either side,
-and each with the same connectance \(\rho\). We will assume that these
-networks \emph{share} a proportion \(p\) of their species from one side
-(and share all species from the other), and that the interactions
-between these species are undergo rewiring with at a rate \(q\). This is
-sufficient information to calculate the values of \(A\), \(S\), and
-\(U\) required to get the values of \(\beta_{os}\) and \(\beta_{wn}\).
-Note that the simplification of assuming that only species from one side
-can vary is merely for the sake of simplicity, but does not decrease the
-generality of the argument.
-
-Each network will have \(\rho(1-p)R^2\) interactions that are unique due
-to species turnover, and so
+The results presented in fig.~\ref{fig:turnrew} include the strong
+assumption that the two networks have equal connectance. Although the
+range of connectances in nature tends to be very strongly conserved
+within a system, we can relax this assumption, by letting one network
+have more interactions than the other. Note that for the sake of
+notation simplicity, I maintain the constraint that the two networks are
+equally species rich. Therefore, the sole variation in this numerical
+experiment is that one network has \(L_1 = \rho\times a\times S^2\), and
+the other network has \(L_2 = \rho\times S^2\); in other words,
+\(L_1 = a\times L\) and \(L_2 = L\). As one step of the components
+calculations involves a \(\text{min}\) operation, I will add the
+constraint that \(L_1 \le L_2\), which is to say \(0 < a \le 1\). The
+value of \(a\) is the \emph{ratio} of connectances of the two networks,
+and the terms \(S^2\) and \(\rho\) being shared across all factors, they
+will be dropped from the calculations.
+
+The maximal number of links that can be shared is \(ap^2\) (\emph{i.e.}
+\(\text{min}(p^2, ap^2)\)), as we cannot share more links than are in
+the sparsest of the two networks. Of these, \(q\) are not rewired,
+leading to \(C = aqp^2\). The number of links that are rewired in
+network 1 is the number of its links between shared species minus \(C\),
+\emph{i.e.} \(r_1 = ap^2 - aqp^2 = ap^2(1-q)\), and similarly
+\(r_2 = p^2 - aqp^2 = p^2(1-aq)\), leading to
+\(R = r_1 + r_2 = p^2 \left[a(1-q)+1\right]\). Using the same approach,
+we can get \(t_1 = a(1-p^2)\) and \(t_2 = (1-p^2)\), leading to
+\(T = t_1 + t_2 = (1-p^2)(1+a)\).
 

### 2021-10-04 notebooks/connectance.jl
🎓 get figure back

@@ -109,7 +109,7 @@ end
 
 # ╔═╡ b074d4e7-7006-438e-abda-e38f94af530e
 begin
-	p_os = contour(_values, _values, rotl90(os), levels=9, fill=true)
+	p_os = contour(_values, _values, os, levels=9, fill=true)
 	xaxis!((0, 1), "Probability of sharing a link")
 	yaxis!((0, 1), "Relative connectance")	
 	title!("\\beta os")
@@ -117,7 +117,7 @@ end
 
 # ╔═╡ a6b0715d-66fc-4d80-828d-bf80b1c36426
 begin
-	p_wn = contour(_values, _values, rotl90(wn), levels=9, fill=true)
+	p_wn = contour(_values, _values, wn, levels=9, fill=true)
 	xaxis!((0, 1), "Probability of sharing a link")
 	yaxis!((0, 1), "Relative connectance")	
 	title!("\\beta wn")
@@ -125,7 +125,7 @@ end
 
 # ╔═╡ 9a2ceb4e-7ebe-4b0e-b942-79f33340ee94
 begin
-	p_st = contour(_values, _values, rotl90(st), levels=9, fill=true)
+	p_st = contour(_values, _values, st, levels=9, fill=true)
 	xaxis!((0, 1), "Probability of sharing a link")
 	yaxis!((0, 1), "Relative connectance")	
 	title!("\\beta st")
@@ -133,7 +133,7 @@ end
 
 # ╔═╡ 855e65e1-a7bb-470d-861b-534b4d80ad17
 begin
-	p_ratio = contour(_values, _values, rotl90(ratio), levels=9, fill=true)
+	p_ratio = contour(_values, _values, ratio, levels=9, fill=true)
 	xaxis!((0, 1), "Probability of sharing a link")
 	yaxis!((0, 1), "Relative connectance")	
 	title!("\\beta st / \\beta wn")

### 2021-10-04 notebooks/empirical_rewiring.jl
📈 empirical rewiring proba

@@ -1,0 +1,1147 @@
+### A Pluto.jl notebook ###
+# v0.16.1
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ 3bbd7c2a-2543-11ec-2e9d-67b4fae64b77
+using StatsPlots, EcologicalNetworks
+
+# ╔═╡ f3bf27fc-4124-40fc-b79f-1c4fabd47c6a
+begin
+	default(; c=:oslo, lw=0.0, levels=9, aspectratio=1, size=(400, 400), dpi=500, frame=:box, clim=(0,1))
+	_EXPNAME = split(split(@__FILE__, '/')[end], ".jl")[begin]
+	_FIGPATH = joinpath(@__DIR__, "..", "figures",  _EXPNAME)
+	ispath(_FIGPATH) || mkpath(_FIGPATH)
+end
+
+# ╔═╡ ec0145fe-bd86-4956-a2c6-749d79b82b4e
+N = simplify.(nz_stream_foodweb())
+
+# ╔═╡ da4168dc-b0d7-49d8-ad6c-58795d030a83
+KGL08(βos(N[1], N[2]))
+
+# ╔═╡ 5fd17750-d3d4-4a8a-83c5-41070639692f
+connectance(N[1])/connectance(N[2])
+
+# ╔═╡ 31b484b4-2b20-434c-a678-db80d17ccb05
+function p_rewiring(m, n)
+	spool = species(m ∩ n)
+	a = connectance(m[spool])/connectance(n[spool])
+	a = a > 1.0 ? 1.0/a : a
+	isnan(a) && return nothing
+	b = KGL08(βos(m,n))
+	q = ((a+1)*(1-b))/(2a)
+	isnan(b) && return nothing
+	isnan(q) && return nothing
+	return [richness(m ∩ n)/richness(m ∪ n), 1-q]
+end
+
+# ╔═╡ c6dddf02-86a7-4ca6-96e8-7e81feaf200b
+rewiring, sharing = Float64[], Float64[]
+
+# ╔═╡ 32884f78-fc06-41a1-86a8-49017e8c2162
+for (i,m) in enumerate(N[1:end-1])
+	for (j,n) in enumerate(N[(i+1):end])
+		pr = p_rewiring(m, n)
+		if !isnothing(pr)
+			push!(rewiring, pr[2])
+			push!(sharing, pr[1])
+		end
+	end
+end	
+
[…]

### 2021-10-04 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@a7617ab9429045dcd289b2570fa69c9f37e4b2a6 🚀

@@
 
-\[U = 2\rho(1-p)R^2\,.\]
+As in the previous section, we can use these values to write
 
-The part of both networks composed of overlapping species has
-\(\rho pR^2\) interactions, of which \(\rho (1-q) p R^2\) are shared,
-and \(\rho qp R^2\) underwent rewiring. This leads to
+\[\beta_{os} = 1 - 2\frac{aq}{1+a}\,,\]
 
-\[A = \rho (1-q) p R^2\,,\]
+\[\beta_{wn} = 1 - 2\frac{ap^2q}{1+a}\,,\]
 
@@
 
-\[S = \rho pq R^2\,.\]
-
-Note that we can drop the multiplicative constant \(R^2\), making the
-result independent of the size of the network. Based on these
-components, we can get the values of \(\beta_{os}\) and \(\beta_{wn}\),
-as presented in fig.~\ref{fig:numexp2}.
+\[\beta_{st} = 2aq\frac{(1-p^2)(1+a)}{a^2 + 2a + 1}\,.\]
 
 \begin{figure}
-\hypertarget{fig:numexp2}{%
+\hypertarget{fig:connectance}{%
 \centering
-\includegraphics{figures/numexp2.png}
-\caption{Response of \(\beta_{os}\) and \(\beta_{wn}\), and the
-consequences on \(\beta_{st}\), to changes in rewiring probability
-(\(q\)) and probability of species sharing (\(p\)). As expected,
-\(\beta_{os}\) is not affected by species turnover, but increases with
-the rewiring probability. By contrast, \(\beta_{wn}\) increases when the
-rewiring probability is higher \emph{and} when fewer species are shared.
-This has important consequences for \(\beta_{st}\): its value is
-maximized for low species sharing, and decreases for high rewiring
-probability.}\label{fig:numexp2}
+\includegraphics{figures/connectance/components.png}
+\caption{Consequences of changing the ratio of connectances between two
+equally species-rich networks on the decomposition of network
+beta-diversity, assuming \(p = 0.8\). Networks with stronger differences
+in connectance will tend to be more similar, because the differences in
+number of links becomes extreme enough that the chances of all the links
+in the sparser network being in the denser network
+increases.}\label{fig:connectance}
 }
@@
 
-The value of \(\beta_{os}\) is entirely unchanged by variations in \(p\)
-(species sharing), and responds \emph{only} to changes in \(q\) (the
-probability of rewiring), whereas as expected, \(\beta_{wn}\) responded
-to changes in both of these parameters: the most dissimilar networks
-have low species sharing (interactions are dissimilar because brought by
-unique species), and high rewiring (shared species do not share
-interactions). The relative changes in \(\beta_{os}\) and \(\beta_{wn}\)
-lead to predictable changes in \(\beta_{st}\): its value is maximized
-when both rewiring \emph{and} species sharing are low. Increasing
-rewiring decreases the impact of species turnover (because, for an equal
-number of interactions, the dissimilarity of interactins in shared
-species contributes more to \(\beta_{wn}\)); increasing the chance of
-sharing species also does decrease \(\beta_{st}\), trivially because
-there is no species turnover anymore. Note that when using the
-correction of \(\beta_{st}/\beta_{wn}\), the effect of species turnover
-is magnified for low probabilities of re-wiring.
+The values of these components are visualized in
+fig.~\ref{fig:connectance}. The introduction of the connectance ratio
+makes these expressions marginally more complex than in the case without
+differences in connectance, but the noteworthy result remains that in
+the presence of differences of connectance, the value of \(\beta_{os}\)
+is still independent from species turnover. In fact, there is an
+important conclusion to be drawn from this expression. The shared
+species component is by definition square, meaning that from an actual
+measurement of \(\beta_{os}\) between two networks for which we know the
+connectance, noted \(\mathbf{b}_{os}\), we can get the probability of
+rewiring by reorganizing the terms of
+\(\mathbf{b}_{os} = 1 - 2aq/(1+a)\) as
 
-In conclusion, this numerical experiment shows that the decomposition as
-initially presented by Poisot \emph{et al.} (2012), \emph{i.e.} using
-denominators that make sense from a network composition point of view,
-succeeds at capturing the relative effect of turnover and rewiring. \#\#
-Does the partition of network dissimilarity needs a new normalization?
+\[q \approx \frac{(1-\mathbf{b}_{os})(a+1)}{2a}\,,\]
+
+which gives the probability of rewiring as \(1-q\); note that this is an
+\emph{approximation}, as it assumes that the connectances of the entire
+network and the connectances of the shared components are the same.
+
+\hypertarget{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}{%
+\subsection{Does the partition of network dissimilarity needs a new
+normalization?}\label{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}}
 
@@
 \begin{figure}
-\hypertarget{fig:numexp3}{%
+\hypertarget{fig:commden}{%
 \centering
-\includegraphics{figures/numexp3.png}
+\includegraphics{figures/common_denominator/components.png}
 \caption{Effects of varying the connectance of the second network
@@
 \(q = 0.15\), and the results are qualitatively robust to changes in
-these parameters.}\label{fig:numexp3}
+these parameters.}\label{fig:commden}
 }

### 2021-10-04 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@a7617ab9429045dcd289b2570fa69c9f37e4b2a6 🚀

@@
 
-\[U = 2\rho(1-p)R^2\,.\]
-
-The part of both networks composed of overlapping species has
-\(\rho pR^2\) interactions, of which \(\rho (1-q) p R^2\) are shared,
-and \(\rho qp R^2\) underwent rewiring. This leads to
-
-\[A = \rho (1-q) p R^2\,,\]
+As in the previous section, we can use these values to write
+
+\[\beta_{os} = 1 - 2\frac{aq}{1+a}\,,\]
+
+\[\beta_{wn} = 1 - 2\frac{ap^2q}{1+a}\,,\]
 
@@
 
-The value of \(\beta_{os}\) is entirely unchanged by variations in \(p\)
-(species sharing), and responds \emph{only} to changes in \(q\) (the
-probability of rewiring), whereas as expected, \(\beta_{wn}\) responded
-to changes in both of these parameters: the most dissimilar networks
-have low species sharing (interactions are dissimilar because brought by
-unique species), and high rewiring (shared species do not share
-interactions). The relative changes in \(\beta_{os}\) and \(\beta_{wn}\)
-lead to predictable changes in \(\beta_{st}\): its value is maximized
-when both rewiring \emph{and} species sharing are low. Increasing
-rewiring decreases the impact of species turnover (because, for an equal
-number of interactions, the dissimilarity of interactins in shared
-species contributes more to \(\beta_{wn}\)); increasing the chance of
-sharing species also does decrease \(\beta_{st}\), trivially because
-there is no species turnover anymore. Note that when using the
-correction of \(\beta_{st}/\beta_{wn}\), the effect of species turnover
-is magnified for low probabilities of re-wiring.
-
-In conclusion, this numerical experiment shows that the decomposition as
-initially presented by Poisot \emph{et al.} (2012), \emph{i.e.} using
-denominators that make sense from a network composition point of view,
-succeeds at capturing the relative effect of turnover and rewiring. \#\#
-Does the partition of network dissimilarity needs a new normalization?
+The values of these components are visualized in
+fig.~\ref{fig:connectance}. The introduction of the connectance ratio
+makes these expressions marginally more complex than in the case without
+differences in connectance, but the noteworthy result remains that in
+the presence of differences of connectance, the value of \(\beta_{os}\)
+is still independent from species turnover. In fact, there is an
+important conclusion to be drawn from this expression. The shared
+species component is by definition square, meaning that from an actual
+measurement of \(\beta_{os}\) between two networks for which we know the
+connectance, noted \(\mathbf{b}_{os}\), we can get the probability of
+rewiring by reorganizing the terms of
+\(\mathbf{b}_{os} = 1 - 2aq/(1+a)\) as
+
+\[q \approx \frac{(1-\mathbf{b}_{os})(a+1)}{2a}\,,\]
+
+which gives the probability of rewiring as \(1-q\); note that this is an
+\emph{approximation}, as it assumes that the connectances of the entire
+network and the connectances of the shared components are the same.
+
+\hypertarget{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}{%
+\subsection{Does the partition of network dissimilarity needs a new
+normalization?}\label{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}}
 

### 2021-10-26 reviews/01_draft.tex
some more replies

--- 2021-10-04 draft.tex
@@
 al.} 2015; Trøjelsgaard \& Olesen 2016) - this variability motivated the
-emergence of methodology to compare ecological networks, including in a
-way that meshes with the core concept for the comparison of ecological
-communities, namely \(\beta\)-diversity (Poisot \emph{et al.} 2012). The
-need to understand network variability through partitioning in
-components equivalent to \(\alpha\), \(\beta\), and \(\gamma\)
-diversities is motivated by the prospect to further integrate the
-analysis of species interactions to the analysis of species
-compositions. Because species that make up the networks do not react to
-their environment in the same way, and because interactions are only
-expressed in subsets of the environments in which species co-occurr, the
-\(\beta\)-diversity of networks may behave in complex ways, and its
-quantification is likely to be ecologically informative.
+emergence of methodology to compare ecological networks, in a way that
+meshes with the usual approaches of comparison of ecological
+communities, \emph{i.e.} \(\beta\)-diversity; although the definiton of
+\(\beta\)-diversity is a contentious topic amongst community ecologists
+(see \emph{e.g.} Tuomisto 2010), the need to understand network
+variability is motivated by the fact that species that make up the
+networks do not react to their environment in the same way, and
+therefore the \(\beta\)-diversity of networks may behave in complex
+ways.
 
@@
 suggested an approach to \(\beta\)-diversity for ecological networks
-which is based on the comparison of the number of shared and unique
-links among species within a pair of networks. Their approach
-differentiates this sharing of links between those established between
-species occurring in both networks, and those established with at least
-one unique species. This framework is expressed as the decomposition
-\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that network
-dissimilarity (\(\beta_{wn}\)) has a component that can be calculated
-directly from the dissimilarity of interactions between shared species
-(\(\beta_{os}\)), and a component that cannot (\(\beta_{st}\)).
-Presumably, the value of these components for a pair of networks can
-generate insights about the mechanisms involved in dissimilarity.
+which is based on the comparison of shared and unique links among
+species, and differentiate this sharing of links between common and
+unique species. This framework can be summarized as
+\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that overall
+network dissimilarity (\(\beta_{wn}\)) has a component that can be
+calculated directly from the dissimilarity of interactions between
+shared species (\(\beta_{os}\)), and a component that cannot, the later
+originating in unique species introducing their unique interactions
+(\(\beta_{st}\)). This approach has been widely adopted since its
+publication, with recent examples using it to understand the effect of
+fire on pollination systems (Baronio \emph{et al.} 2021); the impact of
+rewiring on spatio-temporal network dynamics (Campos-Moreno \emph{et
+al.} 2021); the effects of farming on rural and urban landscapes on
+species interactions (Olsson \emph{et al.} 2021); and as a tool to
+estimate the sampling completeness of networks (Souza \emph{et al.}
+2021). It has, similarly, received a number of extensions, including the
+ability to account for interaction strength (Magrach \emph{et al.}
+2017), the ability to handle probabilistic ecological networks (Poisot
+\emph{et al.} 2016), and the integration into the Local Contribution to
+Beta Diversity (Legendre \& De Cáceres 2013) approach to understand how
+environment changes drive network dissimilarity (Poisot \emph{et al.}
+2017).
 
-This approach has been widely adopted since its publication, with recent
-examples using it to understand the effect of fire on pollination
-systems (Baronio \emph{et al.} 2021); the impact of rewiring on
-spatio-temporal network dynamics (Campos-Moreno \emph{et al.} 2021); the
-effects of farming on rural and urban landscapes on species interactions
-(Olsson \emph{et al.} 2021); the impact of environment gradients on
-multi-trophic metacommunities (\textbf{Ohlmann2018MapImp?}); and as a
-tool to estimate the sampling completeness of networks (Souza \emph{et
-al.} 2021). It has, similarly, received a number of extensions,
-including the ability to account for interaction strength (Magrach
-\emph{et al.} 2017), the ability to handle probabilistic ecological
-networks (Poisot \emph{et al.} 2016), and the integration into the Local
-Contribution to Beta Diversity (Legendre \& De Cáceres 2013) approach to
-understand how environment changes drive network dissimilarity (Poisot
-\emph{et al.} 2017).
-
-\begin{figure}
-\hypertarget{fig:conceptual}{%
-\centering
-\includegraphics{figures/betadiv_response_figure.png}
-\caption{The dissimilarity of two networks (green and orange) of equal
-richness \(S\) (this also holds for unequal richness) depends on three
-families of interactions: those that are unique because of species
-turnover (in a pale color), those that are unique because of rewiring
-(in a saturated color), and those that are shared (in black). Assuming
-that the chance of sharing a species between the two networks is \(p\),
-then there can be at most \(p^2\times S^2\) shared links -- for this
-reason, overall network dissimilarity (\(\beta_{wn}\)) will have a
-component tied to species turnover, which is
-\(\beta_{st}\).}\label{fig:conceptual}
-}
-\end{figure}
-
-Yet, the precise meaning of \(\beta_{st}\), namely the importance of
-species turnover in the overall dissimilarity, has been difficult to
-capture, and a source of confusion for some practitioners. This is not
-particularly surprising, as this component of the decomposition responds
-to unique species introducing their unique interactions both between
-themselves, and with species that are common to both networks
-fig.~\ref{fig:conceptual}. For this reason, it is important to come up
-with guidelines for the interpretation of this measure, and how to use
-it to extract ecological insights.
-
-Furthermore, much like the definition of \(\beta\)-diversity in all its
-forms is a contentious topic amongst community ecologists (see
-\emph{e.g.} Tuomisto 2010), the \(\beta\)-diversity of networks has been
-submitted to methodological scrutiny over the years. A synthesis of some
-criticisms, related to the correct denominator to use to express the
-proportion of different links, has recently been published (Fründ 2021).
-It argues that the calculation of network dissimilarity terms as
-originally outlined by Poisot \emph{et al.} (2012) is incorrect, as it
-can lead to over-estimating the role of interactions between shared
-species in a network (``rewiring''), and therefore underestimate the
-importance of species turnover across networks. As mist-understanding
-either of these quantities can lead to biased inferences about the
-mechanisms generating network dissimilarity, it is important to assess
-how the values (notably of \(\beta_{os}\), and therefore of
-\(\beta_{st}\)) react to methodological choices.
-
-Here, I present a mathematical analysis of the Poisot \emph{et al.}
-(2012) method, explain how information about species turnover and link
-rewiring can be extracted from its decomposition, and conduct numerical
-experiments to guide the interpretation of the \(\beta\)-diversity
-values thus obtained (with a specific focus on \(\beta_{st}\)). These
-numerical experiments establish three core facts. First, the
-decomposition adequately captures the relative roles of species turnover
-and interaction rewiring; second, the decomposition responds to
-differences in network structure (like connectance) as expected;
-finally, the decomposition more accurately captures rewiring than the
-proposed alternative using a different denominator put forth by Fründ
-(2021).
+In a recent contribution, Fründ (2021) argues that the calculation of
+network dissimilarity terms as outlined by Poisot \emph{et al.} (2012)
+is incorrect, as it can lead to over-estimating the role of interactions
+between shared species in a network (``rewiring''), and therefore
+underestimate the importance of species turnover across networks. Here,
+I present a more thorough justification of the methodological choices
+for the Poisot \emph{et al.} (2012) method, explain how information
+about species turnover can be extracted from its decomposition, and
+conduct numerical experiments to guide the interpretation of the
+\(\beta\)-diversity values thus obtained. These numerical experiments
+establish three core facts. First, the decomposition responds to the
+correct sources of network variation; second, the decomposition
+adequately captures the relative roles of species turnover and
+interaction rewiring; finally, the decomposition adequately captures the
+role of turnover vs.~non-turnover (like changes in connectance)
+processes. Although the alternative normalization suggested by Fründ
+(2021) is not without its uses, which I discuss in conclusion, it is
+inadequate as a network \(\beta\)-diversity measurement, as it
+introduces many confounding elements that make the interpretation of the
+results more difficult, and should likely not be used as a default.
 
@@
 overall method by Koleff \emph{et al.} (2003) for species dissimilarity
-based on presence-absence data. The objects to compare, \(X_1\) and
+baed on presence-absence data. The objects to compare, \(X_1\) and
 \(X_2\), are partitioned into three values, \(a = |X_1 \cup X_2|\),
 \(b = |X_2 \setminus X_1|\), and \(c = |X_1 \setminus X_2|\), where
-\(|\cdot|\) is the cardinality of set \(\cdot\) (the number of elements
-it contains), and \(\setminus\) is the set substraction operation. In
-the perspective of species composition comparison, \(X_1\) and \(X_2\)
-are the sets of species in either community, so that if
-\(X_1 = \{x, y, z\}\) and \(X_2 = \{v, w, x, y\}\), we have
-\(X_1 \cup X_2 = \{v, w, x, y, z\}\), \(X_1 \cap X_2 = \{x, y\}\),
-\(X_2 \setminus X_1 = \{v, w\}\), and \(X_1 \setminus X_2 = \{z\}\). The
-core message of Koleff \emph{et al.} (2003) is that the overwheling
-majority of measures of \(\beta\)-diversity can be re-expressed as
-functions that operate on the cardinality of these sets -- this allows
-to focus on the number of unique and common elements, as outlined in
-fig.~\ref{fig:conceptual}.
+\(|x|\) is the cardinality of set \(x\), and \(\setminus\) is the set
+substraction operation. In the perspective of species composition
+comparison, \(X_1\) and \(X_2\) are the sets of species in either
+community, so that if \(X_1 = \{x, y, z\}\) and
+\(X_2 = \{v, w, x, y\}\), we have \(X_1 \cup X_2 = \{v, w, x, y, z\}\),
+\(X_1 \cap X_2 = \{x, y\}\), \(X_2 \setminus X_1 = \{v, w\}\), and
+\(X_1 \setminus X_2 = \{z\}\). The core message of Koleff \emph{et al.}
+(2003) is that the overwheling majority of measures of
+\(\beta\)-diversity can be re-expressed as functions that operate on the
+cardinality (number of elements) of these sets.
 
@@
 definitions. Although ecologists tend to think of networks as their
-adjacency matrix (as is presented in fig.~\ref{fig:conceptual}), this
-representation is not optimal to reach a robust understanding of which
-elements should be counted as part of which set when measuring network
-dissimilarity. For this reason, we need fall back on the definition of a
-graph as a pair of sets, wherein \(\mathcal{G} = (V, E)\). These two
-components \(V\) and \(E\) represent vertices (nodes, species) and edges
-(interactions), where \(V\) is specifically a set containing the
-vertices of \(\mathcal{G}\), and \(E\) is a set of ordered pairs, in
-which every pair is composed of two elements of \(V\); an element
-\(\{i,j\}\) in \(E\) indicates that there is an interaction \emph{from}
-species \(i\) to species \(j\) in the network \(\mathcal{G}\). The
-adjancency matrix \(\mathbf{A}\) of this network would therefore have a
-non-zero entry at \(A_{ij}\).
+adjacency matrix, this representation is far from optimal to get a solid
+understanding of which elements should be counted as part of which set
+when measuring network dissimilarity. For this reason, we need fall back
+on the definition of a graph as a pair of sets, wherein
+\(\mathcal{G} = (V, E)\). These two components \(V\) and \(E\) represent
+vertices (nodes, species) and edges (interactions), where \(V\) is
+specifically a set containing the vertices \(\mathcal{G}\), and \(E\) is
+a set of ordered pairs, in which every pair is composed of two elements
+of \(V\); an element \(\{i,j\}\) in \(E\) indicates that there is an
+interaction \emph{from} species \(i\) to species \(j\) in the network
+\(\mathcal{G}\).
 
@@
 
-\[\mathcal{N} = (V_c \cup V_n, E_c \cup E_{sn} \cup E_{un}) \,,\]
+\[\mathcal{M} = (V_c \cup V_n, E_c \cup E_{sn} \cup E_{un}) \,,\]
 
-where \(V_c\) is the set of common species, \(V_m\) and \(V_n\) are the
-species belonging only to network \(m\) and \(n\) (respectively),
-\(E_c\) are the common edges, and \(E_{sm}\) and \(E_{um}\) are the
-interactions unique to \(k\) involving, respectively, only species in
-\(V_c\), and at least one species from \(V_m\) (the same notation
-applies for the subscript \(_{n}\)).
+where \(V_c\) is the set of shared species, \(V_k\) are the species
+belonging only to network \(k\), \(E_c\) are the shared edges, and
+\(E_{sk}\) and \(E_{uk}\) are the interactions unique to \(k\)
+involving, respectively, only species in \(V_c\), and at least one
+species from \(V_k\).
 
@@
 The decomposition of \(\beta\)-diversity from Poisot \emph{et al.}
-(2012) uses these components to measure \(\beta_{os}\) (``rewiring''),
-and \(\beta_{wn}\) (the overall dissimilarity including non-shared
-species). We can express the components \(a\), \(b\), and \(c\) of
-Koleff \emph{et al.} (2003) as the cardinality of the following sets:
+(2012) uses these components to measure \(\beta_{os}\) (the interaction
+dissimilarity between shared species, which Fründ (2021) terms
+``rewiring''), and \(\beta_{wn}\) (the overall dissimilarity including
+non-shared species). We can express the components \(a\), \(b\), and
+\(c\) of Koleff \emph{et al.} (2003) as the cardinality of the following
+sets:
 
@@
 
-It is fundamental to note that these components can be measured entirely
-from the interactions, and that the number of species in either network
-are never directly involved.
-
-In the following sections, I present a series of calculations aimed at
-expressing the values of \(\beta_{os}\), \(\beta_{wn}\), and therefore
-\(\beta_{st}\) as a function of species sharing probability (as a proxy
-for mechanisms generating turnover), and link rewiring probability (as a
-proxy for mechanisms generating differences in interactions among shared
-species). These calculations are done using \texttt{Symbolics.jl}
-(\textbf{Gowda2021HigSym?}), and subsequently transformed in executable
-code for \emph{Julia} (\textbf{Bezanson2017JulFre?}), used to produce
-the figures.
+These decompositions are used to perform the calculations of
+\(\beta\)-diversity in the \texttt{EcologicalNetworks.jl} package
+(Banville \emph{et al.} 2021) for Julia, which I use for the following
+numerical experiments.
 
 \hypertarget{quantifying-the-importance-of-species-turnover}{%
-\subsubsection{Quantifying the importance of species
+\subsection{Quantifying the importance of species
 turnover}\label{quantifying-the-importance-of-species-turnover}}
@@
 it is easier to understand the effect of turnover by picking a
-dissimilarity measure to work as an exemplar. We will use
-\(\beta = (b+c)/(2a+b+c)\), which in the Koleff \emph{et al.} (2003)
-framework is (Wilson \& Shmida 1984). This measure returns values in
-\([0,1]\), with \(0\) meaning complete similarity, and \(1\) meaning
-complete dissimilarity.
+dissimilarity measure to work as an exemplar. At this point, Fründ
+(2021) introduce a confusin terminology in their work, stating that
+S\o rensen's and Whittaker's measures of dissimilarity are the same in
+the Koleff \emph{et al.} (2003) framework (they are not; in practice,
+\(\beta_{Sor}=1-\beta_{w}\)), and (ii) noting Whittaker's measure as
+\((b+c)/(2a+b+c)\), which in the Koleff \emph{et al.} (2003) framework
+is, in fact, \(\beta_t\) (Wilson \& Shmida 1984). This does not change
+the overall conclusions as these measures can be re-expressed to
+converge to the same value. For the sake of consistency, I will use
+\(\beta_t\) moving forward; it returns values in \([0,1]\), with \(0\)
+meaning complete similarity, and \(1\) meaning complete dissimilarity.
+
+\hypertarget{establishing-that-beta_wn-ge-beta_os}{%
+\subsubsection{\texorpdfstring{Establishing that
+\(\beta_{wn} \ge \beta_{os}\)}{Establishing that \textbackslash beta\_\{wn\} \textbackslash ge \textbackslash beta\_\{os\}}}\label{establishing-that-beta_wn-ge-beta_os}}
 
@@
 So as to simplify the notation of the following section, I will
-introduce a series of new variables. Let \(C = |E_c|\) be the number of
-links that are identical between networks (as a mnemonic, \(C\) stands
-for ``common''); \(R = |E_{sn} \cup E_{sm}|\) be the number of links
-that are not shared, but only involve shared species (\emph{i.e.} links
-from \(\mathcal{M}\cup\mathcal{N}\) established between species from
-\(\mathcal{M}\cap\mathcal{N}\); as a mnemonic, \(R\) stands for
-``rewired''); and \(T = |E_{un} \cup E_{um}|\) the number of links that
-are not shared, and involve at least one unique species (as a mnemonic,
-\(T\) stands for ``turnover'').
+introduce a series of new variables. Let \(A = |E_c|\) be the number of
+links that are identical between networks; \(S = |E_{sn} \cup E_{sm}|\)
+be the number of links that are not shared, but only involve shared
+species (\emph{i.e.} links from \(\mathcal{M}\cup\mathcal{N}\)
+established between species from \(\mathcal{M}\cap\mathcal{N}\)); and
+\(U = |E_{un} \cup E_{um}|\) the number of links that are not shared,
+and involve at least one unique species. Adopting the perspective
+developed in the previous section, wherein networks are sets and the
+measures of \(\beta\)-diversity operates on these sets, highlights the
+conceptual issue in the Fründ (2021) alternative normalization: they are
+using components of the networks that are \emph{not} part of the
+networks being compared.
 
-There are two important points to note here. First, as mentionned
-earlier, the number or proportion of species that are shared is not
-involved in the calculation. Second, the connectance of either network
-is not involved in the calculation. That all links counted in
-\emph{e.g.} \(U\) come from \(\mathcal{M}\), or that they are evenly
-distributed between \(\mathcal{M}\) and \(\mathcal{N}\), has no impact
-on the result. This is a desirable property of the approach: whatever
-quantitative value of the components of dissimilarity can be interpreted
-in the light of the connectance and species turnover \emph{without} any
-risk of circularity; indeed, I present a numerical experiment where
-connectance varies independently later in this manuscript, reinforcing
-this point.
+There are two important points to note here. First, the number or
+proportion of species that are shared is not involved in the
+calculation. Second, the connectance of either network is not involved
+in the calculation. That all links counted in \emph{e.g.} \(U\) come
+from \(\mathcal{M}\), or that they are evenly distributed between
+\(\mathcal{M}\) and \(\mathcal{N}\), has no impact on the result. This
+is a desirable property of the approach: whatever quantitative value of
+the components of dissimilarity can be interpreted in the light of the
+connectance and species turnover \emph{without} any risk of circularity.
+Therefore the argument of Fründ (2021), whereby the \(\beta_{os}\)
+component should decrease with turnover, and be invariant to
+connectance, does not hold: the very point of the approach is to provide
+measures that can be interpreted in the light of connectance and species
+turnover.
 
@@
 (\(\beta_{os}\)), and therefore stems from species turnover. This
-fraction is defined as \(\beta_{st} = \beta_{wn}-\beta_{os}\). The
-expression of \(\beta_{st}\) does not involve a partition into sets that
-can be plugged into the framework of Koleff \emph{et al.} (2003),
+fraction is defined as \(\beta_{st} = \beta_{wn}-\beta_{os}\).
+
+The expression of \(\beta_{st}\) does not involve a partition into sets
+that can be plugged into the framework of Koleff \emph{et al.} (2003),
 because the part of \(\mathcal{M}\) and \(\mathcal{N}\) that are
@@
 
-\[\beta_{os} = \frac{R}{2C+R}\,,\]
+\[\beta_{os} = \frac{S}{2A+S}\,,\]
 
@@
 
-\[\beta_{wn} = \frac{R+T}{2C+R+T}\,.\]
+\[\beta_{wn} = \frac{S+U}{2A+S+U}\,.\]
 
@@
 \(\beta_{wn} \ge \beta_{os}\), and therefore that
-\(0 \ge \beta_{st} \ge 1\); as \(C\), \(T\), and \(R\) are cardinalities
+\(0 \ge \beta_{st} \ge 1\); as \(A\), \(S\), and \(U\) are cardinalities
 of sets, they are necessarily satisfying this condition.
@@
 
-\[\beta_{st} = \frac{2CT}{(2C+R)(2C+R+T)}\,.\]
+\[\beta_{st} = \frac{2AU}{(2A+S)(2A+S+U)}\,.\]
 
@@
 this is obvious when developing the denominator into
-\(4C^2 + R^2 + 4CR + 2CT + RT\). As such, we expect that the value of
-\(\beta_{st}\) will vary in a hump-shaped way with the proportion of
-shared interactions. For this reason, Poisot \emph{et al.} (2012)
-suggest that \(\beta_{st}/\beta{wn}\) (alt. \(1-\beta_{os}/\beta_{wn}\))
-is a better indicator of the \emph{relative} importance of turnover
-processes on network dissimilarity. This can be calculated as
 
-\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2CT}{(2C+S)(2C+R+T)}\times\frac{R+T}{2C+R+T}\,,\]
+\[4A^2 + S^2 + 4AS + 2AU + SU\,,\]
+
+As such, we expect that the value of \(\beta_{st}\) will vary in a
+hump-shaped way with the proportion of shared interactions. For this
+reason, Poisot \emph{et al.} (2012) suggest that
+\(\beta_{st}/\beta{wn}\) (alt. \(1-\beta_{os}/\beta_{wn}\)) is a better
+indicator of the \emph{relative} importance of turnover processes on
+network dissimilarity. This can be calculated as
+
+\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(2A+S+U)}\times\frac{S+U}{2A+S+U}\,,\]
 
@@
 
-\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2CT}{(2C+R)(R+T)}\,.\]
+\[\frac{\beta_{st}}{\beta_{wn}} = \frac{2AU}{(2A+S)(S+U)}\,.\]
 
-The roots of this expression are \(C=0\) (the turnover of species has no
+The roots of this expression are \(A=0\) (the turnover of species has no
 contribution to the difference between \(\beta_{wn}\) and \(\beta_{os}\)
 if there are no shared species, and therefore no rewiring), and for
-\(T = 0\) (the turnover of species has no contribution if all species
+\(U = 0\) (the turnover of species has no contribution if all species
 are shared).
 
-\hypertarget{quantifying-the-response-of-network-beta-diversity-to-souces-of-variation}{%
-\subsection{Quantifying the response of network beta-diversity to souces
-of
-variation}\label{quantifying-the-response-of-network-beta-diversity-to-souces-of-variation}}
+\hypertarget{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}{%
+\subsubsection{Numerical experiment: response of the components to
+different sources of network
+variation}\label{numerical-experiment-response-of-the-components-to-different-sources-of-network-variation}}
 
-\hypertarget{the-relative-effect-of-species-turnover-and-link-rewiring}{%
-\subsubsection{The relative effect of species turnover and link
-rewiring}\label{the-relative-effect-of-species-turnover-and-link-rewiring}}
-
-As the decomposition of beta diversity into sets presented above
-reveals, the value of the components \(\beta_{os}\) and \(\beta_{st}\)
-will respond to two family of mechanisms: the probability of sharing a
-species between the two networks, noted \(p\), which will impose bounds
-on the value of \(T\); and the probability of an interactions between
-shared species \emph{not} being rewired, noted \(q\), which will impose
-bounds on the value of \(C\). These two probabilities represent,
-respectively, mechanisms involved in species turnover and link turnover,
-as per Poisot \emph{et al.} (2015), and the aim of this numerical
-experiment is to describe how these families of processes drive network
-dissimilarity.
-
-In order to simplify the calculations, I make the assumptions that the
-networks have equal species richness (noted \(S\)), so that
-\(S_1 = S_2 = S\), and the same connectance (noted \(\rho\)), so that
-\(\rho_1 = \rho_2 = \rho\). As a consequence, the two networks have the
-same number of links \(L = \rho\times S_1 = \rho\times S_2\). The
-assumption of equal connectance will be relaxed in a subsequent
-numerical experiment. These simplifications allow to express the size of
-\(C\), \(R\), and \(T\) only as functions of \(p\) and \(q\), as they
-would all be multiplied by \(L\), which can therefore be dropped from
-the calculation.
+To illustrate the behavior of \(\beta_{st}\), I conducted a simple
+numerical experiment in which two networks have the same number of
+interactions \(L\) (recall from the previous section that we do not need
+to set a number of species yet), and these interactions are partitionned
+according to proportions \(p_s\) and \(p_r\) into shared (\(A\)),
+rewired (\(S\)), and unique (\(U\)) links, with \(A = p_s\times L\),
+\(S = (1-p_s)\times p_r\times L\), and
+\(U = (1-p_s)\times (1-p_r)\times L\). The results are represented in
+fig.~\ref{fig:numexp1}.
 
@@
 \centering
-\includegraphics{figures/sharing_v_rewiring/components.png}
+\includegraphics{figures/numexp1.png}
 \caption{Values of \(\beta_{os}\), \(\beta_{wn}\), \(\beta_{st}\), and
-\(\beta_{st}/\beta_{wn}\) as a function of the probability \(q\) or
-sharing a link (\(x\)-axis), and the probability \(p\) of sharing a
-species (\(y\)-axis). Larger values indicate \emph{more} dissimilarity,
-such that for \(p=q=1\) the dissimilarity as measured by
-\(\beta_{wn}=0\), and for \(p=q=0\) the dissimilarity as measured by
-\(\beta_{wn}=1\).}\label{fig:numexp1}
+\(\beta_{st}/\beta_{wn}\) as a function of the proportion of rewired
+links and the proportion of shared links.}\label{fig:numexp1}
 }
@@
 
-\hypertarget{sensibility-of-the-decomposition-to-differences-in-connectance}{%
-\subsubsection{Sensibility of the decomposition to differences in
-connectance}\label{sensibility-of-the-decomposition-to-differences-in-connectance}}
+\hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
+\subsection{Is this decomposition over-estimating the effect of
+``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}
+
+One of the arguments put forth by Fründ (2021) is that the decomposition
+outlined above will overestimate the effect of rewiring; I argue that
+this is based on a misunderstanding of what \(\beta_{st}\) achieves. It
+is paramount to clarify that \(\beta_{st}\) is not a direct measure of
+the importance of turnover: it is a quantification of the relative
+impact of rewiring to overall dissimilarity, which, all non-turnover
+mechanisms being accounted for in the decomposition, can be explained by
+turnover mechanisms. In this section, I present two numerical
+experiments showing (i) that the \(\beta_{os}\) component is in fact an
+accurate measure of rewiring, and (ii) that \(\beta_{st}\) captures the
+consequences of species turnover, and of the interactions brought by
+unique species.
+
+\hypertarget{illustrations-on-arbitrarily-small-networks-are-biased}{%
+\subsubsection{Illustrations on arbitrarily small networks are
+biased}\label{illustrations-on-arbitrarily-small-networks-are-biased}}
+
+We can re-calculate the illustration of Fründ (2021), wherein a pair of
+networks with two shared interactions (\(A = 2\)) receive either an
+interaction in \(S\), in \(U\), or in both:
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+\(A\) & \(S\) & \(U\) & \(\beta_{os}\) & \(\beta_{wn}\) & \(\beta_{st}\)
+& \(\beta_{st}/\beta_{wn}\)\tabularnewline
+\midrule
+\endhead
+2 & 0 & 0 & 0 & 0 & 0 &\tabularnewline
+2 & 1 & 0 & \(1/5\) & \(1/5\) & 0 & 0\tabularnewline
+2 & 0 & 1 & 0 & \(1/5\) & \(1/5\) & 0\tabularnewline
+2 & 1 & 1 & \(1/5\) & \(1/3\) & \(2/15\) & \(2/5\)\tabularnewline
+\bottomrule
+\end{longtable}
+
+The over-estimation argument hinges on the fact that
+\(\beta_{st} < \beta_{os}\) in the last situation (one interaction as
+rewiring, one as turnover). Reaching the conclusion of an overestimation
+from this is based on a mis-interpretation of what \(\beta_{st}\) means.
+The correct interpretation is that, out of the entire network
+dissimilarity, only three-fifths are explained by re-wiring. The fact
+that this fraction is not exactly one-half comes from the fact that the
+Wilson \& Shmida (1984) measure counts shared interactions \emph{twice}
+(\emph{i.e.} it has a \(2A\) term), which over-amplifies the effect of
+shared interactions as the network is really small. Running the same
+calculations with \(A = 10\) gives a relative importance of the turnover
+processes of 47\%, and \(\beta_{st}\) goes to \(1/2\) as \(A/(S+U)\)
+increases. As an additional caveat, the value of \(\beta_{st}\) will
+depend on the measure of beta-diversity used. Measures that do not count
+the shared interaction twice are not going to amplify the effect of
+rewiring.
+
+\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}{%
+\subsubsection{Numerical experiment: the decomposition captures the
+roles of rewiring and turnover
+accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-rewiring-and-turnover-accurately}}
 
@@
 denominators that make sense from a network composition point of view,
-succeeds at capturing the relative effect of turnover and rewiring. \#\#
-Does the partition of network dissimilarity needs a new normalization?
-
-\hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
-\subsubsection{Is this decomposition over-estimating the effect of
-``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}
-
-One of the arguments put forth by Fründ (2021) is that the decomposition
-outlined above will overestimate the effect of rewiring; I argue that
-this is based on a misunderstanding of what \(\beta_{st}\) achieves. It
-is paramount to clarify that \(\beta_{st}\) is not a direct measure of
-the importance of turnover: it is a quantification of the relative
-impact of rewiring to overall dissimilarity, which, all non-turnover
-mechanisms being accounted for in the decomposition, can be explained by
-turnover mechanisms. In this section, I present two numerical
-experiments showing (i) that the \(\beta_{os}\) component is in fact an
-accurate measure of rewiring, and (ii) that \(\beta_{st}\) captures the
-consequences of species turnover, and of the interactions brought by
-unique species.
-
-\hypertarget{illustrations-on-arbitrarily-small-networks-are-biased}{%
-\subsubsection{Illustrations on arbitrarily small networks are
-biased}\label{illustrations-on-arbitrarily-small-networks-are-biased}}
-
-We can re-calculate the illustration of Fründ (2021), wherein a pair of
-networks with two shared interactions (\(A = 2\)) receive either an
-interaction in \(S\), in \(U\), or in both:
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-\(A\) & \(S\) & \(U\) & \(\beta_{os}\) & \(\beta_{wn}\) & \(\beta_{st}\)
-& \(\beta_{st}/\beta_{wn}\)\tabularnewline
-\midrule
-\endhead
-2 & 0 & 0 & 0 & 0 & 0 &\tabularnewline
-2 & 1 & 0 & \(1/5\) & \(1/5\) & 0 & 0\tabularnewline
-2 & 0 & 1 & 0 & \(1/5\) & \(1/5\) & 0\tabularnewline
-2 & 1 & 1 & \(1/5\) & \(1/3\) & \(2/15\) & \(2/5\)\tabularnewline
-\bottomrule
-\end{longtable}
-
-The over-estimation argument hinges on the fact that
-\(\beta_{st} < \beta_{os}\) in the last situation (one interaction as
-rewiring, one as turnover). Reaching the conclusion of an overestimation
-from this is based on a mis-interpretation of what \(\beta_{st}\) means.
-The correct interpretation is that, out of the entire network
-dissimilarity, only three-fifths are explained by re-wiring. The fact
-that this fraction is not exactly one-half comes from the fact that the
-Wilson \& Shmida (1984) measure counts shared interactions \emph{twice}
-(\emph{i.e.} it has a \(2A\) term), which over-amplifies the effect of
-shared interactions as the network is really small. Running the same
-calculations with \(A = 10\) gives a relative importance of the turnover
-processes of 47\%, and \(\beta_{st}\) goes to \(1/2\) as \(A/(S+U)\)
-increases. As an additional caveat, the value of \(\beta_{st}\) will
-depend on the measure of beta-diversity used. Measures that do not count
-the shared interaction twice are not going to amplify the effect of
-rewiring.
-
-Based on the arguments presented above, I do not think the suggestion of
-Fründ (2021) to change the denominator of \(\beta_{os}\) makes sense as
-a default; the strength of the original approach by Poisot \emph{et al.}
-(2012) is indeed that the effect of turnover is based on a rigorous
-definition of networks as graphs (as opposed to networks as matrices),
-in which the induction of vertices from the edgelist being compared
-gives rise to biologically meaningful denominators. The advantage of
-this approach is that at no time does the turnover of species itself (or
-indeed, as shown in many places in this manuscript, the network
-richness), or the connectance of the network, enter into the
-calculation. As such, it is possible to use \(\beta_{os}\) and
-\(\beta_{wn}\) in relationship to these terms, calculated externally (as
-was recently done by \emph{e.g.} Higino \& Poisot 2021), without
-creating circularities.
-
-\textbf{TK} Therefore the argument of Fründ (2021), whereby the
-\(\beta_{os}\) component should decrease with turnover, and be invariant
-to connectance, does not hold: the very point of the approach is to
-provide measures that can be interpreted in the light of connectance and
-species turnover.
-
-\textbf{TK} Adopting the perspective developed in the previous section,
-wherein networks are sets and the measures of \(\beta\)-diversity
-operates on these sets, highlights the conceptual issue in the Fründ
-(2021) alternative normalization: they are using components of the
-networks that are \emph{not} part of the networks being compared.
+succeeds at capturing the relative effect of turnover and rewiring.
 
@@
 
-\hypertarget{measuring-network-beta-diversity-recommendations}{%
-\subsection{Measuring network beta-diversity:
-recommendations}\label{measuring-network-beta-diversity-recommendations}}
+\hypertarget{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}{%
+\subsection{Does the partition of network dissimilarity needs a new
+normalization?}\label{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}}
+
+Based on the arguments presented above, I do not think the suggestion of
+Fründ (2021) to change the denominator of \(\beta_{os}\) makes sense as
+a default; the strength of the original approach by Poisot \emph{et al.}
+(2012) is indeed that the effect of turnover is based on a rigorous
+definition of networks as graphs (as opposed to networks as matrices),
+in which the induction of vertices from the edgelist being compared
+gives rise to biologically meaningful denominators. The advantage of
+this approach is that at no time does the turnover of species itself (or
+indeed, as shown in many places in this manuscript, the network
+richness), or the connectance of the network, enter into the
+calculation. As such, it is possible to use \(\beta_{os}\) and
+\(\beta_{wn}\) in relationship to these terms, calculated externally (as
+was recently done by \emph{e.g.} Higino \& Poisot 2021), without
+creating circularities.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Banville2021ManJl}{}%
+Banville, F., Vissault, S. \& Poisot, T. (2021). Mangal.jl and
+EcologicalNetworks.jl: Two complementary packages for analyzing
+ecological networks in Julia. \emph{Journal of Open Source Software}, 6,
+2721.
+
 \leavevmode\hypertarget{ref-Baronio2021NatFir}{}%

### 2021-10-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@08cf85b4d29a6f6a51e60325d9d93a4211aecce3 🚀

@@
 
-We can similarly define the intersection (similarly commutative) of two
+We can similarly define the intersection (also commutative) of two
 networks:
@@
 
-One of the arguments put forth by Fründ (2021) is that the decomposition
-outlined above will overestimate the effect of rewiring; I argue that
-this is based on a misunderstanding of what \(\beta_{st}\) achieves. It
-is paramount to clarify that \(\beta_{st}\) is not a direct measure of
-the importance of turnover: it is a quantification of the relative
-impact of rewiring to overall dissimilarity, which, all non-turnover
-mechanisms being accounted for in the decomposition, can be explained by
-turnover mechanisms. In this section, I present two numerical
-experiments showing (i) that the \(\beta_{os}\) component is in fact an
-accurate measure of rewiring, and (ii) that \(\beta_{st}\) captures the
-consequences of species turnover, and of the interactions brought by
-unique species.
+One of the arguments put forth in a recent paper by Fründ (2021) is that
+the decomposition outlined above will overestimate the effect of
+rewiring; I argue that this is based on a misunderstanding of what
+\(\beta_{st}\) achieves. It is paramount to clarify that \(\beta_{st}\)
+is not a direct measure of the importance of turnover: it is a
+quantification of the relative impact of rewiring to overall
+dissimilarity, which, all non-turnover mechanisms being accounted for in
+the decomposition, can be explained by turnover mechanisms. In this
+section, I present two numerical experiments showing (i) that the
+\(\beta_{os}\) component is in fact an accurate measure of rewiring, and
+(ii) that \(\beta_{st}\) captures the consequences of species turnover,
+and of the interactions brought by unique species.
 
@@
 We can re-calculate the illustration of Fründ (2021), wherein a pair of
-networks with two shared interactions (\(A = 2\)) receive either an
-interaction in \(S\), in \(U\), or in both:
+networks with two shared interactions (\(C = 2\)) receive either an
+interaction in \(T\), in \(R\), or in both:
 
@@
 \toprule
-\(A\) & \(S\) & \(U\) & \(\beta_{os}\) & \(\beta_{wn}\) & \(\beta_{st}\)
+\(C\) & \(T\) & \(R\) & \(\beta_{os}\) & \(\beta_{wn}\) & \(\beta_{st}\)
 & \(\beta_{st}/\beta_{wn}\)\tabularnewline
@@
 Wilson \& Shmida (1984) measure counts shared interactions \emph{twice}
-(\emph{i.e.} it has a \(2A\) term), which over-amplifies the effect of
+(\emph{i.e.} it has a \(2C\) term), which over-amplifies the effect of
 shared interactions as the network is really small. Running the same
-calculations with \(A = 10\) gives a relative importance of the turnover
-processes of 47\%, and \(\beta_{st}\) goes to \(1/2\) as \(A/(S+U)\)
+calculations with \(C = 10\) gives a relative importance of the turnover
+processes of 47\%, and \(\beta_{st}\) goes to \(1/2\) as \(C/(T+R)\)
 increases. As an additional caveat, the value of \(\beta_{st}\) will
@@
 indeed, as shown in many places in this manuscript, the network
-richness), or the connectance of the network, enter into the
-calculation. As such, it is possible to use \(\beta_{os}\) and
-\(\beta_{wn}\) in relationship to these terms, calculated externally (as
-was recently done by \emph{e.g.} Higino \& Poisot 2021), without
-creating circularities.
+richness), or the connectance of the network, enter into the calculation
+of the beta-diversity components. As such, it is possible to use
+\(\beta_{os}\) and \(\beta_{wn}\) in relationship to these terms,
+calculated externally (as was recently done by \emph{e.g.} Higino \&
+Poisot 2021), without creating circularities.
 
-\textbf{TK} Therefore the argument of Fründ (2021), whereby the
-\(\beta_{os}\) component should decrease with turnover, and be invariant
-to connectance, does not hold: the very point of the approach is to
-provide measures that can be interpreted in the light of connectance and
-species turnover.
-
-\textbf{TK} Adopting the perspective developed in the previous section,
+Therefore the argument of Fründ (2021), whereby the \(\beta_{os}\)
+component should decrease with turnover, and be invariant to
+connectance, does not hold: the very point of the approach is to provide
+measures that can be interpreted in the light of connectance and species
+turnover. Adopting the perspective developed in the previous section,
 wherein networks are sets and the measures of \(\beta\)-diversity
 operates on these sets, highlights the conceptual issue in the Fründ
-(2021) alternative normalization: they are using components of the
-networks that are \emph{not} part of the networks being compared.
+(2021) alternative normalization: they are using components (namely,
+interactions) of the networks that are \emph{not} directly part of the
+two networks being compared.
 

### 2021-10-26 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@08cf85b4d29a6f6a51e60325d9d93a4211aecce3 🚀

@@
 indeed, as shown in many places in this manuscript, the network
-richness), or the connectance of the network, enter into the
-calculation. As such, it is possible to use \(\beta_{os}\) and
-\(\beta_{wn}\) in relationship to these terms, calculated externally (as
-was recently done by \emph{e.g.} Higino \& Poisot 2021), without
-creating circularities.
-
-\textbf{TK} Therefore the argument of Fründ (2021), whereby the
-\(\beta_{os}\) component should decrease with turnover, and be invariant
-to connectance, does not hold: the very point of the approach is to
-provide measures that can be interpreted in the light of connectance and
-species turnover.
-
-\textbf{TK} Adopting the perspective developed in the previous section,
+richness), or the connectance of the network, enter into the calculation
+of the beta-diversity components. As such, it is possible to use
+\(\beta_{os}\) and \(\beta_{wn}\) in relationship to these terms,
+calculated externally (as was recently done by \emph{e.g.} Higino \&
+Poisot 2021), without creating circularities.
+
+Therefore the argument of Fründ (2021), whereby the \(\beta_{os}\)
+component should decrease with turnover, and be invariant to
+connectance, does not hold: the very point of the approach is to provide
+measures that can be interpreted in the light of connectance and species
+turnover. Adopting the perspective developed in the previous section,
 wherein networks are sets and the measures of \(\beta\)-diversity
 operates on these sets, highlights the conceptual issue in the Fründ
-(2021) alternative normalization: they are using components of the
-networks that are \emph{not} part of the networks being compared.
+(2021) alternative normalization: they are using components (namely,
+interactions) of the networks that are \emph{not} directly part of the
+two networks being compared.
 

### 2021-10-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@5e43ee994e53d4600e33b8d78c7c7714955c5c2b 🚀

@@
 
-\hypertarget{is-this-decomposition-over-estimating-the-effect-of-rewiring}{%
-\subsubsection{Is this decomposition over-estimating the effect of
-``rewiring?''}\label{is-this-decomposition-over-estimating-the-effect-of-rewiring}}
-
 One of the arguments put forth in a recent paper by Fründ (2021) is that
@@
 
-\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}{%
-\subsubsection{Numerical experiment: the decomposition captures the
-roles of species turnover and connectance
-accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}}
+\hypertarget{using-an-alternative-normalization-trivializes-the-results}{%
+\subsubsection{Using an alternative normalization trivializes the
+results}\label{using-an-alternative-normalization-trivializes-the-results}}
 
-Consider now two bipartite networks, which still have \(R\) species on
-either side, but differ in their connectance (\(\rho_1\) and \(\rho_2\))
--- by maintaining the assumption that species on one side are shared
-with probability \(p\), and that interactions between shared species are
-rewired at probability \(q\), we can examine the effect of varying both
-connectance and turnover on the value of the \(\beta\)-diversity
-components. Note that, although not presented, we will drop the
-multiplicative constant \(R^2\) from all calculations, as it is a common
-factor for all values; again, this implies that the results presented
-here are independant of network richness.
-
-The number of unique links due to species turnover is
-
-\[U = (1-p)(\rho_1 + \rho_2)\,,\]
-
-which decreases with the proportion of shared species, but increases
-with connectance. The number of links between shared species takes a
-little more steps to calculate. First, amongst the \(pR^2\) species in
-both sub-graphs, network 1 will have \(\rho_1 pR^2\), and network 2 will
-have \(\rho_2 pR^2\). Because \(\rho_1 \neq \rho_2\), there are only
-\(\text{min}(\rho_1, \rho_2)pR^2\) links that can be shared, a
-proportion \(q\) of which will undergo re-wiring, and a proportion
-\((1-q)\) of which will be shared. This leads to the expression (after
-dropping \(R^2\)) for the number of shared links:
-
-\[A = p (1-q) \text{min}(\rho_1, \rho_2)\,.\]
-
-The number of unique links due to shared species is the sum of all links
-in network 1 (\(\rho_1 R^2\)), minus the sum of the shared links
-(\(AR^2\)) and the unique links due to species turnover
-(\((1-p)\rho_1R^2\)); this same quantity is calculated in the same way
-for the second networks, leading to (after dropping the multiplicative
-constant \(R^2\) and some simplifications)
-
-\[S = p (\rho_1 + \rho_2) - 2A\,.\]
-
-Note that as expected, this last quantity scales with the proportion of
-shared species (\(p\)) \emph{and} with connectance (as shared species
-bring more of their interactions), but decreases with the size of the
-shared links components. The consequences of varying \(\rho_2\) and
-\(p\) are presented in fig.~\ref{fig:numexp3}.
+In this numerical experiment, we reproduce the results in
+fig.~\ref{fig:turnrew}, but using the alternative normalization
+described above. The results are presented in fig.~\ref{fig:commden}.
+Producing the analytical solutions for the various components, following
+the expressions for \(C\), \(T\), and \(R\) given for
+fig.~\ref{fig:turnrew}, yields a similar value for \(\beta_{wn}\)
+(\emph{i.e.} the two approaches estimate the same value for total
+dissimiliarity), but different values for \(\beta_{st}\) and
+\(\beta_{os}\). Specifically, \(\beta_{os}\) becomes \(p^2(1-q)\), which
+becomes dependent on species turnover. This, from an ecological point of
+view, makes no sense: the quantification of how much shared species
+interact in a similar way should not depend on how much species actually
+overlap. The opposite problem arises for \(\beta_{st}\), which becomes
+\(1-p^2\). In short, the relative importance of species turnover is
+simply species turnover itself, and has no information on interaction
+dissimilarity. Therefore the core issue of the Fründ (2021) alternative
+is that, by attempting to fix a non-issue (namely the over-estimate of
+the importance of re-wiring, which is only true in trivially small
+networks), it blurs the meaning of \(\beta_{os}\), and renders
+\(\beta_{st}\) useless as it is a re-expression of species
+beta-diversity.
 
@@
 \includegraphics{figures/common_denominator/components.png}
-\caption{Effects of varying the connectance of the second network
-(\(\rho_2\)) and the proportion of shared species (\(p\)) on the values
-of the \(\beta\)-diversity components. As expected, \(\beta_{os}\) is
-still independent of species turnover, and \(\beta_{wn}\) increases when
-species turnover increases, or when the connectances become more
-dissimilar. These figures have been generated with \(\rho_1 = 0.25\) and
-\(q = 0.15\), and the results are qualitatively robust to changes in
-these parameters.}\label{fig:commden}
+\caption{Reproduction of fig.~\ref{fig:turnrew} with the alternative
+denominators proposed by Fründ (2021).}\label{fig:commden}
 }
 \end{figure}
-
-Although \(\beta_{os}\) is only responding to changes in connectance (as
-is expected, seeing that the relative connectances of both networks
-appear in the expression for \(S\) and \(A\)), \(\beta_{wn}\) changes in
-response to both parameters. Specifically, increasing the difference in
-connectance between the two networks, especially when also increasing
-the species dissimilarity, results in more dissimilar networks -- this
-is because unique species from both networks bring their own
-interactions (at rate \(\rho_1\) and \(\rho_2\)), and therefore
-contribute to dissimilarity. It is particularly noteworthy that
-\(\beta_{st}\), regardless of the differences in connectance, increases
-with the proportion of unique species. At an equal proportion of shared
-species, \(\beta_{st}\) \emph{decreases} with differences in
-connectance: this is an equally expected result, which indicates that
-the difference between \(\beta_{os}\) and \(\beta_{wn}\) is in part
-explained by non-turnover mechanisms (here, changes in connectance).
-Relying on the \(\beta_{st}/\beta_{wn}\) correction again magnifies this
-effect, without changing their interpretation.
 
@@
 
-The choice of changing the denominator hinges on what one admits as a
-definition for \(\beta_{st}\). If the point of \(\beta_{st}\) is to be a
-component of overall \(\beta\)-diversity as advocated by Fründ (2021)
-and Novotny (2009), a change of numerator \emph{might} be acceptable.
-Nevertheless, this change of numerator contributes to blurring the
-frontier between a measure of interaction dissimilarity and a measure of
-community dissimilarity which starts to add the effect of relative
-richness; this later case warrants a thorough methodological assessment.
-Conversely, if as we argue in Poisot \emph{et al.} (2012),
-\(\beta_{st}\) is to be meant as a \emph{guide} to the interpretation of
-\(\beta_{wn}\) and \(\beta_{os}\), and related to actual measures of
-species turnover and network connectance, one must not change the
-denominator.
-
-It is essential to recognize that there are multiple reasons to
-calculate network dissimilarity, and it is our opinion that the
-arguments levied by Fründ (2021) against the original partition stem
-from a misunderstanding of what it intends to do (and does, indeed, do
-well), not from intrinsic methodological issues in the partition itself.
-Based on the results presented in this contribution, I argue that the
-original partition of network \(\beta\)-diversity from Poisot \emph{et
-al.} (2012) should remain the default.
+Based on the numerical experiments and the derivations presented in this
+paper, we can establish a number of recommendations for the measurement
+and analysis of network dissimilarity. First, \(\beta_{os}\) allows to
+estimate the rate of rewiring, which is an important ecological
+information to have; quantifying it properly can give insights as to how
+networks differ. Second, \(\beta_{st}\) captures both turnover and
+rewiring mechanisms, but its interpretation is easier to accomplish in
+the context of total network dissimilarity, and therefore
+\(\beta_{st}/\beta_{wn}\) should be interpreted more thoroughly.
+Finally, because the alternative denominator from Fründ (2021) removes
+the interesting property of \(\beta_{os}\) (independent estimate of
+rewiring rate), and trivializes the meaning of \(\beta_{st}\) (by
+turning it into species dissimilarity), there seems to be no valid
+reason to use it.
 
@@
 
-\leavevmode\hypertarget{ref-Novotny2009BetDiv}{}%
-Novotny, V. (2009). Beta diversity of plant--insect food webs in
-tropical forests: A conceptual framework. \emph{Insect Conservation and
-Diversity}, 2, 5--9.
-
 \leavevmode\hypertarget{ref-Olsson2021IntPla}{}%

### 2021-10-26 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@5e43ee994e53d4600e33b8d78c7c7714955c5c2b 🚀

@@
 
-\hypertarget{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}{%
-\subsubsection{Numerical experiment: the decomposition captures the
-roles of species turnover and connectance
-accurately}\label{numerical-experiment-the-decomposition-captures-the-roles-of-species-turnover-and-connectance-accurately}}
-
-Consider now two bipartite networks, which still have \(R\) species on
-either side, but differ in their connectance (\(\rho_1\) and \(\rho_2\))
--- by maintaining the assumption that species on one side are shared
-with probability \(p\), and that interactions between shared species are
-rewired at probability \(q\), we can examine the effect of varying both
-connectance and turnover on the value of the \(\beta\)-diversity
-components. Note that, although not presented, we will drop the
-multiplicative constant \(R^2\) from all calculations, as it is a common
-factor for all values; again, this implies that the results presented
-here are independant of network richness.
-
-The number of unique links due to species turnover is
-
-\[U = (1-p)(\rho_1 + \rho_2)\,,\]
-
-which decreases with the proportion of shared species, but increases
-with connectance. The number of links between shared species takes a
-little more steps to calculate. First, amongst the \(pR^2\) species in
-both sub-graphs, network 1 will have \(\rho_1 pR^2\), and network 2 will
-have \(\rho_2 pR^2\). Because \(\rho_1 \neq \rho_2\), there are only
-\(\text{min}(\rho_1, \rho_2)pR^2\) links that can be shared, a
-proportion \(q\) of which will undergo re-wiring, and a proportion
-\((1-q)\) of which will be shared. This leads to the expression (after
-dropping \(R^2\)) for the number of shared links:
-
-\[A = p (1-q) \text{min}(\rho_1, \rho_2)\,.\]
-
-The number of unique links due to shared species is the sum of all links
-in network 1 (\(\rho_1 R^2\)), minus the sum of the shared links
-(\(AR^2\)) and the unique links due to species turnover
-(\((1-p)\rho_1R^2\)); this same quantity is calculated in the same way
-for the second networks, leading to (after dropping the multiplicative
-constant \(R^2\) and some simplifications)
-
-\[S = p (\rho_1 + \rho_2) - 2A\,.\]
-
-Note that as expected, this last quantity scales with the proportion of
-shared species (\(p\)) \emph{and} with connectance (as shared species
-bring more of their interactions), but decreases with the size of the
-shared links components. The consequences of varying \(\rho_2\) and
-\(p\) are presented in fig.~\ref{fig:numexp3}.
+\hypertarget{using-an-alternative-normalization-trivializes-the-results}{%
+\subsubsection{Using an alternative normalization trivializes the
+results}\label{using-an-alternative-normalization-trivializes-the-results}}
+
+In this numerical experiment, we reproduce the results in
+fig.~\ref{fig:turnrew}, but using the alternative normalization
+described above. The results are presented in fig.~\ref{fig:commden}.
+Producing the analytical solutions for the various components, following
+the expressions for \(C\), \(T\), and \(R\) given for
+fig.~\ref{fig:turnrew}, yields a similar value for \(\beta_{wn}\)
+(\emph{i.e.} the two approaches estimate the same value for total
+dissimiliarity), but different values for \(\beta_{st}\) and
+\(\beta_{os}\). Specifically, \(\beta_{os}\) becomes \(p^2(1-q)\), which
+becomes dependent on species turnover. This, from an ecological point of
+view, makes no sense: the quantification of how much shared species
+interact in a similar way should not depend on how much species actually
+overlap. The opposite problem arises for \(\beta_{st}\), which becomes
+\(1-p^2\). In short, the relative importance of species turnover is
+simply species turnover itself, and has no information on interaction
+dissimilarity. Therefore the core issue of the Fründ (2021) alternative
+is that, by attempting to fix a non-issue (namely the over-estimate of
+the importance of re-wiring, which is only true in trivially small
+networks), it blurs the meaning of \(\beta_{os}\), and renders
+\(\beta_{st}\) useless as it is a re-expression of species
+beta-diversity.
 

### 2021-10-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@20a20e3718237e86663379d645b2d33c4128342a 🚀

@@
 (Olsson \emph{et al.} 2021); the impact of environment gradients on
-multi-trophic metacommunities (\textbf{Ohlmann2018MapImp?}); and as a
+multi-trophic metacommunities (Ohlmann \emph{et al.} 2018); and as a
 tool to estimate the sampling completeness of networks (Souza \emph{et
@@
 proxy for mechanisms generating differences in interactions among shared
-species). These calculations are done using \texttt{Symbolics.jl}
-(\textbf{Gowda2021HigSym?}), and subsequently transformed in executable
-code for \emph{Julia} (\textbf{Bezanson2017JulFre?}), used to produce
-the figures.
+species). These calculations are done using \texttt{Symbolics.jl} (Gowda
+\emph{et al.} 2021), and subsequently transformed in executable code for
+\emph{Julia} (\textbf{Bezanson2017JulFre?}), used to produce the
+figures.
 
@@
 
+\leavevmode\hypertarget{ref-Gowda2021HigSym}{}%
+Gowda, S., Ma, Y., Cheli, A., Gwozdz, M., Shah, V.B., Edelman, A.,
+\emph{et al.} (2021). High-performance symbolic-numerics via multiple
+dispatch. \emph{arXiv:2105.03949 {[}cs{]}}.
+
 \leavevmode\hypertarget{ref-Higino2021BetPhy}{}%
@@
 Koleff, P., Gaston, K.J. \& Lennon, J.J. (2003). Measuring beta
-diversity for presence--absence data. \emph{Journal of Animal Ecology},
+diversity for presenceabsence data. \emph{Journal of Animal Ecology},
 72, 367--382.
@@
 blooming of mass-flowering crops. \emph{Ecography}, n/a--n/a.
+
+\leavevmode\hypertarget{ref-Ohlmann2018MapImp}{}%
+Ohlmann, M., Mazel, F., Chalmandrier, L., Bec, S., Coissac, E., Gielly,
+L., \emph{et al.} (2018). Mapping the imprint of biotic interactions on
+\textbackslash beta-diversity. \emph{Ecology Letters}, 21, 1660--1669.
 

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@cbfd5cdbdcfbacffeec53d9f575e8be2dd9348a7 🚀
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@9853410081187b5af0c9187e2a3e047b0cb4c087 🚀

@@
 directly from the dissimilarity of interactions between shared species
-(\(\beta_{os}\)), and a component that cannot (\(\beta_{st}\)).
-Presumably, the value of these components for a pair of networks can
-generate insights about the mechanisms involved in dissimilarity.
+(\(\beta_{os}\)), and a component that cannot (\(\beta_{st}\)). The
+\(\beta_{st}\) component differs slightly from the others, in that it is
+a quantification of the \emph{relative} rewiring to overall
+dissimilarity, and not an \emph{absolute} measure of interaction
+turnover. Presumably, the value of these components for a pair of
+networks can generate insights about the mechanisms involved in
+dissimilarity, when interpreted within the context of species turnover
+and differences in network connectance.
 
@@
 themselves, and with species that are common to both networks
-fig.~\ref{fig:conceptual}. For this reason, it is important to come up
+(fig.~\ref{fig:conceptual}). For this reason, it is important to come up
 with guidelines for the interpretation of this measure, and how to use
@@
 \emph{et al.} 2021), and subsequently transformed in executable code for
-\emph{Julia} (\textbf{Bezanson2017JulFre?}), used to produce the
-figures.
+\emph{Julia} (Bezanson \emph{et al.} 2017), used to produce the figures.
 
@@
 
-So as to simplify the notation of the following section, I will
-introduce a series of new variables. Let \(C = |E_c|\) be the number of
-links that are identical between networks (as a mnemonic, \(C\) stands
-for ``common''); \(R = |E_{sn} \cup E_{sm}|\) be the number of links
-that are not shared, but only involve shared species (\emph{i.e.} links
-from \(\mathcal{M}\cup\mathcal{N}\) established between species from
+Note that this measure is written as \(\beta_t\) for consistency with
+Koleff \emph{et al.} (2003). So as to simplify the notation of the
+following section, I will introduce a series of new variables. Let
+\(C = |E_c|\) be the number of links that are identical between networks
+(as a mnemonic, \(C\) stands for ``common'');
+\(R = |E_{sn} \cup E_{sm}|\) be the number of links that are not shared,
+but only involve shared species (\emph{i.e.} links from
+\(\mathcal{M}\cup\mathcal{N}\) established between species from
 \(\mathcal{M}\cap\mathcal{N}\); as a mnemonic, \(R\) stands for
@@
 is not involved in the calculation. That all links counted in
-\emph{e.g.} \(U\) come from \(\mathcal{M}\), or that they are evenly
+\emph{e.g.} \(T\) come from \(\mathcal{M}\), or that they are evenly
 distributed between \(\mathcal{M}\) and \(\mathcal{N}\), has no impact
@@
 Using the \(\beta_t\) measure of dissimilarity, we can re-write (using
-the notation with \(A\), \(S\), and \(U\))
+the notation with \(R\), \(C\), and \(T\))
 
@@
 
+\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
+Bezanson, J., Edelman, A., Karpinski, S. \& Shah, V. (2017). Julia: A
+Fresh Approach to Numerical Computing. \emph{SIAM Review}, 59, 65--98.
+
 \leavevmode\hypertarget{ref-Campos-Moreno2021ImpInt}{}%
